### 2015-11-07 journal-formats/elsevier-article/template-elsarticle-combined.tex
first deployment

\documentclass[preprint,12pt]{elsarticle}

\usepackage{amssymb}

\usepackage{lineno}

\usepackage{url}

\newcommand{\filenamebase}{drones}

\lefthyphenmin=3
\righthyphenmin=2

\usepackage{graphicx,epsfig,verbatim,enumerate}
\usepackage{amssymb,amsmath}
\usepackage{ifthen}

\usepackage{longtable}

\usepackage{mathtools}

\newboolean{twocolswitch}

\newcommand{\avg}[1]{\left\langle#1\right\rangle}
\newcommand{\tavg}[1]{\langle#1\rangle}

\newcommand{\LavgHa}[2]{\tavg{L_{#1,#2}}}
\newcommand{\LavgHaN}[2]{\tavg{L_{#1,#2}}}
\newcommand{\Lavg}{\tavg{L}}

\newcommand{\sindex}[1]{}
\newcommand{\nindex}[1]{}

\newcommand{\etal}{\textit{et al.}}
\newcommand{\www}[1]{\url{#1}}
\newcommand{\req}[1]{(\ref{#1})}
\newcommand{\Req}[1]{Eq.~(\ref{#1})}

% Lettrines
\usepackage{lettrine}

\newcommand{\yes}{}
\newcommand{\no}{}
\newcommand{\tbf}{\textbf}
\newcommand{\tit}{\textit}

\newcommand{\dee}[1]{\mbox{d}#1}
\newcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pdiffsq}[2]{\frac{\partial^2 #1}{{\partial #2}^2}}
\newcommand{\diff}[2]{\frac{{\rm d}#1}{{\rm d}#2}}
\newcommand{\diffsq}[2]{\frac{{\rm d}^{2}#1}{{\rm d} {#2}^2}}
\newcommand{\tdiff}[2]{\mbox{d} #1/\mbox{d} #2}
\newcommand{\tdiffsq}[2]{\mbox{d}^{2} #1/\mbox{d} {#2}^2}
\newcommand{\tpdiff}[2]{\partial #1/\partial #2}
\newcommand{\tpdiffsq}[2]{\partial^2 #1/\partial {#2}^2}
\newcommand{\postdee}[1]{\,\mbox{d}#1}

\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}

\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}

% \newcommand{\kstar}{d^{\ast}}
% \newcommand{\kstari}{k_i^\ast}
\newcommand{\kstar}{d^\ast}
\newcommand{\kstari}{d_i^\ast}

\newcommand{\dstar}{d^\ast}
\newcommand{\dstari}{d_i^\ast}

\newcommand{\phifix}{{\phi^{\ast}}}
\newcommand{\phifixc}{{\phi_c^{\ast}}}
\newcommand{\phifixb}{{\phi_b^{\ast}}}

\newcommand{\phiiactive}{\phi_{i,t}}

\newcommand{\phiiup}{{\phi_{i,\rm on}}}
\newcommand{\phiidown}{{\phi_{i,\rm off}}}

\newcommand{\phiup}{{\phi_{\rm on}}}
\newcommand{\phidown}{{\phi_{\rm off}}}

\newcommand{\Gfun}{G}

\newcommand{\dstardist}{g}
\newcommand{\dosedist}{f}
\newcommand{\dosedistk}{f^{k\star}}

% chaotic contagion
\newcommand{\edgeinfprob}{\rho}
\newcommand{\nodeinfprob}{\phi}

\newcommand{\avgdegree}{k_{\rm avg}}

\newcommand{\stateA}{S_{0}}
\newcommand{\stateB}{S_{1}}

\newcommand{\lcce}{H}

\newcommand{\effectivecount}{f_{q,\textnormal{exp}}}
\newcommand{\effectivecountq}[1]{f_{#1,\textnormal{exp}}}

\newcommand{\simonrho}{\rho}
\newcommand{\zipfexponent}{\alpha}

\newcommand{\partitionprob}{q}
\newcommand{\partitiontype}[1]{$\partitionprob$$=$$#1$}

\newcommand{\onehalf}{\frac{1}{2}}
\newcommand{\onequarter}{\frac{1}{4}}

\newcommand{\clgamma}{\hat{\zipfexponent}}
\newcommand{\xmin}{r_{\textnormal{max}}}
\newcommand{\kstat}{D}
\newcommand{\clp}{\mbox{$p$-value}}
\newcommand{\sigamma}{1-\simonrho}

\journal{Physica A}

\begin{document}

\begin{frontmatter}

\title{Quantitative patterns in drone wars}

\author{
Javier Garcia-Bernardo
}

\ead{javier.garcia-bernado@uvm.edu}

\address{Department of Computer Science,
  The University of Vermont,
  Burlington, VT 05401.}

\author{
Peter Sheridan Dodds
}

\ead{peter.dodds@uvm.edu}

\address{Department of Mathematics \& Statistics,
  Vermont Complex Systems Center,
  Computational Story Lab,
  \& the Vermont Advanced Computing Core,
  The University of Vermont,
  Burlington, VT 05401.}

\author{
Neil F. Johnson
}

\ead{njohnson@physics.miami.edu}

\address{Department of Physics,
University of Miami,
Coral Gables, FL 33124.
}

\begin{abstract}
  Attacks by drones (i.e., unmanned combat air vehicles) continue to generate heated political and ethical debates. Here we examine instead the quantitative nature of drone attacks, focusing on how their intensity and frequency compares to other forms of human conflict. Instead of the power-law distribution found recently for insurgent and terrorist attacks, the severity of attacks is more akin to lognormal and exponential distributions, suggesting that the dynamics underlying drone attacks lie beyond these other forms of human conflict. Meanwhile the pattern in the timing of attacks is consistent with one side having almost complete control. We show that these novel features can be reproduced and understood using a generative mathematical model in which resource allocation to the dominant side is regulated through a feedback loop.

\end{abstract}

\begin{keyword}

warfare \sep drones \sep terrorism \sep scaling

\PACS 89.65.-s \sep 89.75.Da \sep 89.75.Fb \sep 89.75.-k

\end{keyword}

\end{frontmatter}

\linenumbers

\section*{Introduction}
Dating back to physicist L.~F.~Richardson's pioneering
work nearly 100 years ago~\cite{Richardson194}, the quantitative
analysis of human conflict has attracted research interest from across
the social, biological, economic, mathematical and physical
sciences~\cite{Morgenstern2013,Cederman2003,Clauset2009,Clauset,Bohorquez2009a,Dedeo2011}.
As in other human activities~\cite{Barabasi2005,Gabaix2003}, power
laws have been identified in the severity distribution of individual
attacks in insurgencies and
terrorism~\cite{Johnson2013b,Clauset2009,Clauset,Bohorquez2009a}, and
in the temporal trend in
events~\cite{Johnson2013b,Johnson2011a,Clauset}. These studies found
that across a wide range of insurgent wars in which a relatively small
opponent such as an insurgency (Red Queen~\cite{Johnson2011a}) fights
a larger one such as a state (Blue King~\cite{Johnson2011a}), the
probability distribution for the severity $s$---the number of fatalities---of an event (i.e., clash
or attack) is given by $P(s) \propto s^{-\alpha}$ where $\alpha \sim
2.5$, while the trend in the timing of attacks is given by $\tau_{n}
= \tau_1 n^{-b}$, where $\tau_n$ is the time interval
between events $n$ and $n+1$, $n=1,2,\dots$ and $b$ is the escalation
parameter. When $b=0$, the Blue King and Red Queen are evenly matched,
with both effectively running on the same spot---hence the terminology
surrounding the Red Queen~\cite{Johnson2011a}. When $b\neq 0$, there
is an escalation in the frequency of attacks which can be interpreted
as a relative advantage between the Red Queen and the Blue King~\cite{Johnson2011a}. The power-law finding for the distribution
of event severities is consistent with the Red Queen (i.e., insurgent
force) evolving dynamically as a self-organizing system composed of
cells (i.e., clusters) that sporadically either fragment under the
pressure of the Blue King (e.g., state) or coalesce to create larger
cells, and taking the severity of attacks as proportional to the sizes
of the resulting cells~\cite{Bohorquez2009a}.

This paper examines, for the first time, event patterns in the new
form of human conflict offered by unmanned combat air vehicles
(drones)~\cite{wars}. We focus on Pakistan and Yemen because of their
association with drone strike campaigns, using data from the New
America Foundation and the South Asia Terrorism Portal
databases. The situation of drone wars differs from the typical
situation for insurgencies and terrorism in that the attacks are now
carried out by the Blue King on the Red Queen. Moreover, the
sophistication of the action-at-a-distance technology means that any
delay in the Blue King's next attack is likely to have come from a
constraint within Blue itself (e.g., political opposition) as opposed
to any direct counter-adaptation by the Red Queen. Our findings show that drone attacks tend to deviate from the universal patterns observed in the
severity and timing for insurgencies and terrorism, and instead suggest a new regime in which the Blue
King has almost complete control over the conflict. We develop a
generative model in which the timing of attacks is
determined solely by the resources of the Blue King, but are regulated
by a positive feedback loop due to the Blue King's internal
sociopolitical and economic constraints. We show that this simple
model reproduces the main features of the original data and hence
yields novel insight into the unique nature of drone warfare.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{drawing-2.pdf}  
  \caption{
    \textbf{The severity of drone attacks approximately follows a lognormal distribution.}
    Complementary Cumulative Distribution Function
    (CCDF) of the severity of attacks (blue dots and solid line) and
    best fits to power-law (solid black) and lognormal (dashed green) distributions for drone attacks in Pakistan (A) and Yemen
    (B). The optimal parameters for each distribution are (A)
    Power-law: $\alpha = 4.82$, Log-normal: $\mu = 1.60$ and $\sigma = 0.64$, (B) Power-law: 
    $\alpha = 2.21$, Log-normal: $\mu = 1.65$ and $\sigma = 0.77$.    
    (C-F)  CCDFs of the severity of attacks and best fits to log-normal distributions. (C-D) The attack size is drawn from a 
    normal distribution ($N(\mu,\sigma$) with $\mu$ and $\sigma$ corresponding to (C) the largest value in 100 random numbers drawn from a power-law ($\alpha = 4$) and (D) a random value from a exponential distribution ($\lambda = 5$). (E-F) The attack size is drawn from a normal distribution with $\mu$ and $\sigma^2$ corresponding to (E) the largest value in 100 random numbers drawn from a power-law ($\alpha = 4$) and (F) a random value from a exponential distribution ($\lambda = 5$). The maximum likelihood parameters for the lognormal fits are (C) $\mu = 1.48$ and $\sigma = 0.73$, (D) $\mu = 1.51$ and $\sigma = 0.93$, (E) $\mu = 1.33$ and $\sigma = 0.63$, (F) $\mu = 1.44$ and $\sigma = 0.86$. }
  \label{fig:drones.data}
\end{figure*}

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{drawing-3.pdf}  
  \caption{
    \textbf{The severity of drone attacks approximately follows an exponential distribution.}
    Complementary Cumulative Distribution Function
    (CCDF) of the severity of attacks (blue dots and solid line) and
    best fits to power-law (solid black) and exponential (solid red) distributions for drone attacks in Pakistan (A) and Yemen
    (B). The optimal parameters for each distribution are (A)
    $\lambda = -0.21$ (B) $\lambda = -0.11$.    
    (C-F)  CCDFs of the severity of attacks and best fits to exponential distributions. (C-D) The attack size is drawn from a 
    normal distribution with $\mu$ and $\sigma$ ($\sigma^2$ for D) corresponding to  a random value from a exponential distribution ($\lambda = 5$). (E-F) The attack size is drawn from a normal distribution with $\mu$ and $\sigma$ ($\sigma$ for F) corresponding to the largest value in 100 random numbers drawn from a power-law ($\alpha = 4$). $\lambda$ is measured from the slope of the least squared fit in semi-log scale and corresponds to (C) $\lambda = -0.13$, (D) $\lambda = -0.17$, (E) $\lambda = -0.15$,(F) $\lambda = -0.17$. }
  \label{fig:drones.dataExp}
\end{figure*}

\begin{figure*}[tp!]
  \centering
  \includegraphics[width=\textwidth]{drawing-1.pdf}
  \caption{
    \textbf{The timing between attacks reveals power-law relationships.}
    (A-B) The severity of the attacks (vertical lines,
    left axis) and their escalation parameter $b$ (right axis) are plotted for a moving window of 50 attacks in (A) Pakistan and (B) Yemen. (C) A simple model of the process. The Blue King's
    resources (funding, units, experience, etc.) influence the
    frequency of attacks. Resources are invested and create a
    positive feedback loop. The civilian population and other
    variables influence the strength
    of this positive feedback. The amount of resources
    available corresponds to \textit{A}, the advantage of the
    Blue side. (D) Symmetric plot to (A--B) using data generated by the model. For the severity of attacks, the size was assumed to be drawn from the largest known group, where the group size is distributed as a power-law with $\alpha = 4$. }
  \label{fig:drones.model}
\end{figure*}

\section*{Results and Discussion}
Figs.~\ref{fig:drones.data}A--B shows the complementary cumulative distribution function 
(CCDF) of the severity of drone attacks using the New America Foundation
database. We fit power-law and lognormal 
distributions (dashed green and solid black lines respectively; see methods)
for attacks in Pakistan (Figs.~\ref{fig:drones.data}A) and 
Yemen (Figs.~\ref{fig:drones.data}B).
We find that the severity of the strikes is approximately described by lognormal
distributions, particularly in the case of Pakistan. In the case of Yemen, for which we have far less data, the lognormal is more tentative with the larger events deviating most. This finding of approximate log-normality is
consistent with the notion that a drone has a specific design and targets (predominantly houses and vehicles) and hence a pre-determined order of magnitude of the range of
destruction and likely severity. This contrasts with attacks by terrorist or insurgent clusters whose size and hence lethality crosses multiple scales, yielding scale-free (power-law) severity distributions. In drone attacks, an approximate lognormal distribution can arise through at least two mechanisms: First, the fact that the severity of the attack is the result of many independent processes (e.g., successful reporting, good visibility, compact target group, etc.) will itself produce a lognormal distribution in the attack size. Second, if we take the uncertainty in the casualty number to scale with the target size, this also produce an approximate lognormal distribution for many underlying distributions of target sizes. For example, suppose attacks target the largest known or available Red group, drawn from a power-law distribution. Setting the mean and standard deviation of a zero-truncated normal distribution to this value, then reproduces the approximate lognormal distribution (Figs.~\ref{fig:drones.data}C). Similarly, we can imagine that most attacks target small groups, where the chances of civilian casualties is lower, and that the probability of targeting larger groups decreases exponentially. Setting the mean and standard deviation to a random value from a exponential distribution again yields an approximate lognormal distribution (Figs.~\ref{fig:drones.data}D). The same pattern is recovered if the standard deviation is set to scale with the square root of the mean (Figs.~\ref{fig:drones.data}E--F). We note that further reduction in the uncertainty of group size increases the weight of the underlying distribution: For the case where the largest known group is targeted, this can explain the fat tail observed for the Yemen data.

Although we have choses to focus on fitting lognormal distributions as the alternative to power laws, other distributions can also provide good fits. For example, the data agrees well with an exponential distribution (see Fig.~\ref{fig:drones.dataExp}A--B). Scenarios where the size of the Red groups is exponentially distributed, as is the case if the probability of joining a group is constant and independent of the number of members, would naturally yield exponential distributions (Fig.~\ref{fig:drones.dataExp}C--D). Approximate exponential distributions can also be achieved if the groups are power-law distributed Fig.~\ref{fig:drones.dataExp}E--F). Our purpose is not to identify the best alternative to a power-law distribution, but to show that in contrast with conventional warfare and terrorism, the data does not follow a power-law distribution and hence feedback processes are not present across all scales.

We now turn to the timing of attacks in order to gain insight into the temporal
dynamics of the Blue King-versus-Red Queen activity. Following
previous work~\cite{Johnson2013b}, we plot the time interval between
consecutive attacks $\tau_n$ as a function of the cardinal number of
the attack $n=1,2,3,\dots$. The escalation parameter $b$ is the
exponent of the power-law fit $\tau_n = \tau_1 n^{-b}$, which will be
the slope of the best-fit line on a double logarithmic plot. In the
organizational development literature in which subsequent events are related to production, this is referred to as a development curve
while in the psychology literature, where subsequent events correspond
to completing a certain task, it is referred to as a learning 
curve~\cite{Johnson2011a}. In this sense, the `production' or `completion'
of drone attacks has a natural connection to human activity in these
wider fields. For both Pakistan and Yemen, we find that the
parameter $b$ fails to stabilize around zero 
(Figs.~\ref{fig:drones.model}A--B), which is the expected value in a steady state where both sides are adapting well to the opponent's advances.
Instead, the
drone attacks exhibit a large initial escalation (i.e., large positive
$b$) which then transitions to a de-escalation (i.e., large negative
$b$). Given the difficulty for a Red Queen without air defenses to thwart drone attacks
directly, Figs.~\ref{fig:drones.model}A--B suggest that one side (Blue King) effectively
holds complete control for an extended period of time, and that some
internal constraints then arise within the Blue King entity that
eventually de-escalate drone attacks. This is consistent with the decrease in the escalation rate following the closure of a main drone base in 2011 ~\cite{benjamin2013drone}. Interestingly, there is some evidence of Red adaptation to Blue attacks in a recent report by ``The Bureau of Investigative Journalism", which shows a decrease in Red vehicle usage after 2011 corresponding to a peak in attacks on vehicles \url{http://wherethedronesstrike.com/report/76}. This suggests that the Red Queen may be able to limit the severity of attacks.

Fig.~\ref{fig:drones.model}C shows our simple model for explaining these drone attack patterns. This model is of course over-simplified given the wealth of unknowns, yet we believe it is a plausible first step in explaining the empirical observations. We regard the Blue King as possessing
certain resources, for example experience, units, and funding. These
resources degrade over time if no investment is made in the Blue
King's activity, i.e. if the government does not invest in its own
drone development or information research. We assume that if there exists
investment (i.e., funding, time, etc.) then the available resources increase due to a
positive feedback loop, according to the escalation observed $A \propto n^b$, 
where $A$ corresponds to the advantage of the Blue King over the Red Queen. Similar feedback loops has been proposed
in models of conventional terrorism~\cite{Clauset2012}, and can be
affected by external agents, for example public opinion or budget changes. For simplicity we take the
frequency of attacks as directly proportional to the resource level,
while the severity of the attack is independent of resources.

These minimal features are able to replicate the drone strike data
(Figs.~\ref{fig:drones.data}C--F,Figs.~\ref{fig:drones.data}D, is achieved if the
resources increase as a power-law --- hence this is only sustainable for
short periods of time. A constant $b<0$, corresponding to the
frequency of attacks decreasing continuously, is achieved if the
resources decrease continuously, i.e. when there is little or no
investment. Assuming that each drone acts individually, that the attack
severity varies slowly with the available resources (which is in turn
consistent with some form of adaptation by the Red side) and that an increase in precision
requires significant amounts
of development effort, we are able to recreate approximate
lognormal and exponential distributions for the severity of attacks.

In summary, our analysis reveals and helps explain patterns in the severity and
timing of attacks in drone wars, which themselves represent a new form
of action-at-a-distance human conflict. We have purposely stepped
aside from issues of ethics or technology, choosing instead to focus
on the event data since they represent a quantitative measure of drone
war impact. We have shown that a simple model, in which the production of
drones evolves from a shared pool of resources controlled by a
feedback loop, is able to recreate the original data and therefore
explain the overall dynamics of the Blue King's drone campaign.
Going forward, our model could be also used to 
explore how wars might unfold when drones are used 
by two or more sides in conflicts.

\section*{Methods}
We obtained all data from the New America database:
\url{http://securitydata.newamerica.net/.newamerica.net/} and crosschecked with the South Asia Terrorism Portal
database:
\url{http://www.satp.org/satporgtp/countries/pakistan/}.

We obtained the best fit to power-law distributions following Clauset et
al.~\cite{Clauset2009}. We fitted lognormal distributions using the
maximum likelihood estimators. For the escalation rate analysis, $\tau_n
= \tau_1 n^{-b}$, we plotted the number of attack vs. the time between
attacks on a log-log scale. We used a rolling window of 50 attacks and accepted every value of $b$ that
allowed for a correlation greater than 20\%, which allows us to
measure fast transitions.

We simulated 200 attacks with our
model. The initial advantage was set to $1$. The time to the
next attack is equal to $323 \cdot A^{-1}$, where the 323 mimics $t_0$ for the Pakistan conflict. At every step (attack) the advantage
of the Blue King changed by the factor 
$\left((n+1)/n\right)^b$, where $n$ is the attack number.
For the first 50 attacks 
$b = 0.5$; 
for the last 50 attacks 
$b = -0.5$.

\section*{Acknowledgments}
PSD was supported by NSF CAREER Grant No. 0846668.

\bibliographystyle{elsarticle-num} 
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\expandafter\ifx\csname href\endcsname\relax
  \def\href#1#2{#2} \def\path#1{#1}\fi

\bibitem{Richardson194}
L.~F. Richardson, {Variation of the frequency of fatal quarrels with
  magnitude}, Journal of the American Statistical Association 43~(224) (1948)
  523--546.

\bibitem{Morgenstern2013}
A.~P. Morgenstern, N.~Velasquez, P.~Manrique, Q.~Hong, N.~Johnson, N.~Johnson,
  {Resource Letter MPCVW-1: Modeling Political Conflict, Violence, and Wars: A
  Survey}, American Journal of Physics 81~(11) (2013) 805.

\bibitem{Cederman2003}
L.-E. Cederman, {Modeling the Size of Wars: From Billiard Balls to Sandpiles},
  American Political Science Review 97~(01) (2003) 135.

\bibitem{Clauset2009}
A.~Clauset, C.~R. Shalizi, M.~E.~J. Newman, {Power-law distributions in
  empirical data}, SIAM Review 51~(4) (2009) 661--703.

\bibitem{Clauset}
A.~Clauset, M.~Young, K.~Gleditsch, {On the frequency of severe terrorist
  events}, Journal of Conflict Resolution 51~(1) (2007) 58--88.

\bibitem{Bohorquez2009a}
J.~C. Bohorquez, S.~Gourley, A.~R. Dixon, M.~Spagat, N.~F. Johnson, Common
  ecology quantifies human insurgency, Nature 462 (2009) 911--914.

\bibitem{Dedeo2011}
S.~Dedeo, D.~Krakauer, J.~Flack, {Evidence of strategic periodicities in
  collective conflict dynamics.}, Journal of the Royal Society Interface 8~(62)
  (2011) 1260--73.

\bibitem{Barabasi2005}
A.-L. Barab\'{a}si, {The origin of bursts and heavy tails in human dynamics.},
  Nature 435~(7039) (2005) 207--11.

\bibitem{Gabaix2003}
X.~Gabaix, P.~Gopikrishnan, V.~Plerou, H.~E. Stanley, {A theory of power-law
  distributions in financial market fluctuations.}, Nature 423~(6937) (2003)
  267--70.

\bibitem{Johnson2013b}
N.~F. Johnson, P.~Medina, G.~Zhao, D.~S. Messinger, J.~Horgan, P.~Gill, J.~C.
  Bohorquez, W.~Mattson, D.~Gangi, H.~Qi, P.~Manrique, N.~Velasquez,
  A.~Morgenstern, E.~Restrepo, N.~Johnson, M.~Spagat, R.~Zarama, {Simple
  mathematical law benchmarks human confrontations.}, Scientific Reports 3
  (2013) 3463.

\bibitem{Johnson2011a}
N.~Johnson, S.~Carran, J.~Botner, K.~Fontaine, N.~Laxague, P.~Nuetzel,
  J.~Turnley, B.~Tivnan, {Pattern in escalations in insurgent and terrorist
  activity.}, Science Magazine 333~(6038) (2011) 81--4.

\bibitem{wars}
A.~R. Hall, C.~J. Coyne, {The political economy of drones}, Defence and Peace
  Economics 25~(5) (2013) 445--460.

\bibitem{benjamin2013drone}
M.~Benjamin, Drone warfare: Killing by remote control, Verso Books, 2013.

\bibitem{Clauset2012}
A.~Clauset, K.~S. Gleditsch, {The developmental dynamics of terrorist
  organizations.}, PloS ONE 7~(11) (2012) e48633.

\end{thebibliography}

\end{document}

### 2015-11-07 journal-formats/journal-of-the-royal-society-interface/drones-royal-society.tex
first deployment

% To compile:
% latex drones
% bibtex drones
% latex drones
% latex drones
% dvipdf drones
\documentclass[12pt]{article}

\usepackage{hyperref}
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

\usepackage[T1]{fontenc}

\usepackage[nofiglist,nofighead,nomarkers]{endfloat}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{sidecap}
\usepackage[pdftex]{graphicx}
\usepackage{caption}
\usepackage{subcaption}
% \usepackage{cite}
\usepackage{color} 
\usepackage{soul}
\usepackage[super,comma,sort&compress]{natbib}
\usepackage{url}
\urlstyle{rm} % typeset urls in roman instead of in typewriter style
\makeatletter
\renewcommand\@biblabel[1]{#1} %from [1] to 1
\makeatother

\usepackage[margin=21mm]{geometry}
\usepackage[labelfont=bf,font=small,labelsep=period,justification=raggedright]{caption}
%% \bibliographystyle{./vancouver}
\bibliographystyle{unsrtabbrv}

\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\pagestyle{myheadings}
\begin{document}

\begin{flushleft}
{\huge
\textbf{\input{drones.title}}
}\smallbreak

Javier Garcia-Bernardo\textsuperscript{1}, 
Peter Sheridan Dodds\textsuperscript{2}, 
Neil F. Johnson\textsuperscript{3}
\\
\textsuperscript{1}Department of Computer Science, University of Vermont, Burlington VT 05405, USA \\
\textsuperscript{2}Computational Story Lab, Department of Mathematics and Statistics, Vermont Complex Systems Center \& the Vermont Advanced Computing Core, University of Vermont, Burlington VT 05405, USA \\
\textsuperscript{3}Department of Physics, University of Miami, Coral Gables, FL 33124, USA
\end{flushleft}

\section*{Abstract}
\input{drones.abs}

\noindent \textbf{Keywords}: Drones, escalation, conflict, patterns

\input{drones.body}

\bibliography{drones}

\section*{Acknowledgements}
\input{drones.acknowledgments}

\bigskip
\bigskip

\noindent
\textbf{Short title:} \input{drones.title}

\end{document}

### 2015-11-07 paper-template-revtex4-combined.tex
first deployment

\documentclass[pre,twocolumn,twoside,byrevtex,superscriptaddress,floatfix]{revtex4-1}

\usepackage{currfile}
\lefthyphenmin=3
\righthyphenmin=2

\usepackage{graphicx,epsfig,verbatim,enumerate}
\usepackage{amssymb,amsmath}
\usepackage{ifthen}

\usepackage{longtable}

\usepackage{mathtools}

\newboolean{twocolswitch}

\newcommand{\avg}[1]{\left\langle#1\right\rangle}
\newcommand{\tavg}[1]{\langle#1\rangle}

\newcommand{\LavgHa}[2]{\tavg{L_{#1,#2}}}
\newcommand{\LavgHaN}[2]{\tavg{L_{#1,#2}}}
\newcommand{\Lavg}{\tavg{L}}

\newcommand{\sindex}[1]{}
\newcommand{\nindex}[1]{}

\newcommand{\etal}{\textit{et al.}}
\newcommand{\www}[1]{\url{#1}}
\newcommand{\req}[1]{(\ref{#1})}
\newcommand{\Req}[1]{Eq.~(\ref{#1})}

% Lettrines
\usepackage{lettrine}

\newcommand{\yes}{}
\newcommand{\no}{}
\newcommand{\tbf}{\textbf}
\newcommand{\tit}{\textit}

\newcommand{\dee}[1]{\mbox{d}#1}
\newcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pdiffsq}[2]{\frac{\partial^2 #1}{{\partial #2}^2}}
\newcommand{\diff}[2]{\frac{{\rm d}#1}{{\rm d}#2}}
\newcommand{\diffsq}[2]{\frac{{\rm d}^{2}#1}{{\rm d} {#2}^2}}
\newcommand{\tdiff}[2]{\mbox{d} #1/\mbox{d} #2}
\newcommand{\tdiffsq}[2]{\mbox{d}^{2} #1/\mbox{d} {#2}^2}
\newcommand{\tpdiff}[2]{\partial #1/\partial #2}
\newcommand{\tpdiffsq}[2]{\partial^2 #1/\partial {#2}^2}
\newcommand{\postdee}[1]{\,\mbox{d}#1}

\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}

\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}

% \newcommand{\kstar}{d^{\ast}}
% \newcommand{\kstari}{k_i^\ast}
\newcommand{\kstar}{d^\ast}
\newcommand{\kstari}{d_i^\ast}

\newcommand{\dstar}{d^\ast}
\newcommand{\dstari}{d_i^\ast}

\newcommand{\phifix}{{\phi^{\ast}}}
\newcommand{\phifixc}{{\phi_c^{\ast}}}
\newcommand{\phifixb}{{\phi_b^{\ast}}}

\newcommand{\phiiactive}{\phi_{i,t}}

\newcommand{\phiiup}{{\phi_{i,\rm on}}}
\newcommand{\phiidown}{{\phi_{i,\rm off}}}

\newcommand{\phiup}{{\phi_{\rm on}}}
\newcommand{\phidown}{{\phi_{\rm off}}}

\newcommand{\Gfun}{G}

\newcommand{\dstardist}{g}
\newcommand{\dosedist}{f}
\newcommand{\dosedistk}{f^{k\star}}

% chaotic contagion
\newcommand{\edgeinfprob}{\rho}
\newcommand{\nodeinfprob}{\phi}

\newcommand{\avgdegree}{k_{\rm avg}}

\newcommand{\stateA}{S_{0}}
\newcommand{\stateB}{S_{1}}

\newcommand{\lcce}{H}

\newcommand{\effectivecount}{f_{q,\textnormal{exp}}}
\newcommand{\effectivecountq}[1]{f_{#1,\textnormal{exp}}}

\newcommand{\simonrho}{\rho}
\newcommand{\zipfexponent}{\alpha}

\newcommand{\partitionprob}{q}
\newcommand{\partitiontype}[1]{$\partitionprob$$=$$#1$}

\newcommand{\onehalf}{\frac{1}{2}}
\newcommand{\onequarter}{\frac{1}{4}}

\newcommand{\clgamma}{\hat{\zipfexponent}}
\newcommand{\xmin}{r_{\textnormal{max}}}
\newcommand{\kstat}{D}
\newcommand{\clp}{\mbox{$p$-value}}
\newcommand{\sigamma}{1-\simonrho}

\setboolean{twocolswitch}{true}

\begin{document}

\title{\protect
Quantitative patterns in drone wars
}

\author{
\firstname{Javier}
\surname{Garcia-Bernardo}
}
\email{javier.garcia-bernado@uvm.edu}

\affiliation{Department of Computer Science,
  The University of Vermont,
  Burlington, VT 05401.}

\author{
\firstname{Peter Sheridan}
\surname{Dodds}
}
\email{peter.dodds@uvm.edu}

\affiliation{Department of Mathematics \& Statistics,
  Vermont Complex Systems Center,
  Computational Story Lab,
  \& the Vermont Advanced Computing Core,
  The University of Vermont,
  Burlington, VT 05401.}

\author{
\firstname{Neil F.}
\surname{Johnson}
}
\email{njohnson@physics.miami.edu}

\affiliation{Department of Physics,
University of Miami
Coral Gables, FL 33124.
}

\date{\today}

\begin{abstract}
  \protect
  Attacks by drones (i.e., unmanned combat air vehicles) continue to
generate heated political and ethical debates. Here we examine instead
the quantitative nature of drone attacks, focusing on how their
intensity and frequency compares to other forms of human conflict.
In contrast to the power-law distribution found recently for insurgent
and terrorist attacks, we find that the severity of attacks is well
described by a lognormal distribution, indicating that the dynamics
underlying drone attacks differ from these other forms of human
conflict.
Likewise, the pattern in the timing of attacks is consistent with one
side having almost complete control.
We show that these novel features can be reproduced and understood
using a generative mathematical model in which resource allocation to
the dominant side is regulated through a feedback loop.

\end{abstract}

\pacs{89.65.-s,89.75.Da,89.75.Fb,89.75.-k}

\maketitle

\section*{Introduction}
Dating back to physicist L.~F.~Richardson's pioneering
work nearly 100 years ago~\cite{Richardson194}, the quantitative
analysis of human conflict has attracted research interest from across
the social, biological, economic, mathematical and physical
sciences~\cite{Morgenstern2013,Cederman2003,Clauset2009,Clauset,Bohorquez2009a,Dedeo2011}.
As in other human activities~\cite{Barabasi2005,Gabaix2003}, power
laws have been identified in the severity distribution of individual
attacks in insurgencies and
terrorism~\cite{Johnson2013b,Clauset2009,Clauset,Bohorquez2009a}, and
in the temporal trend in
events~\cite{Johnson2013b,Johnson2011a,Clauset}.  These studies found
that across a wide range of insurgent wars in which a relatively small
opponent such as an insurgency (Red Queen~\cite{Johnson2011a}) fights
a larger one such as a state (Blue King~\cite{Johnson2011a}), the
probability distribution for the severity $s$---the number of fatalities---of an event (i.e., clash
or attack) is given by $P(s) \propto s^{-\alpha}$ where $\alpha \sim
2.5$, while the trend in the timing of attacks is given by $\tau_{n}
= \tau_1 n^{-b}$\textcolor{red}{,} where $\tau_n$ is the time interval
between events $n$ and $n+1$, $n=1,2,\dots$ and $b$ is the escalation
parameter. When $b=0$, the Blue King and Red Queen are evenly matched,
with both effectively running on the same spot---hence the terminology
surrounding the Red Queen~\cite{Johnson2011a}. When $b\neq 0$, there
is an escalation in the frequency of attacks which can be interpreted
as a relative advantage between the Blue King and the Red
Queen~\cite{Johnson2011a}. The power-law finding for the distribution
of event severities is consistent with the Red Queen (i.e., insurgent
force) evolving dynamically as a self-organizing system composed of
cells (i.e., clusters) that sporadically either fragment under the
pressure of the Blue King (e.g., state) or coalesce to create larger
cells, and taking the severity of attacks as proportional to the sizes
of the resulting cells~\cite{Bohorquez2009a}.

This paper examines, for the first time, event patterns in the new
form of human conflict offered by unmanned combat air vehicles
(drones)~\cite{wars}. We focus on Pakistan and Yemen because of their
association with drone strike campaigns, using data from the New
America Foundation (NATSEC) and the South Asia Terrorism Portal (SATP)
databases. The situation of drone wars differs from the typical
situation for insurgencies and terrorism in that the attacks are now
carried out by the Blue King on the Red Queen. Moreover, the
sophistication of the action-at-a-distance technology means that any
delay in the Blue King's next attack is likely to have come from a
constraint within Blue itself (e.g., political opposition) as opposed
to any direct counter-adaptation by the Red Queen. Indeed, we find
that the drone attacks do not follow the universal patterns in the
severity and timing for insurgencies and terrorism. Instead our
analysis indicates a new regime of human conflict in which the Blue
King has almost complete control over the conflict. We develop a
generative model in which the severity and timing of attacks are
determined solely by the resources of the Blue King, but are regulated
by a positive feedback loop due to the Blue King's internal
sociopolitical and economic constraints. We show that this simple
model reproduces the main features of the original data and hence
yields novel insight into the nature of drone warfare.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{drawing-2.pdf}  
  \caption{
    \textbf{The severity of drone attacks approximately follows a lognormal
      distribution.}
    Complementary Cumulative Distribution Function
    (CCDF) of the severity of attacks (blue dots and solid line) and
    best fits (dashed line) to power-law (A, C) and lognormal (B,
    D) distributions for drone attacks in Pakistan (A, B) and Yemen
    (C, D). The optimal parameters for each distribution are (A)
    $\alpha = 4.82$, (B) $\mu = 1.60$, $\sigma = 0.64$, (C)
    $\alpha = 2.21$ and (D) $\mu = 1.65$, $\sigma = 0.77$.
    Bottom panels (E, F) show the severity of attack (vertical lines,
    left axis) and escalation parameter $b$ (right axis) for a
    moving window of 50 attacks in Pakistan (E) and Yemen (F).
  }
  \label{fig:drones.data}
\end{figure*}

\begin{figure*}[tp!]
  \centering
  \includegraphics[width=\textwidth]{drawing-1.pdf}
  \caption{
    \textbf{A simple model of drone wars.}
    (A) The Blue King's
    resources (funding, units, experience, etc.) influence the
    frequency and the magnitude of attacks. Resources are
    degraded continuously, with some being invested to create a
    positive feedback loop. The civilian population and other
    variables, summarized as \textit{I}, influence the strength
    of this positive feedback. The dashed line indicates that
    the Red Queen (i.e., target group) is only able to affect the
    Blue King's attacks indirectly. The amount of resources
    available corresponds to \textit{A}, the advantage of the
    Blue side.
    (B) Our simple model has one unstable fixed point at
    \textit{A} = 0, given that $\alpha \cdot I > \lambda$. (C) A
    `resistance' is added to the model due to internal Blue King
    constraints. The model now has a new fixed point at $A =
    \frac{\alpha I - \lambda}{\beta}$. (D-F) Our model is able
    to replicate the observed findings. Complementary Cumulative
    Distribution Function (CCDF) for event severity (blue dots
    and solid line) and the best fits (dashed line) to power-law
    (D, $\alpha = 2.7519$) and lognormal (E, $\mu=1.7$,
    $\sigma=0.83$) distributions for the model. (F) Severity of
    attack (vertical lines, left axis) and escalation parameter
    $b$ for a moving window of 50 attacks (right axis) for the
    model. Note the resemblance to Fig.~\ref{fig:drones.data}.
  }
  \label{fig:drones.model}
\end{figure*}

\section*{Results and Discussion}
Figs.~\ref{fig:drones.data}A--D show the complementary cumulative distribution function
(CCDF) of the severity and frequency of drone attacks using the NATSEC
database. We fit power-law and lognormal 
distributions (dashed green and solid black lines respectively; see methods)
for attacks in Pakistan (Figs.~\ref{fig:drones.data}A--B) and 
Yemen (Figs.~\ref{fig:drones.data}C--D).
We find
that the severity of the strikes is approximately described by lognormal
distributions,
in contrast to that expected for attacks carried out by terrorist and
insurgent groups which follow a power-law distribution reflecting
their underlying cluster sizes~\cite{Johnson}.
Our observation is most well supported for Pakistan,
and tentative for Yemen for which we have far less data,
and where the best fit lognormal 
is less sympathetic with larger events.
This finding of lognormality is
consistent with the notion that a drone has a specific design which
tends to pre-determine the order of magnitude of its range of
destruction and likely severity, in sharp contrast to attacks by a
terrorist or insurgent cluster where the severity of attack can vary
greatly according to the size of the cluster carrying out that
attack. For drone attacks, a lognormal distribution will then arise
naturally from a sum of normal distributions with variable mean and
standard deviation, which in turn reflects variations away from the
pre-determined impact due to differences in the terrain and the number
of people in the target group.

We now turn to
the timing of attacks in order to gain insight into the temporal
dynamics of the Blue King-versus-Red Queen activity. Following
previous work~\cite{Johnson2013b}, we plot the time interval between
consecutive attacks $\tau_n$ as a function of the cardinal number of
the attack $n=1,2,3,\dots$. The escalation parameter $b$ is the
exponent of the power-law fit $\tau_n = \tau_1 n^{-b}$, which will be
the slope of the best-fit line on a double logarithmic plot. In the
organizational development literature in which progressive events are
related to production, this is referred to as a development curve
while in the psychology literature, where subsequent events correspond
to completing a certain task, it is referred to as a learning 
curve~\cite{Johnson2011a}---in this sense, the `production' or `completion'
of drone attacks has a natural connection to human activity in these
wider fields.

For both Pakistan and Yemen, we find that the
parameter $b$ fails to stabilize around 0 
(Figs.~\ref{fig:drones.data}E--F).
Instead, the
drone attacks exhibit an initial large escalation (i.e., large positive
$b$) which then transitions to a de-escalation (i.e., large negative
$b$). Given the difficulty for the Red Queen to thwart drone attacks
directly, Figs.~\ref{fig:drones.data}E--F suggest that one side (Blue King) effectively
holds complete control for an extended period of time, and that some
internal constraints then arise within the Blue King entity that
eventually de-escalate the drone attacks.

Fig.~\ref{fig:drones.model}A shows our
simple model for drone attacks. In our model, the Blue King possesses
certain resources, for example experience, units, and funding. These
resources degrade over time if no investment is made in the Blue
King's activity, i.e., if the government does not invest in its own
drone development or information research. We assume that if the
investment (i.e., funding, time, etc.) is higher than the degradation
rate $\lambda$, then the available resources increase due to a
positive feedback loop (Fig.~\ref{fig:drones.model}B). This feedback loop has been proposed
in models of conventional terrorism~\cite{Clauset2012}, and can be
affected by external agents, for example public opinion or by the
perceived number of Red Queen cells remaining that need targeting. The
resources available at any particular time will affect the timing and
severity of the next drone attack. For simplicity we take the
frequency of attacks as directly proportional to the resource level,
while the severity of the attack depends on the available resources
and the characteristic of the terrorist cells (see methods). We add
`resistance' to the model to mimic constraints within the Blue King,
for example due to political opposition, and hence obtain the logistic
equation which has a new stable equilibrium at a non-zero value, as
shown in Fig.~\ref{fig:drones.model}C.

This simple model is able to replicate the drone strike data
(Figs.~\ref{fig:drones.model}D--F). The escalation parameter $b$ is proportional to
$(\alpha\cdot I - {\lambda})$, which means that the Blue King has a
positive advantage if the investment term is larger than the
degradation term (Fig.~\ref{fig:drones.model}C). A constant $b>0$, corresponding to the
frequency of attacks increasing continuously, is achieved if the
resources increase exponentially---hence this is only sustainable for
short periods of time. A constant $b<0$, corresponding to the
frequency of attacks decreasing continuously, is achieved if the
resources decrease exponentially, i.e., when there is little or no
investment. Assuming each drone acts individually and the attack
severity varies slowly with the available resources, which is
consistent with increases in precision requiring significant amounts
of development effort (see methods), we are able to recreate the
lognormal distribution for the severity of attacks.

In summary, our analysis reveals new patterns in the severity and
timing of attacks in drone wars, which themselves represent a new form
of action-at-a-distance human conflict. We have purposely stepped
aside from issues of ethics or technology, choosing instead to focus
on the event data since they represent a quantitative measure of drone
war impact. We have shown that a simple model, in which the production of
drones evolves from a shared pool of resources controlled by a
feedback loop, is able to recreate the original data and therefore
explain the overall dynamics of the Blue King's drone campaign.
Going forward, our model could be also used to 
explore how wars might 
unfold when drones are used 
by two or more sides in conflicts.

\section*{Methods}
We obtained all data from the NATSEC database:
\url{http://natsec.newamerica.net/} and crosschecked with the SATP
database:
\url{http://www.satp.org/satporgtp/countries/pakistan/}.

We obtained the best fit to power-law distributions following Clauset et
al.~\cite{Clauset2009}. We fitted lognormal distributions using the
maximum likelihood estimators. For the learning rate analysis, $\tau_n
= \tau_1 n^{-b}$, we plotted the number of attack vs. the time between
attacks on a log-log scale. We accepted every value of $b$ that
allowed for a correlation greater than 20\%, which allows us to
measure fast transitions.

We simulated 100 attacks with our
model. The initial advantage was set to $10^{-3}$. The time to the
next attack is equal to $A^{-1}$. At every step (attack) the advantage
of the Blue King changed by the factor 
$\alpha\cdot I - {\lambda}$.
For the first 50 attacks 
$\lambda = \alpha\cdot I = 1.075$; 
for the last 50 attacks 
$\lambda = 1.075$ 
and 
$\alpha\cdot I = 0$.
We model the attacks as individual drones strikes; each individual
attack was drawn from a zero-truncated normal distribution, whose mean
and standard deviation were set to be equal to the root square of the
largest known Red group. The area of a self-avoiding random walk in
two dimensions increases as $A \propto n^{1/2}$, where n is the number
of individuals in the plane~\cite{Havlin1982}; the magnitude of the
attacks is fixed, hence the number of casualties scales with
$n^{1/2}$. The number of known Red groups was set to $(100+10^4\cdot
A) = [110,500]$ and they were assumed to be power-law distributed with
$\alpha =2.5$~\cite{Bohorquez2009a,Richardson194}.

\acknowledgments
We are grateful for funding from the Vermont Complex
Systems Center.
PSD was supported by NSF CAREER Grant No. 0846668.

%merlin.mbs apsrev4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (-1) disabled
%Control: page (0) single
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{15}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%
\providecommand \@@href[1]{\endgroup#1\@@endlink}%
\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
\providecommand \@@startlink[1]{}%
\providecommand \@@endlink[0]{}%
\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
\providecommand \urlprefix  [0]{URL }%
\providecommand \Eprint [0]{\href }%
\providecommand \doibase [0]{http://dx.doi.org/}%
\providecommand \selectlanguage [0]{\@gobble}%
\providecommand \bibinfo  [0]{\@secondoftwo}%
\providecommand \bibfield  [0]{\@secondoftwo}%
\providecommand \translation [1]{[#1]}%
\providecommand \BibitemOpen [0]{}%
\providecommand \bibitemStop [0]{}%
\providecommand \bibitemNoStop [0]{.\EOS\space}%
\providecommand \EOS [0]{\spacefactor3000\relax}%
\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
\let\auto@bib@innerbib\@empty
%</preamble>
\bibitem [{\citenamefont {Richardson}(1948)}]{Richardson194}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.~F.}\ \bibnamefont
  {Richardson}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Journal of the American Statistical Association}\ }\textbf {\bibinfo
  {volume} {43}},\ \bibinfo {pages} {523} (\bibinfo {year} {1948})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Morgenstern}\ \emph {et~al.}(2013)\citenamefont
  {Morgenstern}, \citenamefont {Velasquez}, \citenamefont {Manrique},
  \citenamefont {Hong}, \citenamefont {Johnson},\ and\ \citenamefont
  {Johnson}}]{Morgenstern2013}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~P.}\ \bibnamefont
  {Morgenstern}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
  {Velasquez}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Manrique}},
  \bibinfo {author} {\bibfnamefont {Q.}~\bibnamefont {Hong}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Johnson}}, \ and\ \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Johnson}},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {American Journal of Physics}\ }\textbf
  {\bibinfo {volume} {81}},\ \bibinfo {pages} {805} (\bibinfo {year}
  {2013})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Cederman}(2003)}]{Cederman2003}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.-E.}\ \bibnamefont
  {Cederman}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {American Political Science Review}\ }\textbf {\bibinfo {volume} {97}},\
  \bibinfo {pages} {135} (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Clauset}\ \emph {et~al.}(2009)\citenamefont
  {Clauset}, \citenamefont {Shalizi},\ and\ \citenamefont
  {Newman}}]{Clauset2009}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Clauset}}, \bibinfo {author} {\bibfnamefont {C.~R.}\ \bibnamefont
  {Shalizi}}, \ and\ \bibinfo {author} {\bibfnamefont {M.~E.~J.}\ \bibnamefont
  {Newman}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {SIAM
  Review}\ }\textbf {\bibinfo {volume} {51}},\ \bibinfo {pages} {661} (\bibinfo
  {year} {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Clauset}\ \emph {et~al.}(2007)\citenamefont
  {Clauset}, \citenamefont {Young},\ and\ \citenamefont {Gleditsch}}]{Clauset}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Clauset}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Young}}, \
  and\ \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Gleditsch}},\
  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Journal of
  Conflict Resolution}\ }\textbf {\bibinfo {volume} {51}},\ \bibinfo {pages}
  {58} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Bohorquez}\ \emph {et~al.}(2009)\citenamefont
  {Bohorquez}, \citenamefont {Gourley}, \citenamefont {Dixon}, \citenamefont
  {Spagat},\ and\ \citenamefont {Johnson}}]{Bohorquez2009a}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~C.}\ \bibnamefont
  {Bohorquez}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Gourley}},
  \bibinfo {author} {\bibfnamefont {A.~R.}\ \bibnamefont {Dixon}}, \bibinfo
  {author} {\bibfnamefont {M.}~\bibnamefont {Spagat}}, \ and\ \bibinfo {author}
  {\bibfnamefont {N.~F.}\ \bibnamefont {Johnson}},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Nature}\ }\textbf {\bibinfo {volume} {462}},\
  \bibinfo {pages} {911} (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Dedeo}\ \emph {et~al.}(2011)\citenamefont {Dedeo},
  \citenamefont {Krakauer},\ and\ \citenamefont {Flack}}]{Dedeo2011}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Dedeo}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Krakauer}}, \
  and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Flack}},\
  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Journal of the
  Royal Society Interface}\ }\textbf {\bibinfo {volume} {8}},\ \bibinfo {pages}
  {1260} (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Barab\'{a}si}(2005)}]{Barabasi2005}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.-L.}\ \bibnamefont
  {Barab\'{a}si}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Nature}\ }\textbf {\bibinfo {volume} {435}},\ \bibinfo {pages} {207}
  (\bibinfo {year} {2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Gabaix}\ \emph {et~al.}(2003)\citenamefont {Gabaix},
  \citenamefont {Gopikrishnan}, \citenamefont {Plerou},\ and\ \citenamefont
  {Stanley}}]{Gabaix2003}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
  {Gabaix}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
  {Gopikrishnan}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
  {Plerou}}, \ and\ \bibinfo {author} {\bibfnamefont {H.~E.}\ \bibnamefont
  {Stanley}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Nature}\ }\textbf {\bibinfo {volume} {423}},\ \bibinfo {pages} {267}
  (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Johnson}\ \emph {et~al.}(2013)\citenamefont
  {Johnson}, \citenamefont {Medina}, \citenamefont {Zhao}, \citenamefont
  {Messinger}, \citenamefont {Horgan}, \citenamefont {Gill}, \citenamefont
  {Bohorquez}, \citenamefont {Mattson}, \citenamefont {Gangi}, \citenamefont
  {Qi}, \citenamefont {Manrique}, \citenamefont {Velasquez}, \citenamefont
  {Morgenstern}, \citenamefont {Restrepo}, \citenamefont {Johnson},
  \citenamefont {Spagat},\ and\ \citenamefont {Zarama}}]{Johnson2013b}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~F.}\ \bibnamefont
  {Johnson}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Medina}},
  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Zhao}}, \bibinfo {author}
  {\bibfnamefont {D.~S.}\ \bibnamefont {Messinger}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Horgan}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Gill}}, \bibinfo {author} {\bibfnamefont {J.~C.}\
  \bibnamefont {Bohorquez}}, \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
  {Mattson}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Gangi}},
  \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Qi}}, \bibinfo {author}
  {\bibfnamefont {P.}~\bibnamefont {Manrique}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Velasquez}}, \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Morgenstern}}, \bibinfo {author}
  {\bibfnamefont {E.}~\bibnamefont {Restrepo}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Johnson}}, \bibinfo {author} {\bibfnamefont
  {M.}~\bibnamefont {Spagat}}, \ and\ \bibinfo {author} {\bibfnamefont
  {R.}~\bibnamefont {Zarama}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
  {journal} {Scientific Reports}\ }\textbf {\bibinfo {volume} {3}},\ \bibinfo
  {pages} {3463} (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Johnson}\ \emph {et~al.}(2011)\citenamefont
  {Johnson}, \citenamefont {Carran}, \citenamefont {Botner}, \citenamefont
  {Fontaine}, \citenamefont {Laxague}, \citenamefont {Nuetzel}, \citenamefont
  {Turnley},\ and\ \citenamefont {Tivnan}}]{Johnson2011a}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
  {Johnson}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Carran}},
  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Botner}}, \bibinfo
  {author} {\bibfnamefont {K.}~\bibnamefont {Fontaine}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Laxague}}, \bibinfo {author} {\bibfnamefont
  {P.}~\bibnamefont {Nuetzel}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Turnley}}, \ and\ \bibinfo {author} {\bibfnamefont
  {B.}~\bibnamefont {Tivnan}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
  {journal} {Science Magazine}\ }\textbf {\bibinfo {volume} {333}},\ \bibinfo
  {pages} {81} (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Hall}\ and\ \citenamefont {Coyne}(2013)}]{wars}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~R.}\ \bibnamefont
  {Hall}}\ and\ \bibinfo {author} {\bibfnamefont {C.~J.}\ \bibnamefont
  {Coyne}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Defence
  and Peace Economics}\ }\textbf {\bibinfo {volume} {25}},\ \bibinfo {pages}
  {445} (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Johnson}\ \emph {et~al.}(2006)\citenamefont
  {Johnson}, \citenamefont {Spagat}, \citenamefont {Restrepo}, \citenamefont
  {Becerra}, \citenamefont {Camilo}, \citenamefont {Su\'{a}rez}, \citenamefont
  {Restrepo},\ and\ \citenamefont {Zarama}}]{Johnson}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~F.}\ \bibnamefont
  {Johnson}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Spagat}},
  \bibinfo {author} {\bibfnamefont {J.~A.}\ \bibnamefont {Restrepo}}, \bibinfo
  {author} {\bibfnamefont {O.}~\bibnamefont {Becerra}}, \bibinfo {author}
  {\bibfnamefont {J.}~\bibnamefont {Camilo}}, \bibinfo {author} {\bibfnamefont
  {N.}~\bibnamefont {Su\'{a}rez}}, \bibinfo {author} {\bibfnamefont {E.~M.}\
  \bibnamefont {Restrepo}}, \ and\ \bibinfo {author} {\bibfnamefont
  {R.}~\bibnamefont {Zarama}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
  {journal} {arXiv:physics/0605035}\ } (\bibinfo {year} {2006})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Clauset}\ and\ \citenamefont
  {Gleditsch}(2012)}]{Clauset2012}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Clauset}}\ and\ \bibinfo {author} {\bibfnamefont {K.~S.}\ \bibnamefont
  {Gleditsch}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {PloS ONE}\ }\textbf {\bibinfo {volume} {7}},\ \bibinfo {pages} {e48633}
  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Havlin}\ and\ \citenamefont
  {Ben-Avraham}(1982)}]{Havlin1982}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Havlin}}\ and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {Ben-Avraham}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
  {Physical Review A}\ }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages}
  {1728} (\bibinfo {year} {1982})}\BibitemShut {NoStop}%
\end{thebibliography}%

\clearpage

\end{document}

### 2015-11-07 paper-template.outtakes.tex
first deployment

\begin{figure*}
  \centering
  \includegraphics[width=0.8\textwidth]{cartoon_highlight.pdf}
  \caption{
    Partioning of Charles Dickens's ``Tale of Two Cities''
    into five different kinds of elements: 
    clauses (red),
    pure random partitioning phrases ($\partitionprob=\frac{1}{2}$, orange),
    words (yellow),
    pure random partitioning graphemes ($\partitionprob=\frac{1}{2}$,
    green),
    and
    letters (blue).
    \textbf{A.}
    Partition examples for the start of the work.
    The specific phrases and graphemes shown are for one realization
    of pure random partitioning.
    \textbf{B.}
    Zipf distributions for the five kinds of partitions
    along with estimates of exponents.
    See supplementary material for measurement details.
    \todo{add connecting line for letters only as a guide}
    \todo{remove estimate for letters? not signficant?}
    \todo{2 decimal places with errors.}
    \todo{Change 'slope' to $\simonalpha$.}
   }
  \label{fig:tpz.twocities}
\end{figure*}

\todo{Insert note that clauses are the most independent and follow something like a Simon's model}

Continuing with the statistical mechanical analogy,
we see that because random partitioning can be viewed as an infinite 
temperature limit ($\beta=0$, $T=\infty$), we may be reassured that the resulting Zipf
distribution is robust for a wide range of adjacent partitions (i.e., $\beta$ near 0).
The traditional zero temperature word-centric partition 
($\beta \rightarrow \infty$, $T \rightarrow 0^{+}$) 
both performs poorly and lacks stability.
In allowing the non-physical but abstractly allowable adjustment to
$\beta \rightarrow -\infty$ ($T \rightarrow 0^{-}$),
we would have clauses partitioned only as themselves,
and the resulting distribution of so many disparate, unique entities
would depart even further from a pure scaling form.\\
\todo{Fix the above up given new figures.}
\todo{Can we say this?}\\
We note that the appearance of a power law is not guaranteed.
As a simple example, if we apply random partitioning to a repeating sequence
of the same word, say `ook', then the resulting rank distribution
will be exponential.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{twoCitiesVertCartoon.pdf}
  \caption{
    \todo{Add caption.}
    \todo{Measure exponents.}
    This is one random 
  }
  \label{fig:tpz.twocities}
\end{figure}

%% Here, we present a fast, efficient, and scalable method for
%% partitioning arbitrary, large-scale texts into meaningful words and phrases.

%% n-grams
%% requiring our collective analysis

\todo{Reconciling with text-mixing: Decay of phrase introduction is
  slower when phrases are involved.  Word order matters.}

\todo{Do following for Fig. 1}\\
\todo{Please convert to $\log_{10} r$, etc.} \\
\todo{Remove bold from headings}\\
\todo{increase font size for axis labels}\\
\todo{Underlay main figure with normal Zipf distribution?
  That is, take the lower subplots out.
  Could have the normal one offset lower.
  Main thing is to make the comparison as clear as possible.}

For each corpus we observe a scaling
break of reduced severity following the regime over which Zipf's
law (black line main plot, dashed red in the lower left) is
present.

This reduction in the severity of the lower tail is
consistent with text mixing theory, as differences of word sets
between the texts of a corpus are compunded with word order, which
creates a greater diversity of symbols---hence a decreased rate of
symbol-introduction decay---between texts.

Also, we observe that
despite having a greater number of unique words, the Twitter
dataset has fewer partitioned elements than either the New York
Times or 
This is possibly due to the
character limit on the tweet lengths, and the fact that alphabetic
sequences are used in creating URLs, which, despite having little
semantic
meaning are picked up by our parser as word forms.

\begin{figure}[tbp!]
  \centering
  \includegraphics[width=.47\textwidth]{Grandma.pdf}
  \caption{
    The possible random partitions of the string `let's eat
    Grandma' ($\sigma$).
    Each partition is equally likely giving rise
    to expected weights for the phrases they contain.
    The six phrases along with their weights are
    `let's' (1/2),
    `eat' (1/4),
    `Grandma' (1/2),
    `let's eat' (1/4),
    `eat Grandma' (1/4),
    and 
    `let's eat Grandma' (1/4).
    These weights sum to that the number of words 
    in the string, $n_\sigma=3$.
    \todo{Add weights for random partitions}
    \todo{Choose a more friendly phrase?}
    \todo{Use / or $\mid$ instead of comma}
  }
  \label{fig:textpart.grandma}
\end{figure}

We show a simple example of random partitioning for the
punctuation-sensitive phrase `let's eat Grandma'.  in
Fig.~\ref{fig:textpart.grandma}.

Random Partitioning gives us a useful way to measure
the `strength' of a given phrase $\tau$ with respect to its local contexts---what we will
call $\tau$'s \textit{local context conditional entropy} $\lcce_\tau$---in the following way.
First, we define the phrase neighborhood $\mathcal{N}_\tau$
`generated' by a phrase $\tau$ as the collection of
all length $n_{\tau}+1$ phrases found in the text $T$
that can be formed from $\tau$ by an insertion of a single word.
We in turn define the neighborhoods to which a phrase 
owes membership as its \textit{internal contexts}.
Our example phrase `let's eat Grandma' 
is thus a member of three internal
contexts (if they exist in $T$): 
$\mathcal{N}_{\textnormal{let's eat}}$, 
$\mathcal{N}_{\textnormal{let's Grandma}}$, 
and $\mathcal{N}_{\textnormal{eat Grandma}}$.
Denoting a context as $C$ and the set of all
internal contexts of $\tau$ as $\mathcal{C}_\tau$, we define 
$\lcce_\tau = \max_{C \in \mathcal{C}_\tau} - \log_{2} P(\tau \mid C)$.
Thus, $\lcce_\tau$ is derived from 
most unlikely and most information rich
inner context of $\tau$.

\todo{Include somewhere?}: Though there is no reason to assume that a distribution of text
separated $\&$ rank-order phrase counts abides the Zipfian hypothesis,
it is known that a pure Zipfian word distribution can only allow for
non-Zipfian $n$-gram distributions \cite{egghe????a}.

\todo{Include somewhere?:
A great deal of research has been done in the
way of MWE extraction, where most successful developments often rely
upon parallel corpora.  
In this investigation we provide an extraction
technique that relies upon the text being interpreted alone.}

\section{Discussion}
\label{sec:textpart.discussion}

A perhaps understated feature of the Serial Partitioning algorithm is
the means by which likelihood data might be accessed.
The method relies upon local likelihoods defined by the internal
context model. 
These likelihoods are determined according to the tree structure
proposed in~\cite{???}.
Interestingly, one could just as well store these likelihoods on the
aforementioned tree.
Accessing this data then becomes computationally efficient (searches
are log-time on data trees), which allows for biologically feasiblity.

As mentioned in~\cite{williams2013a} language occurs on a multitude of
scales. 
This paper focuses on the phrase scale, for its position as a highly
active regime.
Serial Partitioning provides a novel technique for the detection of
meaningful language units.  
In the N.L.P. community these techniques could easily be implemented
to build better translation services with no need for pre-translated
parallel corpora.  
From any serially encoded text, our method extracts likely lexical
content - providing a means to direct the (context-dependent)
identification and translation of idioms in the most advanced
M.T. techniques.

In the phrase distributions section above we describe the distribution
of expected counts.
For a text, this distribtuion is defined by a random and independent
partitioning of text strings into contiguous sequences of words.
Here we perform precisely this operation, heretofore referred to as
random partitioning.

Our Serial Partitioning algorithm is in large effective due to a
strong conditioning present in our context model.
This conditioning on the words is precisely a realization of the
non-independence of word appearances as one reads a text.
Hence, the words of a text do no appear in accord to the original
Simon model~\cite{simon1955a}, i.e. independently.

As a premise for this work, we supposed that meaningfully partitioned
phrases are directly interpretable for meaning.
These phrases

This can be explained by the independance of partitioned phrases from
any other previously partitoned phrases. 
So phrases of a meaningfully partitioned text appear more
independently than the words.

This then is our heuristic justification for why informed text
partitions adehere to Zipf's law more so than words.
What is more, we expect that the phrases of a random partition will be
even more independent (they are without reference to possible
meaning-dependencies), and hence be even more in accord with the
Zipf-Simon relationship.

We present a random partition performed on the NY times text (Figure
2).
Note that this distribution posseses a small number of units scaling
more shallowly than the $-1$ exponent, and then rigidly follows a
scaling quite close to $-1$.
these two regions are indeed in direct accord to the analysis of the
original Simon model.
To test the possibility that a random partition is a realization of
the Simon model, we perform Kolmogorov-Smirnov \todo{check spelling
  and perform test} tests on randomly partitioned texts against
distributions of the same size generated as per the Simon model (Table
2).
\todo{make table of K-S test results for various innovation rates}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{Cite Sole papers}

\todo{Explain partitioning in one or two paragraphs}

\todo{Describe empirical results}

\todo{Describe Model}

\todo{Describe Analysis}

\todo{Give results of model}

\todo{Discussion}

\todo{Include music lyrics again?}

\todo{Show that Random Partitioning does a good job for Zipf?}

In Fig.~\ref{fig:textpart.mobydick_randompart}, we compared
the standard Zipf distribution for Moby Dick with the 
phrase weight distribution we obtain from randomly partitioning.\\
\todo{Expand on these results.}\\
\todo{Add NYT or Wiki?}

\begin{figure*}[tbp!]
  \centering
  \includegraphics[width=\textwidth]{2013-10-generatedLanguage.pdf}
  \caption{
    A. Generated language word distribution. 
    B. Generated language phrase distribution. 
    Both word and phrase distribution upper ($\gamma_{1}$) and lower ($\gamma_{2}$) 
    scaling exponents reported in graph. 
    C. The percentage of usage of words by order (1,\ldots,10 represented cold to hot), 
    with both rank and inception as predictor variable. 
    D. Grey bars indicate the percent of all units counted that each order possesses, 
    and red bars indicate the percent of all unique units each order possesses. 
    E. A log-log heat map of word inception vs word rank. 
    F. Shannon phrase (red) and word (black) distribution entropies over of time. 
    G. Phrase (red) and word (black) distribution nullity over time.
  }
  \label{fig:textpart.generatedlanguage}
\end{figure*}

\todo{Lead into why we must go to serial partitioning}

\textbf{Local Serial Partitioning}---Random Partitioning gives us a background distribution
of phrases to work with, but one that clearly does not account for how a
reader might parse text
(necessarily working with limited cognitive resources~\cite{levy????a}),
but rather provides us with only an ensemble of possible partitions.
To capture the act of interpretation, we introduce a local serial
partitioning method 
which `reads' through strings to find phrases which
maximize entropy local context conditional entropy.

\begin{figure}[tbp!]
  \centering
  \includegraphics[width=.47\textwidth]{Grandma.pdf}
  \caption{
    The possible random partitions of the statement `let's eat Grandma'.
  }
  \label{fig:textpart.grandma}
\end{figure}

Below, we step through an example of Serial Partitioning for the string `let's eat Grandma'.
Starting at the left, we set our first partition to be the null phrase
$\varphi$ 
which has local context conditional entropy $\lcce_\varphi =  0$.
The first word is concatenated onto our partition, producing the phrase `let's',
whose positive entropy keeps the process going.
The word `eat' is then concatenated to form `let's eat'. 
Since the entropy increases, we keep the phrase and continue.
Upon adding the word `Grandma', the entropy decreases, 
and we fix the phrase `let's eat' as the first partition.
The process begins again with the null phrase $\varphi$.
The next phrase continues with the word `Grandma' and the partitioning
is complete (and ethically sound).
\begin{enumerate}
\item let's eat Grandma /  $\lcce_\varphi=0$
\item $\underbracket[0.5pt]{\textnormal{\textbf{let's}}}$ eat Grandma /  $\lcce_\text{let's}>\lcce_\varphi=0$
\item $\underbracket[0.5pt]{\textnormal{\textbf{let's eat}}}$ Grandma /  $\lcce_\text{let's eat}>\lcce_\text{let's}$
\item $\underbracket[0.5pt]{\textnormal{\textbf{let's eat Grandma}}}$ / $\lcce_\text{let's eat Grandma}$$<$$\lcce_\text{let's eat}$
\item $\underbracket[0.5pt]{\textnormal{let's eat}} \,|\,$ Grandma /  $\lcce_\varphi=0$
\item $\underbracket[0.5pt]{\textnormal{let's eat}} \,|\,$ $\underbracket[0.5pt]{\textnormal{\textbf{Grandma}}}$ / $\lcce_\text{Grandma}>\lcce_\varphi= 0$
\end{enumerate}
By moving through the full text $T$, treating string by string 
with Serial Partitioning, we extract a unique partition of $T$ into words and phrases.

\begin{figure*}[tbp!]
  \centering
  \includegraphics[width=\textwidth]{dick_random_english_counts.pdf}
  \caption{
    \todo{Finish caption.}\\
    \textbf{A.} Word frequency as a function of rank for Moby Dick.
    \todo{Describe inset in A.}
    \textbf{B.} The weight distribution for Randomly Partitioned phrases.
    }
  \label{fig:textpart.mobydick_randompart}
\end{figure*}

\begin{figure*}[tbp!]
  \centering
  \includegraphics[width=\textwidth]{2013-07-times_counts.pdf}
  \caption{
    \textbf{A.} The distribution of words from the New York Times corpus.
    \todo{Describe inset in A.}
    \textbf{B.} Serially partitioned phrases from the New York Times corpus. 
    An ideal Zipf distribution will appear as a straight line of slope $-1$ (red dotted line for reference). 
    The inset bar plot shows the percent of all unique symbols
    partitioned at each phrase length (pink), 
    and the percent of the total number of units partitioned at each
    order (grey)
    \todo{more clarify in preceding.}
    The phrases presented are ranks $10^n$ ($n=0,...,6$) in both the word and phrase separated distributions. 
    Even at high ranks, phrase units are interpretable (often idiomatic) regular language components, 
    while high ranking words like `cabbage', `mysteriousness' and
    `fish-slicing' are difficult to glean meaning from outside of
    context.\newline
    \todo{describe double fit and give errors}\newline
    \todo{Change order to 'phrase length' in plot}\newline
    \todo{Move slopes to be near actual fits}\newline
    \todo{Move example phrases so they are attached to points on plot}\newline
    \todo{Add examples of 3, 4, 5, 6 word phrases?}
    }
  \label{fig:textpart.nytimes}
\end{figure*}

We now employ this Serial Partitioning scheme to examine
two disparate, large-scale corpora: 
the English version of Wikipedia and the New York Times (see Methods).
We show how the standard word frequency distribution compares with
the serially partitioned  phrase distribution for the New York Times in
Fig.~\ref{fig:textpart.nytimes}
(see Fig.~\ref{fig:textpart.wikipedia} for the Wikipedia plots).
The characteristic scaling break (or `knee') of the Zipf  word distribution
(Fig.~\ref{fig:textpart.nytimes}A)
is in strong evidence, occurring between word ranks of $10^3$ and $10^4$.
For low rank words, frequency $f_r$ scales as $r^{-0.98 \pm ??}$, in
accordance with Zipf's original observations~\cite{zipf1949a} and many that have followed.
Above the break, the slope becomes much steeper with 
$f_r \sim r^{-1.85 \pm ??}$.
\todo{Describe the inset in A}

But for both corpora, this break in scaling largely evaporates
for the distribution for phrases.
For the New York Times, the scaling for infrequent phrases 
follows  $r^{-1.19 \pm}$, weakly different from $r^{-1}$ (Fig.~\ref{fig:textpart.nytimes}B).
Points are colored according to phrase length, revealing a gradual,
mixed transition through phrases of increasing length as rank increases.

The reduction of a break in scaling comes about 
becomes words that appear frequently and are very likely to appear in sequence
will agglomerate into higher order units (see `new york', `york city', and
`new york city' in Fig.~\ref{fig:textpart.nytimes}B \todo{Bring these back?}). 
This process not only diminishes the frequency of the individual words
(suppressing the scaling break), but introduces new, lower-frequency
atoms (phrases with two or more words),
which fill out the tail.

\todo{Explain insets}

Apart from the phrase distribution's close adherence to an idealized Zipf's
relationship, we also find that the most frequently occurring phrases are
interpretable, quite representative of the source corpora, and often
idiomatic (see Tabs.~\ref{tab:textpart.nytimesphrases} and~\ref{tab:textpart.wikipediaphrases}).
Given the relatively simple nature of Serial Paritioning,
the dramatic movement toward a pure Zipf distribution along with the extraction
of reasonable phrases suggests that improved or alternate methods of
partitioning should produce similar results.

\todo{Add more description}

Plotting the resulting phrase distributions for phrases up to length
10, we see that the Zipfian
hypothesis is supported remarkably well.

\section{Stochastic Model}
\label{sec:textpart.model}

In~\cite{araujo2013}, the argument is made that a `mature' natural
langauge is of a Ziphan statistical structure.
Their argument hinges on the property that the Zipfian (harmonic)
distribution has an entropy sum that diverges (in the limit).
Consequently, a Zipfian distributuion may have an arbitrarily large
average entropy as more units are incorporated.
The perspective then is that there is little rank-mobility; the
language is matured or saturated.

However as we and others~\cite{ha2009a,gerlach2013a} have observed, large
enough samples of text produce a break in scaling in their rank-frequency
distributions.
Though some hold the perspective that those words beyond the scaling break are
`junk'~\cite{???}, we view this as case in point of a
non-Zipfian word distribution.
Assuming then that the word distributions are in some way non-Zipfian,
we develop a model based on Simon's~\cite{simon1955a}.
In this model we seek to determine what language development processes
can lead to the formation of a scaling break.

Integral to this development are two interacting distributions: a
`phrase distribution' of constructed units, and a `word
distribution' that is the atomization of the former.
With each time step there are three probabilities, $\alpha$,
$\beta_1$, and $\beta_2$ that govern the construction process. With
each step one will:
\begin{itemize}
\item[$\cdot$] 
  choose a word with probability $\beta_1$;
\item[$\cdot$] 
  choose no word with probability $1-\beta_1$;
\item[$\cdot$] 
  choose a phrase with probability $\beta_2$;
\item[$\cdot$] 
  choose no phrase with probability $1-\beta_2$;
\end{itemize}
where if a word is chose, it will be
\begin{itemize}
\item[$\cdot$] 
  a new word with probability $\alpha$;
\item[$\cdot$] 
  an old word with probability $1-\alpha$.
\end{itemize}

In general any phrase chosen will be an old one, and old words and
phrase are chosen in proportion to the number of times they have
previously appeared. If no word and no phrase are chosen, then the
process is null, rendering no phrase. If both word and phrase are
chosen, then the two will be composed to form the newly constructed
phrase that is counted into the phase distribution, and whose
constituent words are counted into the word distribution.

Perhaps the most straightforward manner to compose phrase and word
then would be to concatenate the word onto the phrase, but, as it
changes our analysis of the resulting word distribution in no way, we
compose the word and phrase together in a more nuanced manner.
In particular, if our phrase has length $n$, we keep track of the
insertion frequncy of usage of each of the $n+1$ possible insertion
points (every phrase insertion history is initialized to a nuiform
state upon its inception). Then, chooseing an insertion point in
proportion to its frequency of usage, the word is inserted
accordingly.

\section{Model Analysis}
\label{sec:textpart.modelanalysis}

To describe the the growth of the word distribution in the stochastic
model presented in the previous section we define $\mathcal{N}_k$ to be the
number of $k$-count words in the word distribution, $n_w$ to be the
total number of words in the word distribution, $n_p$ to be the number
of phrases in the phrase distribution, and $M_k(i)$ to be the number
of (non-unique) phrases in the phrase distribution with $i$ words of
word-count $k$. Then, measuring time by $t$, the number of non-null
construction steps, the growth in $\mathcal{N}_k$ per non-null construction
event is given by
\begin{equation}
\frac{\rm{d}\mathcal{N}_k}{\rm{d}t}
=
\beta_1\frac{\rm{d}\mathcal{N}_k}{\rm{d}n_w}
+
\frac{\beta_2}{n_p}
\left[
  \sum_{i=1}^\infty iM_{k-1}(i)-\sum_{i=1}^\infty iM_{k}(i)
\right],
\end{equation}
where as in~\cite{simon1955a,krapivsky2001a} we know
$$\frac{\rm{d}\mathcal{N}_k}{\rm{d}n_w}=\frac{1}{n_w}\left[(k-1)\mathcal{N}_{k-1}-k\mathcal{N}_k\right](1-\alpha)+\alpha\delta_{k,1}$$
and similarly assume that $n_wp_k=\mathcal{N}_k$. On the left hand side we
immediately find the differential,
$\text{d}n_{w}/\text{d}t$.  
So, since the $M_k(i)$ have the nice property:
$$
\sum_{i=1}^{\infty}
iM_{k}(i) 
=
k\mathcal{N}_k,
$$
the right hand side becomes,
$$
\left[
\beta_1(1-\alpha)
+
\beta_2
\frac{n_w}{n_p}
\right]
\left[
(k-1)p_{k-1}-kp_k
\right]
+\alpha\beta_1\delta_{k,1}.
$$
On average a constructed phrase will have been the result of $\beta_1$
word-drawn words and $\beta_2(n_w/n_p-\beta_1)$ phrase-drawn words. As
such we expect that the approximation
$$\frac{\rm{d} n_w}{\rm{d}t}\approx\beta_1+\beta_2\left(\frac{n_w}{n_p}-\beta_1\right)$$
will hold for large $t$. Putting this all together we find the recursion equation
$$p_k=p_{k-1}\frac{(k-1)}{k+1/\eta},$$
where
$$\eta=\frac{\beta_2\frac{n_w}{n_p}+\beta_1(1-\alpha)}{\beta_2\frac{n_w}{n_p}+\beta_1(1-\beta_2)}.$$

As usual the recursion equation is solved in terms of gamma functions,
where the resulting fit applies to the low-$k$ words. In particular,
if $t_b$ is the time at which the $b^\text{th}$ new word
appears, those words in teh vicinity of rank $b$ will scale with
exponent $-\eta_{t_b}$ in the rank-frequency distribution. The
proportion $n_w/n_p$ does grow (though slowly) over time, and the
result there is tail, initial steep, but gradually increasing.

While foregoing an analysis of the phrase distribution, we do take
time to consider what value $\eta$ may take. Note that if $\beta_2=0$
the model collapses back to that of Simon, with rank-frequency
exponent $\eta=1-\alpha$. In general, we see that if $\beta_2>\alpha$
one ahs $\eta>1$, and that $\eta$ indeed rises with $\beta_2$. If
$\beta_2$ is quite nearly $1$, we obtain
$$\eta\approx1+\beta_1(1-\alpha)\frac{n_w}{n_p},$$
which has the upper limit $1+n_p/n_w$ as $\beta_1\rightarrow 1$ and
$\alpha\rightarrow 0$.

\section{Simulation Results}
\label{sec:textpart.simresults}

We find that indeed the analysis of our model aligns with those runs
observe, where the fit is indeed better as $t$ becomes large. In fact
the fit converges faster for larger values of $\beta_2$ (the phrase
construction process is more dominant then). A peculiar aspect of the
model as presented is that the words with the highest count will
frequently reside in a family of very repetitive phrases. Aside from a
small class of lyrics and chat phrases in our partitioning experiments
(e.g. `ha ha ha...', `la la la...', etc.), most realistic phrases
do not involve word reptition (and certinly no the mose frequent
ones). As such a more nuanced model may have some form of repetition
inhibition. As seen in Fig. 4, such a low-reptition accommodation does
not cause the model to deviate from the analysis, while producing a
more reasonable collection of phrases.

In th previous section our analysis shows that our model is capable of
producing rank-frequency exponents steeper than $-1$. Further, since
$\beta_2=0$ produces the Simon model we may create a knee (see Fig. 4)
in the word distribution via the stepwise procedure: initiate the
model with $\beta_2=0$ to produce a Zipf-like scaling of $f\propto
r^{1-\alpha}$, and after the $b^\text{th}$ word appears begin
the phrasing process by setting $\beta_2>0$.

Supposing the growth of a language follows the course outlined above, it 
initiates with words forming and recurring independently, i.e. the
Simon model. For this period the words bear no relation to one another, 
i.e. are free of interword context. At some point a language participant 
changes the game, lifting a sequence of words from their literal meaning, 
now as an entity of its own. In the war of communication the process
snowballs with other participants borrowing meaning from existing words
to build their own colorful and abstracted interpretations. Though in
the model the point of transition is not endogenously artificed, it is
indeed conceivable in the realm of social dynamics.

%%%%%%

\todo{Sort the following out.  The main problem is that
serial partitioning does not seem to be exactly the plan
when we consider all possible internal contexts.  Also,
all decisions are made on global information, and only
in part on past experience.}

\todo{internal versus local contexts.}

%% jake:
%% http://www.longtail.com/the_long_tail/2006/09/is_zipfs_law_ju.html

In general, we consider the joint pdf for phrases and contexts
$P(\tau,C)$ with $\tau \in \mathcal{T}$ and $C \in \mathcal{C}$, their
full sets.
The conditional distribution $P(\tau \,|\, C)$
then gives the probability of a phrase $\tau$ appearing
given context $C$.
We refer to these probabilities as \textit{local likelihoods}, 
and use them as a basis for our serial partitioning method.

Let $\mathcal{T}$ be the usual collection of all phrases constructable from a
finite word set.
With our context collection $\mathcal{C}$, we must now make some
assumptions to form a valid joint pdf on $\mathcal{T}\times\mathcal{C}$.
Suppose that an interpreter experiences each internal context upon
presentation with a phrase.

For local contexts, we define the phrase neighborhood generated by a phrase
$\tau$, $\mathcal{N}_\tau$, as the collection of
all length $n_{\tau}+1$ phrases found in the text $T$
that can be formed from $\tau$ by an insertion of a single word.
We in turn define the neighborhoods to which a phrase 
owes membership as its \textit{internal contexts}.
Our example phrase `let's eat Grandma' 
is a member of three internal
contexts: $\mathcal{N}_{\textnormal{let's eat}}$, 
$\mathcal{N}_{\textnormal{let's Grandma}}$, 
and $\mathcal{N}_{\textnormal{eat Grandma}}$.

Upon presentation of the phrase
`let's eat Grandma', each of the contexts
$\mathcal{N}_{\textnormal{let's eat}}$,
$\mathcal{N}_{\textnormal{let's Grandma}}$, and
  $\mathcal{N}_{\textnormal{eat Grandma}}$ are experienced once.
So, using Random Partitioning to generate
a distribution of phrase weights for a text, we draw a
phrase from $\mathcal{T}$ in proportion to its count, and define a joint pdf
from the marginals:
$$
P(\tau)=P(\textnormal{an internal context is experienced from $\tau$})
$$
and
$$
P(C=\mathcal{N}_\tau)
=
P(\textnormal{the internal context $\mathcal{N}_\tau$ is experienced}).
$$

Note that we can represent the first marginal by weighting the counts
of each phrase by its length, but the framing does not aid in the
formation of a joint distribution, as it is without reference to
context experience.

We then take that a written body is separable, with ideal separations
accessible via some ``minimal effort'' algorithmic optimization.

Our objective is to break partition a text into meaningful units.

To the computational linguist, what we refer to here as phrases are
often called multi-word expressions (MWEs) or $n$-grams.

The identification of MWEs in natural
language poses one of the largest obstacles to accurate machine
translation~\cite{sag????a}.

A great deal of research has been done in the
way of MWE extraction, where most successful developments often rely
upon parallel corpora.

In this investigation we provide an extraction
technique that relies upon the text being interpreted alone.

Others have noted similar results by combining frequency distributions of n-grams,
though this approach overcounts words~\cite{ha2002a}.

A natural language speaker must consider the likelihood of a statement
or phrase's interpretation in context, that much is clear. 
But what precisely is context? 
For us it is both the previously experienced corpus and the content of
the utterance thus far. 
Previous approaches~\cite{piantadosi2011a} define the context-- for us,
\emph{local context}-- as follows. 
Given a phrase of length $n$, the context of the final word is the
sequence of $n-1$ preceding words.
In \cite{williams2013a} the bent on phrases (as opposed to just words)
allows for a more general definition.

If words formed a layer of units of meaning,
then Simon's model would suffice for an approximate
origin of language model.

for phrase extraction to partition natural language,
we

At the core of much quantitative language analysis
is simple word counting.

Models

Zipf's law has

In previous work \cite{Williams2013a} a natural language is defined
recursively, with the notion of lexicon made relative to the scale of
language being considered.
Assuming this to be the case, we ask how a text's lexicon of
meaningful phrases may be uncovered.
It is easy to ask questions about the collections of fundamental
pieces (e.g. letters in words, or words in clauses), but the trickier
business lies with `meaningful' units in clauses and words.
Since these non-trivial regimes are defined by interpretation through
similarity of past experience, we develop a model for the
identification of lexical atoms based on morphological similarity and
accumulated inferential data.

It is of some surprise then that Zipf's law is surprisingly limited
for language, applying typically to only around the first few thousand
most frequently used words whereupon a clear break in scaling
appears~\cite{bentknee,ferrericancho,ha2009a}.

\begin{figure}[tbp!]
  \centering
  \includegraphics[width=0.48\textwidth, angle=0]{mary2.pdf}
  \caption{
    X-bar theoretic decomposition of the sentence `Mary had a ball at
    her dance recital and knocked it out of the park'. Observe that
    the grammatical dissection finds both `ball' and `it' as subjects
    possessed by `Mary', acted on in conjoined verb phrases. From this
    perspective it seems that `it' is the `ball'.
  }
  \label{fig:textpart.mary}
\end{figure}

Let's observe the severity of the phrase identification problem and consider whether or not the statement
\begin{center}
\begin{minipage}{.9\linewidth}
\center
Mary had a ball at her dance recital and knocked it out of the park.
\end{minipage}
\end{center}
\noindent is so easily interpreted grammatically (Diagram 1), or
literally.  Were this the only option, we might believe that Mary had
a dance recital in a park where she somehow knocked a toy sphere
beyond the park's limits. This is troubling since the colloquial
speaker will immediately know that Mary's ball was not a toy sphere,
but instead a positive experience. The root of this concern may be
seen through a naive grammatical interpretation that implies ``it'' in
the phrase, ``knocked it out of the park'', is the ``ball'' (Diagram
1). Though ``it'' was indeed the positive experience that Mary
excelled at, equating the two easily leads a non-native speaker to the
incorrect literal interpretation, after all, a ball might certainly be
batted about. For the native speaker, knowledge of context directs
their interpretation of ``knocked it out of the park'' as a closed
structure; i.e. ``it'' is internal and fixed within an atomic phrase--
that is, it cannot be broken down any further.

\clearpage

Previous:

Let's observe the severity of the phrase identification problem and
consider whether or not the statement
\begin{center}
\begin{minipage}{.9\linewidth}
\center
Mary had a ball at her dance recital and knocked it out of the park.
\end{minipage}
\end{center}
is so easily interpreted grammatically~\ref{fig:textpart.mary}, or literally.
Were this the only option, we might believe that Mary had a dance
recital in a park where she somehow knocked a toy sphere beyond the
park's limits. This is troubling since the colloquial speaker will
immediately know that Mary's ball was not a toy sphere, but instead a
positive experience. The root of this concern may be seen through a
naive grammatical interpretation that implies `it' in the phrase,
`knocked it out of the park', is the `ball'~\ref{fig:textpart.mary}. 
Though `it'
was indeed the positive experience that Mary excelled at, equating the
two easily leads a non-native speaker to the incorrect literal
interpretation, after all, a ball might certainly be batted about. For
the native speaker, knowledge of context directs their interpretation
of `knocked it out of the park' as a closed structure; i.e. `it' is
internal and fixed within an atomic phrase-- that is, it cannot be
broken down any further.

The linguist Brame \cite{Brame01} viewed natural language as a
recursive system - one where complex relational structures build up
from smaller ones until they gel into the fundamental pieces of the
next scale. Assuming this to be the case, how do we know which
portions of an utterance are atomic and cannot be parsed further? It
is easy to ask questions about the collections of fundamental pieces
(e.g., letters in words, or words in clauses), but the thornier
territory is always in the mixing regimes (e.g., the morphemes in
words, or `meaningful' phrases in clauses). Being that these
non-trivial regimes are defined by interpretation through similarity
of past experience, we develop a model for the identification of
lexical atoms based on morphological similarity and accumulated
inferential data.

A natural language speaker must consider the likelihood of a statement
or phrase's interpretation in context, that much is clear. But what
precisely is context? For us it is both the previously experienced
corpus, and the content of the utterance thus far. Previous approaches
\cite{Piantadosi01} define the context-- for us, \emph{local
context}-- as follows. Given a phrase of length $n$, the context of
the final word is the sequence of $n-1$ preceding words. We may, by
allowing phrases (as opposed to just words) to be the context-bearing
units, find a more general definition. Instead of just the $n-1$
leading words it is well to consider any $n-1$ length subsequence of
words. Here we build a top-down approach to local context which
replaces the assumption that the role filled by a word is fixed at
its' utterance, and replaces it with a windowed flexibility. We thus
look to organize language by collections of phrases with similarly
ordered sub-structures, this being the shape of languages of recursion
grown.

\section{The tree structure of phrases and resulting metric}

Let $W$ be the collection of words in a language including the null
word, $\varnothing$, and $S$ be the collection of all finite-length
sequences of words from $W$. We will refer to elements of $S$ as
phrases. For phrases $\sigma,\tau\in S$ we define the relation
$\preceq$ by setting $\tau\preceq\sigma$ if and only if $\sigma$ may
be constructed from $\tau$ through successive insertions. For example,
if $\tau=(the\:brown\:fox)$, and
$\sigma=(the\:quick\:brown\:fox\:jumped)$, then $\tau\preceq\sigma$.

The collection of all phrases in a language is immediately an
infinite, unbounded tree - a partially ordered set of phrases
branching out from the null word. With this structure in mind, we
refer to the first neighborhood of an element $\sigma$, or $N_\sigma$,
as the set of all phrases constructible from $\sigma$ by a single
insertion, e.g., with $\sigma$ as in the previous we see that $(the\:
quick\: brown\: fox)\in N_\sigma$.

As is shown in Appendix III, the collection of phrases has a wealth of
structure endowed to it via the metric induced by $\preceq$. First one
defines the \emph{order} map $\#: S\rightarrow\mathbb{Z}$ to send
$\sigma$ to the number terms (words) it contains. Then, for any
$\sigma,\tau\in S$ define $\eta$ to be a \emph{maximal sequential
intersection} of the pair, if $\eta$ is a subsequence of both $\sigma$
and $\tau$ that has order at least as big as any other subsequence
common to the pair. Then the function $d:S^2\rightarrow\mathbb{Z}$
defined by
$$d(\sigma,\tau)=\#\sigma+\#\tau-2\#\eta$$
is a metric on $S$ (Appendix III). The main value of the metric is the
opportunity to build a notion of similarity on $S$. This is easily
accomplished by defining the normalized value
$\rho:S^2\rightarrow\mathbb{Q}$
$$\rho(\sigma,\tau)=\frac{d(\sigma,\tau)}{\#\sigma+\#\tau}.$$
The similarity function then shows us that $N_\sigma$ is for $\sigma$
the set of $\rho$-maximal similar elements (Appendix III), which we
then take as justification of the choice of the $N_\sigma$ as the
local contexts. The graph and resulting metric framework provide a
means to tease apart the regular recursive forms, but as we will see
shortly, it is these sequential perspectives in combination with
inferential data that directs the process of interpretation.

\section{Counting distributions}
In \cite{Zipf01}, George K. Zipf demonstrated that most frequently
appearing words in the English language strongly adhered to a simple
relationship. If the words were ranked ascending according to their
counts descending (i.e. the most frequently appearing word gets rank
1), then a word's rank, $R$, would be inversely proportional to its
count $R$, that is $R\times S= C$, for some universal constant $C$. In
part this observation achieved such fame for it's ubiquity and
simplicity. Zipf offered a simple formula representing the
interrelations between the words of a language, a possible
quantitative lead on the psychological processes underpinning a
language - his hypothesis being the `principle of least
effort''\cite{Zipf04}.

Regardless of the meaning behind Zipf's observation, implicit in his
analysis was the assumption that words form the natural atomization
for the clauses of a text. This assumption offers a very convenient
separation of a corpus, i.e. a counting by words assures that no
portion of a text is accounted for more than once. Here we seek
alternate and natural ways to separate a text.  In particular, we
develop distributions of phrases which arise from text separations.

To date the most extensively developed corpus of higher order lexical
in the computational linguistics community is that created by the
Google corporation \cite{Google01}. Despite its recent appearance as
data in various scholarly articles\cite{Garcia01,Piantadosi01} , the
perspective taken by Google and much of the $n$-grams community lacks
a thorough consideration of the active function of non-alphanumeric
symbols, ignoring punctuative boundaries and essentially considering
sequences embedded with punctuation to be potentially interpretable
phrases. In the Google corpus the entity `it's' is considered a phrase
of three words, with the apostrophe given the same status as a
word. This perspective is stuck, straddling the character-atomic and
word-atomic views simultaneously. Furthermore, as a consequence of the
sliding-window method of counting $n$-grams, the resulting
distribution cannot be integrated into a single distribution
meaningfully (each word forms a portion of many different $n$-grams).

Previous work has shown that a Zipfian word distribution can only
allow for non-Zipfian phrase distributions \cite{Egghe01}.
We, as
\cite{Ha01} have done, take a posteriori that the word distribution is
in fact non-Zipfian (since the relationship does break down).
Though
\cite{Ha01} tried to extend Zipf's work to $n$-grams, their attempts
utilized the sliding-window method of counting $n$-grams, immediately
destroying the statistical integrity of an integrated distribution.

As an alternative to these approaches this work takes punctuation
characters as rigid boundaries, and then bends the notion of counting
to be probabilistic.
Behind the punctuation-boundary assumption is the
observation that punctuation gives the interpreter an understanding of
where interpretable phrases start or end.
By `counting' words in a
probabilistic way, we construct a distribution of phrases respectful
to the distribution of words in a text.

Suppose now that of the (infinite?) collection of possible
interpretations of a clause, some reasonable interpretations are
modeled by the various separations, or, partitions of that clause.  In
particular, we assume that different partitions correlate to different
inflections, which correlate to different meanings.  We formalize this
as follows: given a clause $\sigma\in S$ of order $n$, define the
\emph{naive partition liklihood}, $p_\sigma(\tau)$, for any
\emph{full} (contiguous) phrase $\tau\preceq\sigma$, to be the
probability that a partition chosen at random (among the possible
$2^{n-1}$) isolates $\tau$.

As an example we let $\sigma=(let's\:eat\:grandma)$.  The possible
partitions of a clause nicely represented through the placement of
commas~\ref{fig:textpart.grandma}.
Of the four possible partitions of $\sigma$ we see
\begin{align*}
p_\sigma(let's)=1/2 \hspace{30pt}&\hspace{10pt} p_\sigma(let's\:eat)=1/4\\
p_\sigma(eat)=1/4 \hspace{36pt}&\hspace{10pt} p_\sigma(eat\:grandma)=1/4\\
p_\sigma(grandma)=1/2 \hspace{10pt}&\hspace{10pt} p_\sigma(let's\:eat\:grandma)=1/4.
\end{align*}
Among the first of things to note is, that each boundary a phrase
shares with the clause increases its partition probability by a factor
of 2. Also, since these probabilities are not mutually exclusive, they
do not sum to 1. In general we see that for any $\sigma\in S$ of order
$n$ and $\tau\preceq\sigma$, a full phrase of order $k$, that
$$
p_\sigma(\tau)
=
2^{b-(k+1)},
$$
where $b$ is the number of boundaries (left or right) that $\sigma$
and $\tau$ share (Appendix III).

\section{Local-context partitioning}

To rigorously develop this material we formally define a \emph{text}
to be a pair $(S,m)$, where $S$ is the space of finite-length
sequences of words from $W$, and $m:S\rightarrow \mathbb{Z}$ is the
clause multiplicity map, where a clause once again is defined to be
any sequence of words bounded by punctuation. Then, if we restrict our
analysis to any record of language, $m(\sigma)$ is the number of times
$\sigma\in S$ appeared as a clause in the record. With this, for any
$\tau\in S$ we define the \emph{naive expected count}, $c_\tau$, of
$\tau$ in the text to be
$$
c_\tau
=
\sum_{\sigma\in S}
m(\sigma)
p_{\sigma}(\tau),
$$
where $p_{\sigma}(\tau)=0$ if $\tau$ is not a full subsequence of
$\sigma$. Then, the \emph{expectation of occurrence} $E:2^S\rightarrow
\mathbb{Q}$ for any subset $A\subseteq S$ is well-defined by
$$
E(A)=\sum_{\tau\in A}c_\tau.
$$
The function $E$ is essentially describes the mass of a collection of
phrases in a text.

Given $\tau\in S$ with $\#\tau=n>0$, we now wish to determine which of
the $n$ possible local contexts $\tau$ represents. In section II we
define the first neighborhood of a phrase $\sigma$ to be the
collection of order $\#\sigma+1$ phrases constructible from
$\sigma$. Now, for a phrase $\tau$ the local contexts are precisely
the first neighborhoods that it belongs to. Therefore we take the
determination of a local context for $\tau$ to be the inference of
$\tau$'s construction. A reasonable approach, perhaps, assumes
construction of $\tau$ by $\sigma$ (i.e. assumes the context
$N_\sigma$) in proportion to $N_\sigma$'s expectation of
occurrence. Then, upon observing $\tau$ one assumes the context
$N_\sigma$ with probability
$$\frac{E(N_\sigma)}{\underset{\tau\in N_\eta}{\sum}E(N_\eta)}.$$
However, in the main body of this work, we take the simplification
that $\tau$ belongs to the context $N_\sigma$ with greatest
expectation of occurrence.

With our simplification, each phrase has a fixed context. So, we
consider the likelihood of a phrase's appearance given the presence of
a particular context. We define his quantity, the \emph{local
likelihood}, for any $\tau$ with context $N_\sigma$ to be
$$P(\tau\mid N_\sigma)=\frac{E(\tau)}{E(N_\sigma)}.$$
Local likelihood measurements form the basis of the partitioning
schemes in this work.

To develop a partitioning scheme we consider cognitive theories
\cite{Levy01}, which pose that the natural human syntactic parser
operates with limited resources. Typically the limited resource
assumed is memory. We may certainly average the local likelihood of
each clause partition and choose the one with greatest mean likelihood
(Appendix III). However, we must consider again that a clause of $n$
words has $2^{n-1}$ distinct partitions. Is it reasonable to expect a
human mechanism to scan all partitions? Given the memory-resource
limitation hypothesis, this brute force method seems unnatural, and
for long clauses totally unreasonable. Hence we expect that ideal
partitions of human language should be accessible through some
restriction of the search space.

To simplify the search problem we observe that English (along with
many other languages) relies heavily on a directional flow. This
property is exemplified by the `garden path' sentences of the English
language. Though these statements often lack punctuation and function
words, we find them instructive to understanding the limits of the
human lexical parser, most importantly the extent to which English
commonly reads from left to right. Consider the statement
\begin{center} 
  The prime number few.
\end{center} 
Since we immediately see the beginning of the phrase `the
prime number...'', we are forced to think of mathematics and attempt
to understand what sort of a number `few' is. However, a more
internally consistent interpretation considers `prime' to be a noun,
and verbs the word `number' into meaning `existence in a quantity
of''. Though this confusion is easily remedied through the insertion
of an implicit noun such as `people' into the statement `the prime
people number few'', the point is clear.  Given this nuance in the
English language we then expect that a reasonable mechanism `breaks
phrases off' from a clause, assessing the local likelihood of each
phrase from left to right. We refer to this process as \emph{serial
partitioning}~\ref{fig:textpart.grandma}.

\section{Experiment, Results, and Discussion}

Partitioning the Wikipedia.en, NY Times and Lyrics corpora via the
serial partitioning scheme with probabilistically determined contexts
(materials and methods discussed in Appendix II), we count the
occurrence of phrases up to order 10. Plotting the resulting phrase
distributions, we see that the Zipfian hypothesis is supported
remarkably well (Figure 1-A). The characteristic `knee' of the classic
word-separated distribution is greatly diminished and can be
attributed to the following. Words that appear frequently (see new,
york, and city in Figure 1-A) and are very likely to appear in
sequence will agglomerate into higher order units (see new york, york
city, and new york city in Figure 1-B). This process not only
diminishes the frequency of the individual words (suppressing the
knee), but introduces new, lower-

\noindent frequency atoms, which fill out the tail. Irrespective of
the phrase distribution's adherence to Zipf's relationship, we find
that the most frequently occurring phrases are interpretable, quite
representative of the source corpora and often idiomatic (Table 1).

Investigation to the Zipfian hypothesis immediately begs the question:
why a scaling exponent of $-1$? A well-known measure quite relevant to
this is that of Shannon and others \cite{Kolmogorov01, Shannon01},
known as information content.  This measure is most easily understood
as the average quantity of bits (when the logarithm is taken base 2)
per symbol required to efficiently store or represent a message.

The average representation-information required per symbol in a
Zipfian language of $N$ atoms with rank-decreasing frequencies
$\{p_k\}_{k=1}^N$, and scaling exponent $\gamma>1$ is given by the sum
$$-\sum_{k=1}^{N}p_k\log{p_k}=\frac{\gamma}{M_N(\gamma)}\left[\frac{\log{M_N(\gamma)}}{\gamma M_N(\gamma)}+\sum_{k=1}^{N}\frac{\log{k}}{k^\gamma}\right],$$
where $M_N(\gamma)$ is the $N$th generalized harmonic number of
exponent $\gamma$. Note that $M_N(\gamma)$ and the sum are finite in
the limit as $N\rightarrow\infty$ so long as $\gamma>1$. Hence a
growing language that maintains the Zipfian relationship may expand
its vocabulary indefinitely, i.e. the number of meaningful atoms may
increase without bound while the average representation-information
per symbol approaches a finite limit.  Furthermore, the value of the
theoretical, infinite sum, may become arbitrarily large as
$\gamma\rightarrow1$, meaning that such a language may maintain
discernible atoms which, on average, may represent any quantity of
information thinkable. For a system of encoding this seems to be a
powerful and provocative feature.

Regardless of what the information measure represents and what the
slope of a phrase distribution means, we once again note the value of
the context-dependent parser in application. Without any necessity for
parallel corpora, this development provides a novel MWE detection
technique.  From any serially encoded text, this method extracts the
likely lexical units, which in turn may be assessed for proper
translation.

\section{Sub-word partitioning}

As explained in [citation needed] a language posses multiple scales for representing meaning.
So far in this work, we have partitioned units at the word-sequence scale.
We can, however, look inside words (i.e. look at words like clauses), and partition letter-sequence units.
This partitioning can be done analagous to the method for phrases,
or, since word formation rules are more stringent
(e.g. pronouncable English units generally contain vowels and are composed of graphemes),
we can leverage known sub-word structure to obtain more realistic partitions.

Nevertheless,
the goal remains a partition that carries meaning.
At the word-seqence scale there is no qantative way to determine if a phrase is meaningful.
However, at the sub word scale we know nearly all combining forms [cite combining forms list],
in addition to roots, prefixes, and suffixes.
So, we present the sub-word partitioning results from two variants of the serial partitioning algorithm.
in addition to standard algorithm and random partitioning (see Table 1).
Each of these variations are performed by only taking expected counts from partitions adherent certain restrictions.
The first restriction requires that all partition components possess a vowel,
and the second requires that all partition componets possess a vowel or be a grapheme.

\begin{center}

{Coverage of Sub-Words Partitioned}\newline

{\fontsize{6.8}{8.16}\selectfont

\begin{tabular}{|c|c|c|c|c|}
\hline Source & Random & None  & Vowel & Grapheme\\\hline
NYTimes& 62.42\% (2435) & 72.79\% (100) &68.26\% (226) & 72.16\%(89) \\\hline
Wikipedia& 63.42\% (3191)& 0 & 0 &0 \\\hline
\end{tabular}
}
\end{center}

\section{Glossory of terms}
{\footnotesize
\textbf{Clause:} Any phrase in a corpus that abuts punctuation characters.\newline
\textbf{Corpus:} A sequential record of language.\newline
\textbf{First neighborhood:} For a phrase, The collection of all sequences constructible by it though a single insertion.\newline
\textbf{Full sub-phrase:} Of a phrase, any contiguous sub-phrase of terms.\newline
\textbf{Lexical unit/atom:} A minimal-length interpretable phrase. \newline
\textbf{Local likelihood:} The conditional likelihood of a phrase in context.\newline
\textbf{Maximal Sequential Intersection:} For two phrases, any maximal length sub-phrase common to the pair.\newline
\textbf{Naive expected count:} Sum of naive partition likelihood for a phrase across a text.\newline
\textbf{Naive partition likelihood:} The probability that a partition of a clause delineates a full sub-phrase of the clause.\newline
\textbf{Occurrence expectation:} The sum of naive expected count of the phrases in a collection.\newline
\textbf{Order of a phrase:} The number of terms in a phrase.\newline
\textbf{Partition-mean local likelihood:} The harmonic mean of a partition's components' local likelihood.\newline
\textbf{Partition of a clause:} A collection of full sub-phrases that when composed in sequence produce the clause.\newline
\textbf{Phrase:} Any sequence of words.\newline
%%\textbf{Positionally disjoint sub-phrases:} For a phrase, sub-phrases whose terms have differing indices.\newline
\textbf{Separated corpus:} A distribution of phrases that arises when every clause in a corpus is partitioned.\newline
\textbf{Sub-phrase:} For a phrase, any other phrase comparable in the partial order of no larger order.\newline
\textbf{Terms:} The (not possibly non-unique) words composing a phase.\newline
\textbf{Text:} Representing a corpus, $S$ paired with a map $m$ into the integers describing the multiplicity of occurrence of the clauses in the corpus. \newline
\textbf{Word:} Character sequences delineated by spaces and the null word, $\varnothing$.
}

\begin{thebibliography}{1}

\bibitem{Brame01} Brame, \emph{Recursive categorical syntax and morphology I: Semigroups, monoids and categories}, Linguistic Analysis,1984.

\bibitem{Egghe01} Egghe, \emph{On the law of Zipf-Mandelbrot for multi-word phrases}, Journal of the American Society for Information Science and Technology, 1999.

%%\bibitem{Frank01} Frank, Jaeger \emph{Speaking rationally: Uniform information density as an optimal strategy for language production}, Proceedings of the 30th Annual Meeting of the Cognitive Science Society, 2008.

\bibitem{Garcia01} Garcia, Garas and Schweitzer, \emph{Positive words carry less information than negative words}, 2012.

\bibitem{Google01} \emph{All Our N-gram are Belong to You},\\ http://googleresearch.blogspot.com/2006/08/all-our-n-gram-are-belong-to-you.html
\bibitem{Ha01}Ha, Hanna, Ming and Smith, \emph{Extending Zipf's law to n-grams for large corpora}, Artificial Intelligence Review, 2009.

\bibitem{Kolmogorov01} Kolmogorov, \emph{Three approaches to a quantitative definition of information}, Problems of Information Transmission, 1965.

\bibitem{Lakoff01}Lakoff and Johnson, \emph{Metaphors We Live By}, Univ. of Chicago Press, 1980.
\bibitem{Lakoff02}Lakoff, \emph{Women, Fire, and Dangerous Things}, University of Chicago Press, 1987.

\bibitem{Levy01} Levy, \emph{Probabilistic Models of Word Order and Syntactic Discontinuity}, Stanford
University Press, 2005.

%%\bibitem{Mandelbrot01} Mandelbrot, \emph{Information theory and psycholinguistics}, Scientific Psychology: Principles and Approaches, 1965.
%%\bibitem{Mandelbrot02} Mandelbrot, \emph{A note on a class of skew distribution function. analysis and critique of a paper by H.A. Simon}, Information and Control, 1955.
%%\bibitem{Mandelbrot03} Mandelbrot, \emph{Final note on a class of skew distribution functions: analysis and critique of a model due to H.A. Simon}, Information and Control, 1961.
%%\bibitem{Mandelbrot04} Mandelbrot, \emph{Post scriptum to `final note'}, Information and Control, 1961.

%%\bibitem{Miller01} Miller and Chomsky, \emph{Handbook of Mathematical Psychology II}, Wiley, pp. 419-491, 1963.

%%\bibitem{Nicolis} Nicolis and Tsuda, \emph{On the parallel between Zipf's Law and 1/f processes in chaotic systems possessing coexisting attractors: a possible mechanism for language formation in the cerebral cortex}, Progress of Theoretical Physics, 1989.

%%\bibitem{Orlov01} Orlov, \emph{Why, how, and when does the Zipf-Mandelbrot law fail?}, in Statistical Methods in Linguistics, 1977.

\bibitem{Piantadosi01} Piantadosi, Tily and Gibson, \emph{Word lengths are optimized for efficient communication}, PNAS, 2010
%%\bibitem{Piantadosi02} Piantadosi, Tily and Gibson, \emph{The communicative function of ambiguity in language}, Cognition, 2010.
%%\bibitem{Piantadosi03} Piantadosi, \emph{Learning and the language of thought}, PhD thesis, MIT, 2011.
%%\bibitem{Piantadosi04} Piantadosi, Tily and Gibson, \emph{Information content and word length in natural language}, (under review), 2012.

\bibitem{Reilly} Reilly and Kean \emph{Information content and word frequency in natural language: Word length matters}, Proc Natl Acad Sci USA, 2011.

%%\bibitem{Simon01} Simon, \emph{On a class of skew distribution functions}, Biometrika, 1955. 
%%\bibitem{Simon02} Simon, \emph{Some further notes on a class of skew distribution functions}, Information and Control, 1960.
%%\bibitem{Simon03} Simon, \emph{Reply to `final note' by Benoit Mandelbrot}, Information and Control, 1961.
%%\bibitem{Simon04} Simon, \emph{Reply to Dr. Mandelbrot's post scriptum}, Information and Control, 1961.

\bibitem{Sag01} Sag, Baldwin, Bond, Copestake, and Flickinger, \emph{Multiword Expressions: A Pain in the Neck for NLP}, Proc. of the 3rd International Conference on Intelligent Text Processing and Computational Linguistics.
\bibitem{Shannon01} Shannon, \emph{A Mathematical Theory of Communication}, Bell System Technical Journal, 1948.
%%\bibitem{Shannon02} Shannon, \emph{Prediction and entropy of printed English}, Bell System Technical Journal, 1951.

%%\bibitem{Yavuz} Yavuz, \emph{Zipf's law and entropy}, IEEE Transactions on Information Theory, 1974.

\bibitem{Zipf01} Zipf, \emph{Relative frequency as a determinant of phonetic change}, Harvard Studies in Classical Philology, 1929.
\bibitem{Zipf02} Zipf, \emph{the psycho-biologogy of language: an introduction to dynamic philology}, The M.I.T. Press, 1935.
%%\bibitem{Zipf03} Zipf, \emph{National Unity and Disunity: The Nation As a Bio-Social Organism}, Principia Press, 1941.
\bibitem{Zipf04} Zipf, \emph{Human Behavior and the Principle of Least Effort}, Addison-Wesley, 1949.

%%\bibitem{Zornig01} Zornig and Altmann, \emph{The repeat rate of phoneme frequencies and the Zipf-Mandelbrot law}, Glottometrika, 1983.

\end{thebibliography}

We further argue that this view of phrases as the meaningful unit of
language reveals the English language possesses
optimal information content with 
the ability to describe an infinite set of things.  Meh.

minimal meaningful phrases rather than individual words,

that the frequency of words is inversely proportional
the remarkable 
to their rank

and its origins

Zipf's law is famous, much debate over its origin
-1 exponent

We show how Zipf's law applies to English for
meaningful phrases rather than words

The words of human languages 
usage frequency
exhibit  Zipf's law
The Zipf's law is

Zipf's law is

Here, we employ a simple probabilistic algorithm
that identifies meaningful phrases by context in large-scale
corpora.
We use our parsing scheme to extend the work
of Zipf  to distributions of English phrases.
We find Zipf's power-law hypothesis holds for phrases
rather than words, 
a strong argument for the semantic correlation of our technique.
We find strong support for our observations 
with a modified rich-gets-richer 
mechanism for language evolution.

In doing so we meet Zipf's power-law hypothesis better than
with English words alone, and use phrase considerations to develop a
stochastic model akin to that of H. A. Simon \cite{Simon01}.  
Using
this stochastic model we identify parameter regimes which produce the
characteristic "bent knee" Zipf distributions, a strong argument for
the semantic correlation of our theory.

Words are easy.

The description of a language is historically split into two parts:
grammar and the lexicon.
While the former consists of rules that
govern composition, the lexicon is the wealth of words and phrases
understood through context by interpreters.
Though grammar provides us
with a formal system of analysis, in its rigidity it often fails to
represent the underlying cognitive processes by which the language
itself likely forms.

Given modern perspectives from the cognitive
linguistics community \cite{Lakoff01,Lakoff02}, it is important to ask

how we may represent and identify the interpreted units of the
lexicon, phrases which cannot be sensibly broken apart 
(e.g, ``War on Terror'' and ``New York City'')

\begin{figure*}[tbp!]
  \centering
  \includegraphics[width=1\textwidth, angle=0]{generatedLanguage44_2.pdf}
  \caption{
    A - Generated language word distribution. Open circles and dashed
    line are measured and smoothed from the derived scaling (Appendix IV).
    B - Generated language phrase distribution.  C - Shannon phrase (red)
    and word (black) distribution entropies over of time.  D - Grey bars
    indicate the percent of all units counted that each order possesses,
    and red bars indicate the percent of all unique units each order
    possesses.
  }
  \label{fig:textpart.generatedlanguagezipf}
\end{figure*}

### 2015-11-07 breakup-manuscript-supp.pl

@@ -1,0 +1,19 @@
+#!/usr/bin/perl
+
+$pdffile = $ARGV[0];
+($manufile = $pdffile) =~ s/\.pdf$/-manuscript.pdf/;
+($suppfile = $pdffile) =~ s/\.pdf$/-supplementary.pdf/;
+
+$supp_firstpage = `cat startsupp.txt`;
+chomp($supp_firstpage);
+$manulastpage = $supp_firstpage - 1;
+
+$pdfinfo = `pdfinfo $pdffile`;
+$pdfinfo =~ m/^Pages:\s+?(\d+)\s/ms;
+$lastpage = $1;
+
+`gs -sDEVICE=pdfwrite -dNOPAUSE -dBATCH -dSAFER -dFirstPage=1 -dLastPage=$manulastpage -sOutputFile=$manufile $pdffile $1>&2`;
+
+`gs -sDEVICE=pdfwrite -dNOPAUSE -dBATCH -dSAFER -dFirstPage=$supp_firstpage -dLastPage=$lastpage -sOutputFile=$suppfile $pdffile $1>&2`;
+
+

### 2015-11-07 log/publicationtimeline.m

@@ -1,0 +1,11 @@
+publog = {
+    [2014 06 03 11 33 00],'submission','Science';
+    [2014 06 21 01 46 00],'submission','PNAS';
+    [2014 10 28 12 00 00],'resubmission','PNAS';
+    [2015 01 02 01 42 00],'resubmission','PNAS';
+    [2015 01 09 13 10 00],'accepted','Nature';
+%%   [2011 10 23 15 27 00],'submission','PLoS ONE';
+%%   [2011 11 14 23 00 00],'resubmission','PLoS ONE';
+%%   [2011 11 29 17 32 03],'accepted','PLoS ONE';
+%%   [2012 01 11 17 00 00],'published','PLoS ONE';
+};

### 2015-11-07 make-single-latex-file.pl

@@ -1,0 +1,121 @@
+#!/usr/bin/perl
+$usage = "make-single-latex-file.pl foo.tex
+
+replaces \inputs{stuff} with stuff.tex,
+inputs bbl file, etc.
+
+improve input detection!!! (need to include file find)
+
+(usually run on foo-revtex4.tex)
+
+presumes that comments start with two percentage signs: %%
+(which works nicely for tab indenting in emacs so you
+should be doing it anyway)
+
+creates foo-combined.tex
+";
+if ($#ARGV < 0)
+{
+    print $usage;
+    exit;
+}
+($basefile = $ARGV[0]) =~ s/\.tex//;
+($realbasefile = $basefile) =~ s/-revtex4//;
+$file = "$basefile.tex";
+($bblfile = $file) =~ s/tex$/bbl/;
+
+$outfile = "$basefile-combined.tex";
+
+$help = "Parses through to remove all indexing commands,
+inputs all \inputs directly,
+and inputs the .bbl file
+";
+
+open (FILE,"$file") or die "can't open $file: $!\n";
+open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
+
+undef $tex;
+foreach $line (<FILE>)
+{
+    $line =~ s/%%.*$//;
+    if ($line =~ m/\\currfile/) {
+	$line = "";
+    }
+    
+    $line =~ s/\\filenamebase/$realbasefile/;
+    
+    if ($line =~ m/\\input/)
+    {
+	$line =~ m/\\input\{(.*?)\}/;
+#	print "$1\n";
+	$inputfile = $1.".tex";
+	open (INPUT,"$inputfile") or die "can't open $inputfile: $!\n";
+	undef $/;
+	$input = <INPUT>;
+	$/ = "\n";
+	close INPUT;
+
+	# special treatment for $realbasefile.title.tex
+	if ($line =~ m/\.title/) {
+	    $input =~ s/%.*?$//msg;
+	    $input =~ s/^\s*?//;
+	    $input =~ s/\s*?$//;
+	    chomp($input);
+	}
+
+	$line =~ s/\\input\{(.*?)\}/$input/;
+
+# one extra layer
+	while ($line =~ m/\\input\{(.*?)\}/) {
[…]

### 2015-11-07 makelinks.pl

@@ -1,0 +1,82 @@
+#!/usr/bin/perl
+# make links into package directory
+$usage = "
+usage: makelinks.pl foo-combined.tex
+(works for any latex file)
+
+use after make-single-latex-file.pl
+
+copies foo-combined.tex to ./package
+and replaces figure names by fig001, fig002, etc.
+
+see end of file for localized pieces
+";
+
+
+
+
+if ($#ARGV < 0)
+{
+    print $usage;
+    exit;
+}
+
+$file = $ARGV[0];
+
+unless (-e $file)
+{
+    print $usage;
+    exit;
+}
+
+# clean out the package directory
+# `\\rm package/*`;
+
+# find figures
+
+$outfile = "package/$file";
+open (FILE,$file) or die "can't open $file: $!\n";
+open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
+
+$i = 1;
+foreach $line (<FILE>)
+{
+    unless ($line =~ m/\s*%/) {
+	if ($line =~ m/\\includegraphics\[.*?\]\{(.*?)\}/) {
+	    push @figures, $1;
+
+	    if ($i<10) {$fignum = "00$i";} 
+	    elsif ($i<100) {$fignum = "0$i";}
+	    else {$fignum = $i;}
+	    $prefix = "fig".$fignum."_";
+	    $line =~ s/(\\includegraphics.*?)\{(.*?)\}/$1\{$prefix$2\}/;
+	    $i = $i + 1;
+	}
+    }
+    print OUTFILE $line;
+}
+
+close FILE;
+close OUTFILE;
+
+foreach $i (0..$#figures)
+{
+    $figure = $figures[$i];
+    $tmp = `find . -follow -name $figure -print 2>/dev/null`;
+    @tmp = split("\n",$tmp);
+    $fullfigures[$i] = $tmp[0];
+    chomp($fullfigures[$i]);
+}
+
+$i=1;
+foreach $fullfigure (@fullfigures)
+{
+    ($figname = $fullfigure) =~ s/.*\///;
+    if ($i<10) {$fignum = "00$i";} 
+    elsif ($i<100) {$fignum = "0$i";}
+    else {$fignum = $i;}
+    $prefix = "fig".$fignum."_";
+    `cp $fullfigure package/$prefix$figname`;
+    `ln -s ../$fullfigure package/.`;
+    $i = $i+1;
+}

### 2015-11-07 makerevtex4.pl

@@ -1,0 +1,103 @@
+#!/usr/bin/perl
+
+foreach $switch (@ARGV) {
+    if (($switch eq "-quick") or ($switch eq "-q"))
+    {
+	$quickswitch = 1;
+    }
+}
+
+$bodylist = `ls *.body.tex`;
+@bodies = split(/\s+/,$bodylist);
+
+($sec,$min,$hour,$numday,$month,$year,$wday,$yday,$isdst) = localtime(time);
+$year += 1900;
+$timestamp = "$year $month $numday $hour $min $sec";
+
+foreach $body (@bodies) {
+    ($filename = $body) =~ s/\.body\.tex$//;
+    if (-e "$filename-revtex4.tex") {
+	($logfilename = "log/progress-$body") =~ s/\.body\.tex$/-log/;
+	print "processing $filename...\n";
+
+	# create $filename-revtex4.settings.tex file
+	$settings = "$filename-revtex4.settings.tex";
+	open (SETTINGS,">$settings") or die "can't open $settings: $!\n";
+	print SETTINGS "\\newcommand{\\filenamebase}{$filename}\n";
+	close SETTINGS;
+
+	# check for postscript
+	open (BODY, "$body") or die "can't open $body: $!\n";
+	undef $/;
+	$bodytex = <BODY>;
+	close BODY;
+	if ($bodytex =~ m/\.ps\}/msg) {
+	    $pdfswitch = 0;
+	}
+	else {
+	    $pdfswitch = 1;
+	}
+	    
+	if ($pdfswitch == 1) {
+	    `pdflatex $filename-revtex4 1>&2`;
+	}
+	else {
+	    `latex $filename-revtex4 1>&2`;
+	}
+	unless ($quickswitch == 1) {
+	    `bibtex $filename-revtex4 1>&2`;
+	    if ($pdfswitch == 1) {
+		`pdflatex $filename-revtex4 1>&2`;
+		`pdflatex $filename-revtex4 1>&2`;
+	    }
+	    else {
+		`latex $filename-revtex4 1>&2`;
+		`latex $filename-revtex4 1>&2`;
+	    }
+	}
+	if ($pdfswitch == 0) {
+	    `dvips -o $filename-revtex4.ps $filename-revtex4 1>&2`;
+	    `epstopdf $filename-revtex4.ps 1>&2`;
+	}
+
+	# if open, bring to front
+        `open $filename-revtex4.pdf`;
+	# `open -a preview`;
+
[…]

### 2017-01-13 journal-formats/proceedings-of-the-national-academy-of-sciences/pnas-template-main.tex
add PNAS latex files

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

\title{Template for preparing your research report submission to PNAS using Overleaf}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,c,1]{Author One}
\author[b,1,2]{Author Two} 
\author[a]{Author Three}

\affil[a]{Affiliation One}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any conflict of interest here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format.  Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

\section*{Guide to using this template on Overleaf}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.

If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

\subsection*{Author Affiliations}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

\subsection*{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a PNAS PLUS research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for PNAS PLUS.

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}. All references, including for the SI, should be included in the main manuscript file. References appearing in both sections should not be duplicated.  SI references included in tables should be included with the main reference section.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{frog}
\caption{Placeholder image of a frog with a long example caption to show justification setting.}
\label{fig:frog}
\end{figure}

\subsection*{Digital Figures}
\label{sec:figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, options are to use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures, or to use the \verb|\begin{widetext}...\end{widetext}| environment as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and footnotes, as mentioned in the \href{http://texdoc.net/pkg/cuted}{cuted package documentation}. In the case of problems with footnotes, it may be possible to correct the situation using commands \verb|\footnotemark| and \verb|\footnotetext|.

%% Do not use widetext if paper is in single column.
\begin{widetext}
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{align*}
\end{widetext}

\begin{table}%[tbhp]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Supporting Information (SI)}

The main text of the paper must stand on its own without the SI. Refer to SI in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc. Authors are limited to no more than 10 SI files, not including movie files. Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text. Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.

\subsubsection*{SI Text}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).

\subsubsection*{SI Figures}

Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be pre-assembled.

\subsubsection*{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

\subsubsection*{SI Tables}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

\subsubsection*{SI Datasets}

Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

\subsubsection*{Still images}

Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.

\subsubsection*{Appendices}

PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% If you see unexpected formatting errors, try commenting out this line
% as it can run into problems with floats and footnotes on the final page.
\pnasbreak

% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2017-12-20 make-name-match-settingsfile.pl
adding make-name-match-settingsfile.pl

@@ -1,0 +1,20 @@
+#!/usr/bin/perl
+
+@files = <*-revtex4.tex>;
+
+foreach $file (@files) {
+    $file =~ m/^(.*?)-revtex4.tex/;
+    $papername = $1;
+
+    $settingsfile = $papername."-revtex4.settings.tex";
+
+    open (SETTINGSFILE, ">$settingsfile") or die "Can't open $settingsfile: $!\n";
+
+    print SETTINGSFILE "\\newcommand{\\filenamebase}{$papername}\n";
+    
+    close SETTINGSFILE;
+    
+
+}
+
+

### 2017-12-20 rename.pl
add rename.pl

@@ -1,0 +1,25 @@
+#!/usr/bin/perl
+# peter sheridan dodds
+# https://github.com/petersheridandodds
+# MIT License
+
+$usage = "rename text1 text2 file [files...]
+
+swap text1 for text2 in file names
+whenever text1 is found.
+
+swaps only the first instance.
+
+rudimentary first version: use only for static text.
+";
+
+$pattern1 = $ARGV[0];
+$pattern2 = $ARGV[1];
+
+foreach $filename (@ARGV[2..$#ARGV]) {
+    if ($filename =~ m/$pattern1/) {
+	($newfilename = $filename) =~ s/$pattern1/$pattern2/g;
+	print "renaming $filename as $newfilename...\n";
+	`mv -i "$filename" "$newfilename"`;
+    }
+}

### 2017-12-20 rename.pl
remove rename.pl

@@ -1,25 +1,0 @@
-#!/usr/bin/perl
-# peter sheridan dodds
-# https://github.com/petersheridandodds
-# MIT License
-
-$usage = "rename text1 text2 file [files...]
-
-swap text1 for text2 in file names
-whenever text1 is found.
-
-swaps only the first instance.
-
-rudimentary first version: use only for static text.
-";
-
-$pattern1 = $ARGV[0];
-$pattern2 = $ARGV[1];
-
-foreach $filename (@ARGV[2..$#ARGV]) {
-    if ($filename =~ m/$pattern1/) {
-	($newfilename = $filename) =~ s/$pattern1/$pattern2/g;
-	print "renaming $filename as $newfilename...\n";
-	`mv -i "$filename" "$newfilename"`;
-    }
-}

### 2021-05-21 correspondence/2017-01-13pre.body.tex
localization of figures, inputs, bibliography for better sharing

%% rob axtell
%% sid redner
%% neil johnson
%% martin gerlach
%% newman?
%% simon dideo
%% aaron clauset
%% dashun wang
%% y.-y. ahn
%% sune lehmann
%% luis amaral

To the Editors,

We are very pleased to submit our manuscript for consideration at
Physical Review E:\\
\textbf{``Simon's fundamental rich-gets-richer model entails a dominant first-mover advantage.''}

Simon's famous 1955 rich-gets-richer model is a fundamental
and elementary system-growth mechanism that leads to power-law 
size distributions of component sizes.  
Also known as cumulative advantage and preferential attachment in networks,
Simon's growth mechanism is of profound theoretical importance and widespread
application across complex systems of all kinds.

\textbf{Remarkably, we show that Simon's model has
been drastically misunderstood for over 60 years.}

In our manuscript, 
\textbf{we demonstrate that the first entrant into the system enjoys a first-mover advantage
completely missed in Simon's analysis and, from what we believe, all subsequent
work.}  Through simulations and theory, we show that 
the first-mover advantage is powerful with the size gap
growing as the inverse of the empirically small innovation rate.

As Figure 1 in our manuscript shows, the first mover may be
enormous and seemingly impossible to miss---yet this is what we
have discovered.

We also explore the dynamics of all groups showing the potential
for a high degree of unpredictability in a group's long term success,
connecting with similar observations in the literature.

\textbf{We provide empirical evidence from citation count data} that a pure Simon model
can be invoked to capture a real system with a clear first-mover advantage.

Our work calls into question all past empirical analysis of real-world
systems
using Simon's model
as well as much theoretical work,
and, in going forward, we provide possible adjustments to Simon's model to
better fit real-world data.

We note that we have received valuable feedback from a number of outstanding colleagues
after our initial posting on the arXiv and have improved the manuscript
substantially.

Our correspondents include:
\begin{itemize}
\item 
  Sidney Redner,
\item 
  James P.\ Gleeson,
\item 
  Mason A.\ Porter,
\item 
  Cesar A.\ Hidalgo,
\item 
  Matus Medo, and 
\item 
  Mark E.\ J.\ Newman.
\end{itemize}

\textbf{We believe our paper will be well received by a large and broad audience 
in the physics community and we hope will prove to 
be an important contribution in the years to come.}

%% \textbf{1.}
%% \textbf{2.}
%% \textbf{3.}
%% \textbf{4.}

%% \textbf{}

We look forward to hearing of your decision.

Yours sincerely and on behalf of the manuscript's authors,

%% Professor
%% Director of the Vermont Complex Systems Center
%% Co-Director, Computational Story Lab
%% Vermont Advanced Computing Center
%% Department of Mathematics and Statistics
%% The University of Vermont

### 2021-05-21 extractbibfile-universal.pl

@@ -1,0 +1,66 @@
+#!/usr/bin/perl
+# 
+# december 22, 2015, psdodds
+# 
+# original:
+# june 30, 1999, psdodds
+# extractbibfile.pl
+# 
+# extract bib file that pertains
+# exactly to an existing tex document
+# (builds on bibfind.pl)
+# requires name of tex file and 
+# .bbl file must exist
+
+# change next line if necessary
+$bibfile = "$ENV{HOME}/papers/biblio/everything.bib";
+
+$file = @ARGV[0];
+# remove any .tex suffix
+$file =~ s/^(.*)\.tex$/$1/;
+# remove any . leftover from tab completion
+$file =~ s/^(.*)\.$/$1/;
+
+$bblfile = "$file.bbl";
+$outfile = "$file-exact.bib";
+
+open (BBLFILE,"$bblfile") or die "can't open $bblfile: $!\n";
+open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
+
+## make a hash
+## only take first version
+open(BIBFILE, "$bibfile") or die "Can't open $bibdir$bibfile.bib: $!\n";
+
+undef $/;
+$bibfiles = <BIBFILE>;
+@bibentries = split(/^@/m, $bibfiles);
+
+foreach $bibentry (@bibentries) {
+ #   print $bibentry;
+    $bibentry =~ m/^.*\{(.*?),/;
+    $bibref = $1;
+#    print "$bibref\n\n";
+    if ($bibhash{$bibref} eq "") {
+	$bibhash{$bibref} = "@".$bibentry;
+    }
+#    print $bibhash{$bibref};
+}
+close BIBFILE;
+
+## print keys %bibhash;
+
+$/ = "\n";
+
+while ($line = <BBLFILE>) {
+    if ($line =~ m/^\\bibitem.*\{(.*?)\}/) {
+	$name = $1;
+	print "$name\n";
+#	print $bibhash{$name};
+	print OUTFILE $bibhash{$name};
+    }
+}
+
+print "$outfile created.\n";
+
+close BBLFILE;
+close OUTFILE;

### 2021-05-21 localizify.pl

@@ -1,0 +1,95 @@
+#!/usr/bin/perl
+
+#############################################
+## localizes figures and text snippet inputs
+#############################################
+
+$usage = "
+usage: localizify.pl
+
+does two things:
+
+1. finds and localizes all figures found in:
+
+*.body.tex *.supplementary.tex *.appendix.tex
+
+- searches for all matches to includegraphics
+- finds first match in figures tree
+  and stores in figures/localized
+- ignores figures/localized
+- overwrites figures
+
+2. finds and localizes all input text files in *.settings.tex
+
+- signal to do this by explicitly using \input{inputs/localized/...}
+- stores in inputs/localized/
+- expects snippets to be in ./figures tree only
+
+";
+
+##############################
+## figures
+##############################
+
+$localfiguresdir = "figures/localized";
+`mkdir -p $localfiguresdir`;
+
+$localinputsdir = "inputs/localized";
+`mkdir -p $localinputsdir`;
+
+$text = `grep includegraphics *.body.tex *.supplementary.tex *.appendix.tex 2>/dev/null`;
+
+@lines = split('\n',$text);
+foreach $line (@lines) {
+    $line =~ s/\%.*$//;
+
+    if ($line =~ m/includegraphics\[.*?\]\{$localfiguresdir\/(.*?)\}/) {
+	push @figures, $1;
+    }
+}
+
+# find first match in tree, could go wrong if multiple versions exist with same name.
+# but then you would have failed as a human anyway.
+foreach $i (0..$#figures) {
+    $figure = $figures[$i];
+	    
+    $filenames = `find figures -follow -name $figure -print 2>/dev/null`;
+    @filenames = split("\n",$filenames);
+    @filenames = grep { $_ !~ m/$localfiguresdir/ } @filenames;
+    
+    $fullfigures[$i] = $filenames[0];
+    chomp($fullfigures[$i]);
+
+#    print "$fullfigures[$i]\n $localfiguresdir/$figure\n";
+    `cp $fullfigures[$i] $localfiguresdir/$figure`;
+}
+
+##############################
+## snippet inputs
+##############################
+
+$text = `grep \\input *.settings.tex *.body.tex *.supplementary.tex *.appendix.tex 2>/dev/null`;
+
+@lines = split('\n',$text);
+foreach $line (@lines) {
+    $line =~ s/\%.*$//;
[…]

### 2021-05-21 make-arxiv-submission.pl

@@ -1,0 +1,12 @@
+#!/usr/bin/perl
+$usage = "make-arxiv-submission foo as in foo-revtex4.tex\n";
+
+$basefilename = $ARGV[0];
+
+print "making a single latex file...\n";
+`make-single-latex-file.pl $basefilename-revtex4.tex`;
+print "finding figures and linking...\n";
+`makelinks.pl $basefilename-revtex4-combined.tex`;
+print "making package...\n";
+`makepackage.sh $basefilename`;
+

### 2021-05-21 make-local-bibfile.pl

@@ -1,0 +1,76 @@
+#!/usr/bin/perl
+# 
+# 2021/05/21, psdodds
+# 
+# original:
+# june 30, 1999, psdodds
+# extractbibfile.pl
+# 
+# extract bib file that pertains
+# exactly to an existing tex document
+# (builds on bibfind.pl)
+# requires name of tex file and 
+# .bbl file must exist
+
+# 
+$usage = "make-local-bibfile.pl fooster.bib foo1.tex [foo2.tex] ...  \n" ;
+
+# change next line if necessary
+$bibfile = "$ENV{HOME}/papers/biblio/everything.bib";
+
+
+$outfile = @ARGV[0];
+
+@files = @ARGV[1..$#ARGV];
+
+## make a hash for bib entries
+## only take first version
+open(BIBFILE, "$bibfile") or die "Can't open $bibdir$bibfile.bib: $!\n";
+
+undef $/;
+$bibfiles = <BIBFILE>;
+@bibentries = split(/^@/m, $bibfiles);
+
+foreach $bibentry (@bibentries) {
+    #   print $bibentry;
+    $bibentry =~ m/^.*\{(.*?),/;
+    $bibref = $1;
+    #    print "$bibref\n\n";
+    if ($bibhash{$bibref} eq "") {
+	$bibhash{$bibref} = "@".$bibentry;
+    }
+    #    print $bibhash{$bibref};
+}
+close BIBFILE;
+
+
+foreach $file (@files) {
+
+    open (FILE,"$file") or die "can't open $file: $!\n";
+    $tex = <FILE>;
+    close FILE;
+
+    while ($tex =~ /\\cite\{(.*?)\}/msg) {
+	$keylist = $1;
+	$keylist =~ s/\s+//g;
+	@keys = split(',',$keylist);
+	foreach $key (@keys) {
+	    # repetition just writes over
+	    $localbibhash{$key} = $bibhash{$key};
+#	    print $bibhash{$key};
+	}
+	
+    }
+    
+
+}
+
+open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
+
+foreach $key (keys %localbibhash) {
+    print OUTFILE $localbibhash{$key};
+}
+
+close OUTFILE;
+
+print "$outfile created.\n";

### 2021-05-21 make-single-latex-file.pl

@@ -37,7 +37,10 @@ open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
 undef $tex;
 foreach $line (<FILE>)
 {
+    # Clean out comments
+    # Note: 
     $line =~ s/%%.*$//;
+    
     if ($line =~ m/\\currfile/) {
 	$line = "";
     }
@@ -48,7 +51,13 @@ foreach $line (<FILE>)
     {
 	$line =~ m/\\input\{(.*?)\}/;
 #	print "$1\n";
-	$inputfile = $1.".tex";
+
+	$inputfile = $1;
+	unless ($inputfile =~ m/\.tex/) {
+	    $inputfile = $inputfile.".tex";
+	}
+
+	
 	open (INPUT,"$inputfile") or die "can't open $inputfile: $!\n";
 	undef $/;
 	$input = <INPUT>;
@@ -68,7 +77,11 @@ foreach $line (<FILE>)
 # one extra layer
 	while ($line =~ m/\\input\{(.*?)\}/) {
 #	    print "$1\n";
-	    $inputfile = $1.".tex";
+	    $inputfile = $1;
+	    unless ($inputfile =~ m/\.tex/) {
+		$inputfile = $inputfile.".tex";
+	    }
+	    
 	    $inputfile =~ s/\\filenamebase/$realbasefile/;
 	    open (INPUT,"$inputfile") or die "can't open $inputfile: $!\n";
 	    undef $/;

### 2021-05-21 make-zip-file-for-overleaf.sh

@@ -1,0 +1,8 @@
+zip \
+    -FS \
+    --exclude=make-zip-file* \
+    --exclude=*.zip \
+    overleaf.zip \
+    * \
+    figures/localized/* \
+    inputs/localized/*

### 2021-05-21 makelinks.pl

@@ -59,8 +59,7 @@ foreach $line (<FILE>)
 close FILE;
 close OUTFILE;
 
-foreach $i (0..$#figures)
-{
+foreach $i (0..$#figures) {
     $figure = $figures[$i];
     $tmp = `find . -follow -name $figure -print 2>/dev/null`;
     @tmp = split("\n",$tmp);
@@ -69,12 +68,9 @@ foreach $i (0..$#figures)
 }
 
 $i=1;
-foreach $fullfigure (@fullfigures)
-{
+foreach $fullfigure (@fullfigures) {
     ($figname = $fullfigure) =~ s/.*\///;
-    if ($i<10) {$fignum = "00$i";} 
-    elsif ($i<100) {$fignum = "0$i";}
-    else {$fignum = $i;}
+    $fignum = sprintf("%03d",$i);
     $prefix = "fig".$fignum."_";
     `cp $fullfigure package/$prefix$figname`;
     `ln -s ../$fullfigure package/.`;

### 2021-05-21 makerevtex4.pl

@@ -1,5 +1,11 @@
 #!/usr/bin/perl
 
+# set up name in settings file
+`make-name-match-settingsfile.pl`;
+
+# localize figures and inputs (for better sharing)
+`./localizify.pl`;
+    
 foreach $switch (@ARGV) {
     if (($switch eq "-quick") or ($switch eq "-q"))
     {

### 2021-06-25 make-local-bibfile.pl
getting overleaf adjustments right

@@ -21,6 +21,13 @@ $bibfile = "$ENV{HOME}/papers/biblio/everything.bib";
 
 $outfile = @ARGV[0];
 
+if ($outfile !~ m/\.bib$/) {
+    print "Oops: $outfile is not a .bib file ...
+Exploding.
+";
+    exit;
+}
+
 @files = @ARGV[1..$#ARGV];
 
 ## make a hash for bib entries

### 2021-06-25 make-single-latex-file.pl

@@ -37,10 +37,7 @@ open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
 undef $tex;
 foreach $line (<FILE>)
 {
-    # Clean out comments
-    # Note: 
     $line =~ s/%%.*$//;
-    
     if ($line =~ m/\\currfile/) {
 	$line = "";
     }
@@ -53,11 +50,10 @@ foreach $line (<FILE>)
 #	print "$1\n";
 
 	$inputfile = $1;
-	unless ($inputfile =~ m/\.tex/) {
+	unless ($inputfile =~ m/\.tex$/) {
 	    $inputfile = $inputfile.".tex";
 	}
 
-	
 	open (INPUT,"$inputfile") or die "can't open $inputfile: $!\n";
 	undef $/;
 	$input = <INPUT>;
@@ -67,8 +63,8 @@ foreach $line (<FILE>)
 	# special treatment for $realbasefile.title.tex
 	if ($line =~ m/\.title/) {
 	    $input =~ s/%.*?$//msg;
-	    $input =~ s/^\s*?//;
-	    $input =~ s/\s*?$//;
+	    $input =~ s/^\s*//;
+	    $input =~ s/\s*$//;
 	    chomp($input);
 	}
 
@@ -76,12 +72,13 @@ foreach $line (<FILE>)
 
 # one extra layer
 	while ($line =~ m/\\input\{(.*?)\}/) {
-#	    print "$1\n";
+	    #	    print "$1\n";
+
 	    $inputfile = $1;
-	    unless ($inputfile =~ m/\.tex/) {
+	    unless ($inputfile =~ m/\.tex$/) {
 		$inputfile = $inputfile.".tex";
 	    }
-	    
+
 	    $inputfile =~ s/\\filenamebase/$realbasefile/;
 	    open (INPUT,"$inputfile") or die "can't open $inputfile: $!\n";
 	    undef $/;
@@ -91,6 +88,7 @@ foreach $line (<FILE>)
 	    $line =~ s/\\input\{(.*?)\}/$input/;
 	}
     }
+    
     $tex = $tex."$line";
 }
 

### 2021-06-25 make-zip-file-for-overleaf.sh

@@ -4,5 +4,7 @@ zip \
     --exclude=*.zip \
     overleaf.zip \
     * \
+    correspondence/* \
+    correspondence/figures/* \
     figures/localized/* \
     inputs/localized/*

### 2021-06-25 makelinks.pl

@@ -43,13 +43,18 @@ foreach $line (<FILE>)
 {
     unless ($line =~ m/\s*%/) {
 	if ($line =~ m/\\includegraphics\[.*?\]\{(.*?)\}/) {
-	    push @figures, $1;
 
-	    if ($i<10) {$fignum = "00$i";} 
-	    elsif ($i<100) {$fignum = "0$i";}
-	    else {$fignum = $i;}
+	    $figurename = $1;
+	    $figurename =~ s/^.*\///;
+	    # will fail for png, jpg
+	    unless ($figurename =~ m/\.pdf$/) {
+		$figurename =~ s/$/\.pdf/;
+	    }
+	    push @figures, $figurename;
+
+	    $fignum = sprintf("%03d",$i);
 	    $prefix = "fig".$fignum."_";
-	    $line =~ s/(\\includegraphics.*?)\{(.*?)\}/$1\{$prefix$2\}/;
+	    $line =~ s/(\\includegraphics.*?)\{(.*?)\}/$1\{$prefix$figurename\}/;
 	    $i = $i + 1;
 	}
     }

### 2021-06-25 makerevtex4.pl

@@ -5,16 +5,20 @@
 
 # localize figures and inputs (for better sharing)
 `./localizify.pl`;
-    
-foreach $switch (@ARGV) {
-    if (($switch eq "-quick") or ($switch eq "-q"))
+
+foreach $argument (@ARGV) {
+    if (($argument eq "-quick") or ($argument eq "-q"))
     {
 	$quickswitch = 1;
+    } elsif ($argument =~ m/\.body\.tex$/) {
+	@bodies = (@bodies, $argument);
     }
 }
 
-$bodylist = `ls *.body.tex`;
-@bodies = split(/\s+/,$bodylist);
+if ($#bodies eq -1) {
+    $bodylist = `ls *.body.tex`;
+    @bodies = split(/\s+/,$bodylist);
+}
 
 ($sec,$min,$hour,$numday,$month,$year,$wday,$yday,$isdst) = localtime(time);
 $year += 1900;
@@ -22,7 +26,16 @@ $timestamp = "$year $month $numday $hour $min $sec";
 
 foreach $body (@bodies) {
     ($filename = $body) =~ s/\.body\.tex$//;
+    ($supplementaryfilename = $body) =~ s/\.body\.tex$/\.supplementary\.tex/;
+    ($appendiexfilename = $body) =~ s/\.body\.tex$/\.appendix\.tex/;
     if (-e "$filename-revtex4.tex") {
+
+	# make local bibliography
+	# will need to be adjusted for more elaborate inputs
+	$texfiles = join(" ",<$filename.*.tex>);
+	$command = "make-local-bibfile.pl $filename.bib $texfiles";
+        system($command);
+
 	($logfilename = "log/progress-$body") =~ s/\.body\.tex$/-log/;
 	print "processing $filename...\n";
 
@@ -106,4 +119,5 @@ foreach $body (@bodies) {
 	print "Number of missing citations = $citenum\n";
 
     }
+
 }

### 2021-10-12 makerevtex4.pl
add singlelineify.pl, incorporate into makerevtex4.pl; small tweet for -revtex.tex file

@@ -36,6 +36,14 @@ foreach $body (@bodies) {
 	$command = "make-local-bibfile.pl $filename.bib $texfiles";
         system($command);
 
+	# make single line versions of title, abstract, logline
+	$command = "./singlelineify.pl $filename.title.tex\n";
+        system($command);
+	$command = "./singlelineify.pl $filename.abs.tex\n";
+        system($command);
+	$command = "./singlelineify.pl $filename.logline.tex\n";
+        system($command);
+
 	($logfilename = "log/progress-$body") =~ s/\.body\.tex$/-log/;
 	print "processing $filename...\n";
 

### 2021-10-12 singlelineify.pl

@@ -1,0 +1,42 @@
+#!/usr/bin/perl
+
+$usage = "singlelineify.pl foo1.tex [foo2.tex ...]
+
+- create foo1.txt [foo2.txt ...]
+  which are single line versions of the tex versions.
+
+- removes all empty lines, newlines, all commented out sections.
+
+- for use with 
+  .title.tex
+  .abs.tex 
+  .logline.tex
+
+- will mercilessly clobber .txt versions
+";
+
+if ($#ARGV < 0) {
+    print $usage;
+    exit;
+}
+
+foreach $file (<@ARGV>) {
+
+    if ($file =~ m/\.tex$/) {
+	($outfile = $file) =~ s/\.tex$/.txt/;
+
+	undef $/;
+	open (FILE, "$file") or die "Can't open $file: $!\n";
+	$text = <FILE>;
+	$text =~ s/\\\\//msg;
+	$text =~ s/^\s*\%.*\n//msg;
+	$text =~ s/\n/ /msg;
+	$text =~ s/ +/ /g;
+
+	open (OUTFILE,">$outfile") or die "Can't open $outfile: $!\n";
+	print "Converting $file to $outfile ...\n";
+	print OUTFILE $text."\n";
+    }
+
+}
+

### 2022-03-28 localizify.pl
Generaly clean up on Aisles Five; adjustments to localizify.pl, makerevtex4.pl

@@ -60,8 +60,12 @@ foreach $i (0..$#figures) {
     $fullfigures[$i] = $filenames[0];
     chomp($fullfigures[$i]);
 
-#    print "$fullfigures[$i]\n $localfiguresdir/$figure\n";
-    `cp $fullfigures[$i] $localfiguresdir/$figure`;
+    ##   print "cp $fullfigures[$i] $localfiguresdir/$figure;\n";
+    if (-e "$fullfigures[$i]") {
+	`cp $fullfigures[$i] $localfiguresdir/$figure`;
+    } else {
+	print "Did not find:\n$figures[$i]\n";
+    }
 }
 
 ##############################
@@ -88,8 +92,12 @@ foreach $i (0..$#inputfiles) {
     $inputfilenames = `find figures -follow -name $inputfile -print 2>/dev/null`;
     @inputfilenames = split("\n",$inputfilenames);
 
-#    print "$inputfilenames[0] $localinputsdir/\n";
-    `cp $inputfilenames[0] $localinputsdir/`;
+    ##    print "cp $inputfilenames[0] $localinputsdir/\n";
+    if (-e "$inputfilenames[0]") {
+	`cp $inputfilenames[0] $localinputsdir/`;
+    } else {
+	print "Did not find\n:$inputfiles[$i]\n";
+    }
     
 }
 

### 2022-03-28 makerevtex4.pl

@@ -1,10 +1,16 @@
 #!/usr/bin/perl
 
-# set up name in settings file
+## set up name in settings file
 `make-name-match-settingsfile.pl`;
 
-# localize figures and inputs (for better sharing)
+## localize figures and inputs (for better sharing)
 `./localizify.pl`;
+
+## local directory (for global recording of writing data)
+## works because paper/book directory names are in the format YYYY-MM-descriptive-tags
+## downfall: directory names may change but that represents a rebuilding
+$fulldir = `pwd`;
+($localdir = `pwd`) =~ s/^.*\///;
 
 foreach $argument (@ARGV) {
     if (($argument eq "-quick") or ($argument eq "-q"))
@@ -30,13 +36,13 @@ foreach $body (@bodies) {
     ($appendiexfilename = $body) =~ s/\.body\.tex$/\.appendix\.tex/;
     if (-e "$filename-revtex4.tex") {
 
-	# make local bibliography
-	# will need to be adjusted for more elaborate inputs
+	## make local bibliography
+	## will need to be adjusted for more elaborate inputs
 	$texfiles = join(" ",<$filename.*.tex>);
 	$command = "make-local-bibfile.pl $filename.bib $texfiles";
         system($command);
 
-	# make single line versions of title, abstract, logline
+	## make single line versions of title, abstract, logline
 	$command = "./singlelineify.pl $filename.title.tex\n";
         system($command);
 	$command = "./singlelineify.pl $filename.abs.tex\n";
@@ -47,13 +53,13 @@ foreach $body (@bodies) {
 	($logfilename = "log/progress-$body") =~ s/\.body\.tex$/-log/;
 	print "processing $filename...\n";
 
-	# create $filename-revtex4.settings.tex file
+	## create $filename-revtex4.settings.tex file
 	$settings = "$filename-revtex4.settings.tex";
 	open (SETTINGS,">$settings") or die "can't open $settings: $!\n";
 	print SETTINGS "\\newcommand{\\filenamebase}{$filename}\n";
 	close SETTINGS;
 
-	# check for postscript
+	## check for postscript
 	open (BODY, "$body") or die "can't open $body: $!\n";
 	undef $/;
 	$bodytex = <BODY>;
[…]

### 2023-03-07 correspondence/2022-05-31epj-data-science-response.body.tex
Many small improvements; It's been a while

\section*{Reply to Reviews}

\textbf{Submission:} EPDS-D-20-00186 \\
\textbf{Manuscript:} 
Allotaxonometry and rank-turbulence divergence: A universal instrument for comparing complex systems
\textbf{Authors:} Dodds et al.

\generalcomment{Assessment from the Associate Editor:}{
  The authors propose a very interesting method. We're sorry this review
  took a long time, but eventually we were able to secure three highly
  competent Reviewers. Reviewer 1 presents a few major reviewercomments
  and recommended a major revision. Upon closer inspection, however,
  those reviewercomments are mostly interpretational, but addressing
  them will make the manuscript more readily accessible to a broader
  readership. The reviewercomments of R1 are in line with the
  reviewercomments from Reviewer 2, who also suggests the authors make a
  few minor clarifications. Reviewer 3 raises no criticism. Taken
  together, all reviews are very positive and only minor clarifications
  are requested.
}

  We greatly thank the editor for the efforts and their overall positive evaluation,
  and we thank the reviewers for their helpful and enthusiastic feedback.

We have revised the manuscript to address all concerns, with individual responses below.
We have also generally improved the manuscript with some minor edits throughout,
as well as by marking up the allotaxonographs in Figures 1 and 2 to better facilitate our
explanation of their elements.
We have also adjusted some terminology for clarity,
and to remove unneeded attributions to individual scientists.
For example, we have largely replaced ``Zipf distributions'' with
the more
plainspoken ``type rankings''.

We also apologize for the delay in our response.

\reviewerheader{1}

\reviewercomment{
Reviewer \#1: Thank you for the opportunity to review this paper. It proposes new methods to compare either multiple systems or a single system over time, including graphs and a measure of divergence. The proposed methods are demonstrated on a variety of data examples. \\

Major Reviewercomments:\\

Can we use the rank-rank histogram to detect differences in the number of component types between the two systems? For example, can we use the histogram to figure out whether there are more species in 1985 vs. 2015 in Figure 3? 
}

As was presented in the manuscript, the total number of types was not made explicit.

\todo{Point to balances}

\todo{Add some text to paper about this.}

\todo{In future iterations, we could add the number of tokens (if counts are the measure), number of types, and number of exclusive types to the inset showing balances.}

\todo{Can see the imbalance from the histogram}

\todo{The numbers can be reported in text accompanying figures}

\reviewercomment{
Also, can we use the graph to gauge the difference in size between the component type of rank r in System 1 and the component type of System 2 of the same rank?
}

For RTD, we wanted to build a tool that only works with ranked data.
While in practice, we will have systems where sizes are typically known, the intent here
was to build a fully ranked-based instrument in the spirit of well established
rank-based statistical methods (e.g., the Spearman coefficient).

In our second paper on allotaxonometry, we introduce and explore probability-turbulence divergence,
where we make size (as measured by probability or rate) explicit.

We note that the allotaxonometer code does include the ability to compare systems where we have sizes that are not counts or probabilities or cannot be interpreted as relative fractions
in some way (e.g., heights of buildings).

This manuscript is also in review and can be found here: \url{https://arxiv.org/abs/2008.13078}

\todo{Zipf distributions comparing cities in 1750 to 2000 in the US}

\todo{Could compare normalized ones ...}

\todo{Best thing would be to overlay the two Zipf distributions. We considered ways of adding Zipf distributions to the plots but we felt that the results would be too complicated (and we acknowledge that the allotaxonographs are complicated enough as is)}

\reviewercomment{
How do we interpret the percentages in the "Balances" section at the bottom of Figure 1? Are these percentages unrelated to your alpha tuning parameter? 
}

The balances are independent of alpha, and indicate relative percentages
of tokens, types, and exclusive types.
\todo{We have added some text to make this more clear}
and
\todo{we have also expanded the balance section to included absolute numbers}.

\reviewercomment{
The paper states that the standard methods lack transparency and flexibility. It may be helpful to apply some standard methods in one of your case studies to demonstrate the differences from your method.
}

Yes, this is well spotted. We have separated out work where we compare
methods (e.g., JSD vs PTD) and that is still in development, and had left
residual guidance in the abstract.

We have removed two sentences from the abstract that misleadingly pointed in this direction.

\reviewercomment{
How sensitive is your method to small perturbations? For example, if you fixed the errors in the Berkshire Hathaway and DowDuPont market caps in the last case study, how much do the divergence and histogram change?
}

In this first paper, we have explored the effects of truncation only.
Generally, using ranks will give a more robust instrument, and is part of our motivation.

We have now noted in the manuscript the following:

"Small errors or perturbations of counts for a Zipf distribution 
will generally not affect the ranks of the dominant types."

\reviewercomment{
What is the computing time to calculate the divergence and generate the histograms in your four case studies?
}

Generally, the calculations and graphics require a few seconds of compute time.
The computational bottleneck in producing the RTD calculations is confined entirely to the counting of elements to produce ranks.
For the Twitter data, for example, enumerating $n$-grams for each day (20GB of compressed text) can take many CPUs running continuously to keep up with real time, with more than one million previously unseen types encountered on a given day.

\reviewercomment{

Minor Reviewercomments:

The terms "Zipf" and "1-gram" should be defined.
}

We have made the initial reference to Zipf more clear, connecting only to "Zipf's law".
We have now defined Zipf distribution and $n$-grams in the manuscript. 
We have replaced ``Zipf ranking'' with ``size ranking'' throughout, as we do not need
to name this simple kind of ranking after an individual.

Added to the manuscript:
`For ease of reference, we will refer to
``component rankings by decreasing size'' as ``size rankings.'''

Definition/clarification of 1-grams now appears in a couple of places, e.g.:

`For Twitter, subsampling $n$-grams---phrases containing $n$ words---allows for robust estimation of the rates of common $n$-grams but not rare ones.'

\reviewercomment{
In Paragraph 2 of Part B in the Introduction, it would be helpful to get a concrete example of a component type.
}

\todo{Do}

\reviewercomment{

Define the notation $r_(\tau,1)$ and $r_(\tau,2)$ , which seems to be first used in Part B of Section II. I believe they are the ranks of component τ in Systems 1 and 2, but this should be clarified.
}

\todo{Do}

\reviewercomment{
In Figure 1, the caption has substantial overlap with the main text. Removing some of the overlap would help highlight the new information.
}

It is our strong preference to make the captions reasonably complete in themselves. Readers may choose to work through the paper in different ways.

\reviewercomment{
In the first two paragraphs on page 6, some of the words that are described (such as "Harambe" and "voted") do not seem to be in Figure 1.
}

Thanks for pointing this out, we have adjusted the language describing Figure 1 to match words found at the edges of the histogram. `Gorilla' appears in the figure, so we have retained description of Harambe for context in the main text.

\todo{double check that remade figure contains words referenced in the caption}

\reviewercomment{
Could you give some intuition of why you chose $\alpha = 0$ for the BCI example? Also, what is the $D_{0;rand}^R$ term mentioned in that example?
}

We make the choice of $\alpha = 0$ by inspection.
In the appropriate supplementary PDF flipbook, the effect of tuning alpha can be observed.
Future work, which we discuss at the end of the paper, would seek to find
a method to determine an optimal alpha for specific comparisons.

% \rtdalphavarsystemsRankRandw$

\todo{Connect rand properly back to normalization and notation definition, or wherever}

Note that we have made the first appearance of our notation for RTD into a standalone equation (the first equation, shifting following equations up).

\reviewerheader{2}

\reviewercomment{
It was my pleasure to review this article introducing a new framework to compare heavy-tailed ranked lists.  Specifically, the work motivates and derives a novel divergence measure for ranked lists, and then illustrates its usefulness across four examples including: Twitter word frequency, baby name popularity, species abundance, and firm sizes.  
The presentation is rich and clear, elucidating the central framework and offering wonderful insights into the motivating examples and applications.  The problem this article addresses is relevant to a broad range of data science applications.
Therefore, I strongly recommend to publish the article in EPJ Data Science.
}

We thank Reviewer 2 for their generous feedback.

\reviewercomment{
Reviewercomments:
Even through the presentation is quite clear, it may benefit from an organization that extracts a few non-technical discussions from the rest of the text - i.e. a clearly identified general audience guide to allotaxonomy and the application of rank-turbulence.  This could support usage in fields outside of physics / data sciences.
}

Thank you.

We are authoring a blog post to accompany publication of the
manuscript that will advertise the utility of the instrument to other
disciplines.

We are also working on julia and python versions of the
allotaxonometer code (presently 6000 lines of Matlab) to encourage
broader adoption.

We and others have also used allotaxonographs in a number of other
papers, and we now cite these in the manuscript's conclusion.

%% nWe and others have also used allotaxonographs in a number of other papers, and we now cite these in the manuscript
%% \cite{gothard2021a,stupinksi2021a,minot2021a,ring2020a,alshaabi2021c,dodds2020n}

\reviewercomment{

Relatedly, it would be helpful if the authors could elaborate on
section IIE and the processes used to select different values of alpha
for the different applications.  I'm imagining the framework being
used by a non-expert who just keeps whatever default value of alpha is
encoded instead of exploring the full range - or on the flip side,
being overwhelmed with the many potential choices of alpha and not
understanding what features of the histograms are being accentuated by
the divergence in different limits. Can you distill a few guiding
principles or questions to help motivate an appropriate selection?
}

In the present work, we justify the choice of alpha by looking at the shape of distribution as a guide.

In the appropriate supplementary PDF flipbook, the effect of tuning alpha can be observed.

In future work, which we discuss at the end of the paper, we will seek
to find a method to determine an optimal alpha for specific
comparisons.

\reviewercomment{
The examples illustrating the effects of partial sampling were helpful, but it would be nice to see how the divergence measure itself changes with subsample size: does it converge or is it sensitive to the influx of different cross-ranks uncovered by larger subsample sizes?  Further, how does the divergence measure work in the presence of unbalanced list sizes, say when comparing the word usage of English tweets between countries?  I noticed that some imbalance likely appears in the baby names example.  Does the skewness of the underlying histogram bias the divergence and resulting understanding of turbulent names?
}

\todo{Performance under sampling: should obey whatever decay rate rank-ordered data of particular scaling law experiences, independent of this methodology.}

\todo{We've seen undersampled tweet data, looks too sparse near center, too many hapax words appearing in bar chart on right.}

\reviewercomment{
  The figures function as standalone illustrations of the concepts and enhance the intuition underlying the measure.  I especially like Figure 2, which manages to compress an impressive amount of information into a clear visualization.  Furthermore, the Supplemental Flipbooks are fantastic and I hope the publication of this article is accompanied by a Twitter campaign sharing these as gifs. I'm already excited to explore the allotaxonomy of my own research questions!  
}

Thank you!

As we noted in the paper, we have an online site for the papers, data, and code that we have produced
(and will produce) around allotaxonometry. There are various gifs on display:

\url{https://storylab.w3.uvm.edu/allotaxonometry/}

\reviewerheader{3}

\reviewercomment{
  Reviewer \#3: This is overall an excellent piece of research. The authors develop an innovative, solid method to compare rankings in systems, and demonstrate its use for word frequencies in social media. The methods are solid, the writing of very quality and the figures as well. It is without hesitation that I recommend this paper for publication as is.
}

We greatly, greatly appreciate Reviewer 3's extremely positive assessment. Thank you!

### 2023-03-07 localizify.pl

@@ -96,7 +96,7 @@ foreach $i (0..$#inputfiles) {
     if (-e "$inputfilenames[0]") {
 	`cp $inputfilenames[0] $localinputsdir/`;
     } else {
-	print "Did not find\n:$inputfiles[$i]\n";
+	print "Did not find:\n$inputfiles[$i]\n";
     }
     
 }

### 2023-03-07 makerevtex4.pl

@@ -109,7 +109,7 @@ foreach $body (@bodies) {
 	
 	undef $/;
 	$log = <TEXLOGFILE>;
-	$log =~ m/Output written on (.*\)\.)$/ms;
+	$log =~ m/Output written on (.*\)[\.]*)$/ms;
 	($line = $1) =~ s/\n//g;
 ##    print "$line\n";
 	$line =~ m/\((\d+) (pag.*?), (\d+) bytes\)/;
@@ -138,11 +138,11 @@ foreach $body (@bodies) {
 	    open (LOGFILE, ">>$logfilename") or die "Can't open $logfilename: $!\n";
 	    print LOGFILE "$timestamp $numtodos $numpages $numbytes $numlines $numwords $numchars\n";
 	    close LOGFILE;
-	    print "Data logged\n";
+	    print "\nData logged\n";
 	}
 	else
 	{
-	    print "No data logged\n";
+	    print "\n\nWarning! No data logged!\n\n";
 	}
 	
 	$citenum = `grep \"Warning--I didn\'t find a database entry for\" $filename-revtex4.blg | wc -l`;

### 2023-03-16 correspondence/2020-06-20cover-letter-epj-data-science.body.tex
Greatly improved response-to-reviewers template with example; restructured cover letter naming

%% editor: y.-y. ahn
%% sune lehmann
%% neil johnson
%% martin gerlach
%% simon dideo
%% aaron clauset
%% dashun wang
%% luis amaral

To the Editors at EPJ Data Science,

We are very pleased to submit our manuscript for consideration at
EPJ Data Science:

\textbf{``Allotaxonometry and rank-turbulence divergence: A universal instrument for comparing complex systems.''}

We introduce and frame the concept of allotaxonometry:
The comparison of complex systems which have broad
internal variation in component type and structure.

We argue against the mismeasurement of complexity by single
numbers, that we must instead put forward instruments and tools
that embrace complexity.
Just as we need dashboards to fly planes,
we need dynamic dashboards for comprehending complex systems in general.

We start from the longstanding but misleading observation that Zipf distributions
for language may vary little across texts.
We show that while the overall form of Zipf's distribution may be conserved,
the ordering of words (or n-grams) can vary greatly, a phenomenon we have
previously dubbed `lexical turbulence'.

For general complex systems,
we build a general `allotaxonometric instrument' that shows a map-like
histogram for Zipf distribution changes along with a ranked list
of the most `important' components according to whatever instrument
we use to measure and hence distinguish two systems (e.g., Shannon's entropy).

In our manuscript, we introduce `rank-turbulence divergence', a pragmatic, tunable
instrument for comparing any pair of complex systems where components can
be ranked according to some criterion
(e.g., frequency of words or species, sizes of cites or companies).

We provide a series of four main case studies in the paper, all comparing
systems evolving in time:
Word usage on Twitter,
species abundance,
baby name popularity,
and
market capitalization for publically traded companies in the US.

In online supplementary material, we expand on these
system-system comparions
with pdf flipbooks, and add more examples including
changes over time for causes of death in Hong Kong
and advertised job titles in the US.

See: \url{http://compstorylab.org/allotaxonometry/}.

Our paper is the first in a series of related papers which will
present allotaxonographs for probability-based measures such
as the Jensen-Shannon Divergence (and generalizations) as well
as our own probability-turbulence divergence.
As is, the figure-making code contains all of these options.

We note that we are in fact separately submitting our
second paper to EPJ Data Science:
\textbf{``Probability-turbulence divergence: A tunable
allotaxonometric instrument for comparing
heavy-tailed categorical distributions,''}
which is available online at \url{https://arxiv.org/abs/2008.13078}).

%% \textbf{1.}
%% \textbf{2.}
%% \textbf{3.}
%% \textbf{4.}

%% \textbf{}

We believe our paper will be of great benefit to a broad academic audience.

Finally, some possible reviewers we have listed are:
Simon Dideo,
Sune Lehman,
Y.-Y. Ahn,
Yu-Ru Lin,
Neil Johnson,
and
Johan Ugander.

We also suggest:
Jake Hofman,
Martin Gerlach,
Dashun Wang,
and
Luis Amaral.

We look forward to hearing of your decision.

\bigskip
\bigskip

Yours sincerely and on behalf of the manuscript's authors,

%% Professor
%% Director of the Vermont Complex Systems Center
%% Co-Director, Computational Story Lab
%% Vermont Advanced Computing Center
%% Department of Mathematics and Statistics
%% The University of Vermont

### 2023-03-16 correspondence/2022-05-31response-epj-data-science.body.tex
Greatly improved response-to-reviewers template with example; restructured cover letter naming

\section*{Response to Reviews}

\textbf{Submission:} EPDS-D-20-00186 \\
\textbf{Manuscript:} 
Allotaxonometry and rank-turbulence divergence: A universal instrument for comparing complex systems
\textbf{Authors:} Dodds et al.

\begin{editorcomment}[Assessment from the Associate Editor:]
  The authors propose a very interesting method. We're sorry this review
  took a long time, but eventually we were able to secure three highly
  competent Reviewers. Reviewer 1 presents a few major reviewer comments
  and recommended a major revision. Upon closer inspection, however,
  those reviewer comments are mostly interpretational, but addressing
  them will make the manuscript more readily accessible to a broader
  readership. The reviewer comments of Reviewer 1 are in line with the
  reviewer comments from Reviewer 2, who also suggests the authors make a
  few minor clarifications. Reviewer 3 raises no criticism. Taken
  together, all reviews are very positive and only minor clarifications
  are requested.
\end{editorcomment}

\subsection*{Overall remarks concerning our response to reviewers:}

\begin{enumerate}
\item 
  We deeply apologize for the delay in our response. The lead author
  has had a series of unfortunate maladies over a full year. This
  paper represents an enormous effort, and our response
  has required sustained time to
  engage with fully.
\item 
  We greatly thank the editor for their efforts and their overall positive evaluation.
\item 
  We greatly appreciate the helpful comments from their reviewers,
  and for their positive feedback.
  Their input has led to a number of substantive improvements.
\item 
  We have taken substantive time to revise the manuscript
  to properly address all concerns put forward by the reviewers, with individual responses below.
\item 
  For ease of reading, we have framed the comments from the reviewers in light blue boxes.
  We also show relevant edited excerpts from the updated manuscript in light gray boxes.
\item
  We have expanded our description of the allotaxonographs in Figs.~1 and 2.
  We have added overlaid dashed boundaries with labels to their many parts
  (10 parts for Fig.~1, and 6 parts for Fig.~2).
  We have extracted these (small) labeled graphical elements
  and then incorporated them in the text
  to better facilitate our explanation.
  We hope that this new structure will
  be acceptable for EPJ Data Science.
\item 
  We have also generally improved the manuscript with some minor edits throughout.
\item
  We note that our script for creating allotaxonographs also accommodates
  Probability Turbulence Divergence (PTD, described in a separate paper)
  and
  entropy-based divergences,
  e.g., Jenson-Shannon divergence (JSD) and a natural generalization.
\item
  We have added three more example allotaxonographs in response to reviewer's comments.
\item
  We have shared more code and data through
  the paper's Gitlab repository, and will aim to continue to build there.
  We intend that the paper's figures should all be rebuildable from scratch.
\item
  We have replaced Flipbook 3 which was a series of bare allotaxonographs
  for Twitter, rather than an intended series of allotaxonometric graphs.
\item
  We have also adjusted some terminology for clarity,
  and to remove unneeded attributions to individual scientists.
  For example, we have largely replaced ``Zipf distributions'' with
  the more
  plainspoken ``size rankings'', ``size-rank distributions'', and ``rankings''.
  (We still cite Zipf appropriately.)
\end{enumerate}

Overall, and while we have expended considerable time and effort
to respond fully and faithfully to the reviewers,
we have introduced no major changes in the manuscript.
As the editor indicated, we have found that only clarification
has been requested.

We note that that we received no comments about the mathematical
construction of rank-turbulence divergence, the core invention of
the paper.
As we have sought to demonstrate with our paper,
we have extensively demonstrated how allotaxonographs and RTD perform
for a range of disparate systems.
The core instrument remains the same as do all original figures,
except for Fig.~10 (was Fig.~7) which now includes more information.

We hope that our revised manuscript is suitable for publication in
EPJ Data Science.

\reviewerheader{1}

\begin{reviewercomment}
  Reviewer \#1: Thank you for the opportunity to
  review this paper. It proposes new methods to compare either
  multiple systems or a single system over time, including graphs and
  a measure of divergence. The proposed methods are demonstrated on a
  variety of data examples.

  Major Reviewer comments:

  Can we use the rank-rank histogram to detect differences in the number
  of component types between the two systems? For example, can we use
  the histogram to figure out whether there are more species in 1985
  vs. 2015 in Figure 3?
\end{reviewercomment}

As is, we do not make the total number of types explicit in the
rank-rank histograms. We have intended that our instrument will
work on bare rankings.

To further help make clear that rank-divergence works purely on rankings we have adjusted
a paragraph from the introduction to read:
\begin{excerpt}
  We will consider systems where
  each component type $\elementsymbol$ has at
  least one measurable---and hence rankable---``size'' $\sizesymbol_{\elementsymbol}$.
  We make clear that we may have no knowledge of these component sizes---we may only have rankings (e.g., book rankings provided by a seller,
  but with numbers of sales withheld).
  Our core instrument functions only on rankings,
  incorporating size data (if available) for minor diagnostics.
\end{excerpt}

That said, and as we note below, various absolute numbers
are returned within the Matlab environment when an allotaxonograph
is generated (and can be computed separately).

We provide three points.

1.
In general, our rank-rank histograms will strongly visually indicate if two
systems are imbalanced in their respective number of distinct types
(on this topic, see also our response to Question 5 from Reviewer 2).

For example, in Figs. 4 and 5, the allotaxonographs for baby girl and boy names
show a distinct line separated from the main histogram only on the 2018 (right hand) side.
Names that fall along this separated line are names that
appear in only 2018 (`2018-exclusive types').
So without the exact numbers of types, a user (with some familiarity)
will immediately see that there are more types
of baby names in 2018.

By contrast, the allotaxonograph for Twitter in Fig. 1 shows a fairly balanced number of types.
On both sides, we see two similarly separated lines from the main histogram, corresponding to 1-grams
that appear 0 times in the comparison corpus
(bottommost outer separated line) or 1 time (inner separated line).

2.
The script for generating allotaxonographs (figallotaxonometer9000.m in Matlab provided in the paper's Gitlab repository)
returns some diagnostics
including a set of absolute numbers (this output can be ignored).

Here's an example of part of the output,
using the 1968--2018 baby girl name comparison from Fig. 4,
and with comments added:
\begin{lstlisting}[language=Matlab]
  details = 
  struct with fields:

  Nunion: 22758    %% # of types in union of both systems
  Nshared: 3552     %% # of types in intersection of both systems
  N1: 8195     %% # of types in system 1 (baby girl names, 1968)
  N2: 18115    %% # of types in system 2 (baby girl names, 1968)
  N1exclusive: 4643     %% # of exclusive types in system 1
  N2exclusive: 14563    %% # of exclusive types in system 2
  totalcounts1: 1640302  %% Total # of girls in system 1 (token count)
  totalcounts2: 1696917  %% Total # of girls in system 2 (token count)
\end{lstlisting}

3.
(Note: Some form of the text below now appears in the manuscript.
See our answer to reviewer 1's third comment below.)

The balances at the bottom right of each allotaxonograph
(now labeled 1G in Fig. 1) present various
\textit{relative} comparisons derived from absolute numbers of types,
and if available, `sizes' of those types.
Size is often a form of counting (words, baby names) but could
be anything and may not be normalizable (e.g., market cap for corporations).

Again for baby girl names in Fig. 4, here's the extracted display of balances:
\begin{center}
  \includegraphics[width=0.3\textwidth]{figallotax-baby-girl-name-balances.pdf}
\end{center}

Each allotaxonograph will display two or three bars.
Per the reviewer's interest,
comparisons for the numbers of types in each system
are recorded in the bottom two bars,
and the top one, described below, is rendered optionally.

Counts of names (and more generally, sizes of types)
are not incorporated in the two bottom bars.

The second balance from the bottom shows the fraction of types in each system
as a percentage of the union of types from both systems.
For baby girl names, the computation is:
\begin{lstlisting}
  >> details.N1 / details.Nunion * 100 
  8195 / 22758 * 100
  ans =
  36.0093
  >> details.N2 / details* 100.Nunion * 100 
  18115 / 22758 * 100
  ans =
  79.5984
\end{lstlisting}
These numbers correspond to the middle of the balances: 36.0\% versus 79.6\%.

The bottom balance shows that given each side's set of types,
what percentage appears only on that side.
The computation is:
\begin{lstlisting}
  >>     details.N1exclusive / details.N1 * 100
  4643 / 8195 * 100
  ans =
  56.6565
  >>     details.N2exclusive / details.N2 * 100
  14563 / 18115 * 100
  ans =
  80.3919
\end{lstlisting}
And these numbers are shown in the bottom bar as 56.7\% and 80.4\%.

Lastly, a third bar will appear on top if we have measures of type size.
For baby names, we have counts and the computation is:
\begin{lstlisting}
  >> sumtc = details.totalcounts1 + details.totalcounts2  
  sumtc =
  3337219
  >> details.totalcounts1 / sumtc * 100                  
  1640302 / 3337219 * 100
  ans =
  49.1518
  >> details.totalcounts2 / sumtc * 100
  1696917 / 3337219 * 100
  ans =
  50.8482
\end{lstlisting}
The topmost of the three bars shows the relative balances for the two years:
49.2\% versus 50.8\%.

What the top balance bar shows is that the number of baby girls registered with social security has
increased only very slightly from 1968 to 2018.
The expansion of the `lexicon' of baby names is thus a strong one
and not an artifact of increases in sheer numbers.
We discuss this further in the manuscript.

We have added the following paragraph to the baby names secton:
\begin{excerpt}
  We emphasize that the balance indicators are for baby names
  appearing at least fives times. For our present work, and in
  attempting to maintain uniformity across allotaxonographs, we do not
  attempt to adjust for names appearing less than 5 times, though this
  would be possible for the topmost balance for total counts given we
  have that information separately. Clearly the balance values would
  shift if we had complete data sets for baby names; estimating errors
  for these estimates would be meaningful future work.
\end{excerpt}

%% For example, for baby girl names appearing at least 5 times in a year in the US,
%% there were 1,640,302 recorded in 1968 and 1,696,917 in 2018 (appearing means being
%% registered with Social Security).
%% 
%% Strong separation in the histogram for names that only appear in 2018 relative to 1968.
%% 
%% This means that we do 
%% 
%% In fact, there is only one optional 
%% 
%% This information is an figure-building code 
%% 
%% The top bar is optional and will only be present if a measure of ``size'' of each type is provided
%% along with the ranks.

%% \done{Point to balances}

%% \done{Add some text to paper about this.}

%% In future iterations, we could add the number of tokens
%% (if counts are the measure), number of types,
%% and number of exclusive types to the inset showing balances.
%% 
%% Add to discussion at end.
%% 
%% We note that the baby name allotaxonographs compare only ranked types.
%% 
%% The balance 
%% 
%% omitting the tail of any names that appeared less than 5 times in a year.
%% We do however have the total number of names for each year.

%% Thus, the three balances reflect the ranked names only

%% \done{Can see the imbalance from the histogram}

%% \todo{The numbers can be reported in text accompanying figures}

\begin{reviewercomment}
  Also, can we use the graph to gauge the difference in size
  between the component type of rank r in System 1 and
  the component type of System 2 of the same rank?
\end{reviewercomment}

As per our response to the question above,
for RTD, we wanted to build a tool that only works with ranked data.
While in practice, we will have systems where type `sizes' are known, the intent here
was to build a fully ranked-based instrument in the spirit of well established
rank-based statistical methods (e.g., the Spearman coefficient).

We add two points:

1. The main allotaxonometer script can produce allotaxonographs that
compare by only type size. These outputs will show only an allotaxonograph
with an annotated histogram and no overlaying measure of difference.

2. In our second paper on allotaxonometry, we introduce and explore
probability-turbulence divergence (PTD),
where we make size (as measured by probability or rate) explicit.
We note that the allotaxonometer code does include the ability to
compare systems where we have sizes that are not counts or
probabilities or cannot be interpreted as relative fractions in some
way (e.g., heights of buildings).
We construct PTD with a tunable parameter $0 \le \alpha < \infty$
in a similar fashion to RTD.
We show that PTD has different
limiting behaviors from RTD,
and that it matches up with a collection of existence
difference measures
including
$L^{(p)}$ norms,
the S{\o}rensen-Dice coefficient
(the $F_1$ statistic),
and 
the Hellinger distance.

We add that we prefer PTD over the commonly used Jenson-Shannon divergence (JSD),
which we have a separate article about, not yet on the arXiv.
JSD creates a mixture of two systems which does not exist
(half Moby Dick, half Pride and Prejudice for example).
PTD works as a true difference and is what one would want
in constructing a form of ``type calculus'' (``lexical calculus'' for text corpora).
Our allotaxonograph script nevertheless will produce JSD comparison plots,
along with a particular generalized entropy divergence.

The PTD manuscript can be found here: \url{https://arxiv.org/abs/2008.13078}

We note that we had mentioned type calculus and lexical calculus in the conclusion:
\begin{excerpt}
  Allotaxonometry may be viewed as part of a larger analytic
  framework of ``lexical calculus'' and, more generally, ``type calculus.''
\end{excerpt}
We have now added ``type calculus'' to the abstract.

\begin{reviewercomment}
  How do we interpret the percentages in the ``Balances'' section at the bottom of Figure 1? Are these percentages unrelated to your alpha tuning parameter? 
\end{reviewercomment}

For an overview of the Balance percentages,
we point back to our response to Reviewer 1's question 1 above.

We add here that the balances are independent of alpha, and indicate relative percentages
of type sizes (e.g., word counts), types, and exclusive types.

As we mention in the overview for our reply,
we have outlined 10 distinct cutouts of the allotaxonograph
in Fig.~1 to help with our discussion.
We have taken these cutouts and added them to
the text to, we hope, greatly help the reader with being
able to comprehend the instrument.

We describe balances in Sec IIB: ``Rank-Rank Histograms for Basic Allotaxonomy''.
The cutout for Fig.~1 for balances is Fig.~1G and can be found looking through
the subsections of Sec IIB.

We reproduce the relevant text here:

\begin{center}
  \includegraphics[width=0.4\columnwidth]{allotaxonograph_balances2_flattened-crop.pdf}
\end{center}
\begin{excerpt}
  \textbf{Fig.~1G}:
  For all allotaxonographs, we show balances at the bottom right
  of the rank-rank histogram.
  Two kinds of exclusive type comparisons for the numbers of types in each system
  are recorded in the bottom two bars,
  while the top bar conveys information about type sizes.
  The top bar is rendered only when sizes are known
  (this is the only part of the allotaxonographs that is not
  determined purely from type rankings).
  In the present paper, we work from datasets with known type sizes,
  and all allotaxonographs will have all three bars.
  All balances show normalized quantities rather than absolute numbers

  As we will see, these three balances can vary greatly across system comparisons.

  The top balance bar shows the relative balance of the two systems' sizes, if known.
  For our Twitter example, 
  this top bar shows the breakdown of total counts of 1-grams (type size)
  between the two dates at 59.9\% and 40.1\%.
  We thus see that the election generated considerably more tweets than Charlottesville.

  The middle balance bar shows the fraction of types in each system
  as a percentage of the union of types from both systems.
  For Twitter, we have that of all words in the combined lexicon for the two days combined,
  just over 60\% appear on each of the two days (63.2\% and 61.6\%).

  The bottom balance bar shows that given a system's set of types,
  what percentage of those types appear only in that system.
  For the Twitter example, we take the  separate lexicons for each day,
  and find that around 60\% of words are exclusive for both days,
  further giving a sense of strong turnover (60.8\% and 59.8\%).

  We add that the script for generating allotaxonographs
  (figallotaxonometer9000.m in Matlab provided in the paper's Gitlab repository)
  returns some diagnostics
  including the underlying numbers used to compute the above balances.
\end{excerpt}

%% \noway{we have also expanded the balance section to included absolute numbers}.

\begin{reviewercomment}
  The paper states that the standard methods lack transparency and flexibility.
  It may be helpful to apply some standard methods in one of your case studies to demonstrate the differences from your method.
\end{reviewercomment}

In preparing this manuscript,
we had separated out work where we compare methods (e.g., JSD vs RTD vs PTD),
but we had left residual guidance in the abstract.
This work is still being written up (as mentioned above).

We have removed two sentences from the abstract that misleadingly pointed
in the direction of other measures being deficient.
We have also removed any similar statements in the main text.

\begin{reviewercomment}
  How sensitive is your method to small perturbations? For example,
  if you fixed the errors in the Berkshire Hathaway and DowDuPont market
  caps in the last case study, how much do the divergence and histogram change?
\end{reviewercomment}

In general, using ranks will give a more robust instrument,
and is a key part of our overall motivation.

From the paper:
\begin{excerpt}
  Fourth, rank orderings
  potentially allow for powerful and robust non-parametric statistical measures
  such as Spearman's rank correlation coefficient.
  All told,
  while in moving from sizes to rankings
  we may
  trade information for some simplification,
  we still preserve a great deal of meaningful structure.
\end{excerpt}

We understand the reviewer's point, however,
that we could perform a deep analysis of sensitivity.
In this foundational paper, we have explored the effects of truncation only,
and reserve further work for later (by us, and, hopefully, others as well).
We believed (and continue to believe) that
the few small clear errors in the market cap database were worth
leaving in to show how outlier kinds of errors are easily observed
with allotaxonographs.

The above said, we corrected the Berkshire Hathaway and DowDuPont data
and recomputed $\rtd{1/3}$, finding it be unchanged to three significant figures
at 0.411.
The adjusted allotaxonograph (which we have added to the paper, and is included in our response below)
now make Berkshire Hathaway clearly ascendent
on the right side, as it was ranked 5th overall by market cap at the end of 2018.

Per the above, we have also added the following to the manuscript:

\begin{excerpt}
  The dataset for market caps does
  have some missing and erroneous data.
  DowDuPont's market cap for the last quarter of 2018 is absent
  and is consequently shown to have plummeted from
  a rank of
  $\sizerank$=91 in 2007
  to equal-to-last in 2018
  Berkshire Hathaway's
  market cap is clearly misrecorded
  for the last three
  quarters of the dataset
  (apparently dropping from
  \$528.33B
  to
  \$0.34B at the end of 2018).

  We take the opportunity to perform a small test of the sensitivity of rank-turbulence divergence
  by correcting the data for these two companies.
  For DowDuPont, with further sourcing, we find
  the year-end 2007 and 2018 market caps were reported as
  \$37.06B
  and
  \$121.34B,
  and for
  Berkshire Hathaway,
  \$149.56B
  and
  \$502.37B.
  Upon making these corrections, we first find again
  that $\rtd{1/3}$=0.411, unchanged to three decimal places.
  both.
  In the corrected allotaxonograph (Fig.~9),
  Berkshire Hathaway's location shifts to the right
  side of the histogram ($\sizerank$=38$\rightarrow$5)
  and is now listed as the 7th overall strongest contribution
  for $\rtd{1/3}$.
  DowDuPont no longer makes the top 40 of
  the list of contributions.
  While these two changes are dramatic,
  the remainder of the allotaxonograph
  remains essentially identical.

  Nevertheless, we have chosen to leave such errors in
  Fig.~8
  to help demonstrate the importance of using a rich,
  graphical allotaxonometric instrument.
  With a naive measurement of divergence, we would easily miss problematic data points.
  Evidently, and beyond our present paper's interests,
  for any further investigations,
  these two errors suggest that considerable effort should
  be made to further clean the market cap dataset
\end{excerpt}

\clearpage

For ease of comparison, we include both the
uncorrected and corrected market cap allotaxonographs below (Figs.~8 and 9)
and the corrected one below.

\includegraphics[width=\textwidth]{figallotaxonometer9000-siblis_alphasweep_05-2007-Q4_2018-Q4_marketcaps004_noname.pdf}

\includegraphics[width=\textwidth]{figallotaxonometer9000-siblis_alphasweep_06-2007-Q4_2018-Q4_marketcaps004_corrected_noname.pdf}

\clearpage

Caption for second figure:
\begin{excerpt}
  \textbf{Allotaxonometric comparison of publicly traded US companies
    in 2007 and 2018 by fourth quarter market capitalization
    with corrections for Berkshire Hathaway and DowDuPont Inc.}
  To be compared with Fig.~8.
  The 2018 market caps for both these companies were recorded incorrectly
  in the original data set.
  The revised allotaxonograph shows Berkshire Hathaway---which in fact rose to
  $\sizerank$=5 in 2018---now appears prominently
  on the right hand side of the histogram and the contribution list.
  DowDuPont Inc's corrected rank for 2018 means that it did not
  contribute as strongly, and now no longer appears in the top 40 contribution list.
  Overall, the allotaxonograph is largely identical to that
  of Fig.~\ref{fig:rankturbdiv.allotaxonometer9000-siblis_2007-Q4_2018-Q4_marketcaps004}.
  The rank-divergence is unchanged at $\rtd{1/3}=0.411$.
\end{excerpt}

\begin{reviewercomment}
  What is the computing time to calculate the divergence and generate
  the histograms in your four case studies?
\end{reviewercomment}

Generally, the calculations and graphics require a few seconds of
compute time.
In our experience, the only speed limitation arises
when loading massive datasets (e.g., Twitter).
Separate from our analyses, computational preparation of such large datasets
is where we see delays.
For the Twitter data, for example, enumerating $n$-grams for each day
(20GB of compressed text) can take many CPUs running continuously to
keep up with real time, with more than one million previously unseen
types encountered on a given day.

The computational bottleneck in producing the RTD
calculations is confined entirely to the counting of elements to
produce ranks.

\begin{reviewercomment}

  Minor Reviewer 1 comments:

  \bigskip

  The terms ``Zipf'' and ``1-gram'' should be defined.

\end{reviewercomment}

We have made the initial reference to Zipf more clear, connecting only to ``Zipf's law''.
We have also now defined Zipf distribution and $n$-grams in the manuscript. 
We have replaced ``Zipf ranking'' with ``size ranking'' throughout, as we do not need
to name this simple kind of ranking after an individual
(we of course cite Zipf and others where appropriate).

Added to the manuscript:
\begin{excerpt}
  Here, we will refer to such
  ``a component ranking by decreasing size'' as a ``size ranking''
  or, for brevity and given our paper's framing, simply a ``ranking''.
\end{excerpt}

In the main text, we have added this definition/clarification of $n$-grams when the term first appears:

\begin{excerpt}
  For Twitter, subsampling $n$-grams---phrases containing $n$ consecutive words and/or other text elements---allows for robust estimation of the rates of common $n$-grams but not rare ones.
\end{excerpt}

And in the caption of the Fig.~1:

\begin{excerpt}
  Words are extracted first as 1-grams (contiguous sets of non-whitespace characters)
  from tweets identified as English \ldots
\end{excerpt}

\begin{reviewercomment}
  In Paragraph 2 of Part B in the Introduction, it would be helpful to
  get a concrete example of a component type.
\end{reviewercomment}

We have added an explanatory paragraph at the start of section B in the introduction,
as well as pointed to this when we first mention size.

In ``Sec. I. Introduction A. Instruments that capture complexity'', we now have:
\begin{excerpt}
  But for complex phenomena made up of
  a great many types of components
  of greatly varying `size'
  (clarified below in Sec.~I B)---language, ecologies,
  stock markets---we
  must confront two major problems with our dashboards of simple instruments [2].
\end{excerpt}
We have put size in quotes to indicate it should be construed more generally, or at least,
that the reader will need to see our description.

Then in ``Sec. B. Size rankings, Zipf’s law, and rank turbulence'' (we have removed ``Zipf distributions'' from the title of this section), the section begins with:

\begin{excerpt}
  While the instrument we develop here will have broader application,
  its construction focuses on two regular features of complex systems:
  Heavy-tailed `size-rank distributions' (rather than laws),
  and
  what we will call `rank turbulence'---a phenomenon of system-system comparison.
  We describe and discuss these two common signatures of complex systems in turn.

  We will consider systems where
  each component type $\elementsymbol$ has at
  least one measurable---and hence rankable---`size' $\sizesymbol_{\elementsymbol}$.
  To help make clear what we mean by component types and component sizes,
  we list some examples in the realms of language, ecology, and stock markets.

  \begin{itemize}[leftmargin=70pt]
  \item[Example types:]
    `the', `love', and `spork'.
  \item[Type size:]
    Number of times a word appears in a book.
  \end{itemize}
  In ``Pride and Prejudice'', for example,
  the corresponding counts are
  %% 498, %% !
  4,058, %% the
  90, %% love
  and
  0. %% spork
  In linguistics, the component type and component size distinction is
  referred to as types and tokens [10],
  while more abstractly in semiotics, we have the signifier and the signified.

  \begin{itemize}[leftmargin=70pt]
  \item[Example type:]
    The species Ornithorhynchus anatinus, the platypus.\\
  \item[Type size:]
    The number of platypuses (`instances' of the species) living in Australia in the wild.\\
  \end{itemize}
  Given an ecology, species may be ranked by their population numbers.

  \begin{itemize}[leftmargin=70pt]
  \item[Example types:]
    The publicly traded companies of Apple and Microsoft.
  \item[Type size:]
    Market capitalization.
  \end{itemize}
  Apple and Microsoft may be viewed as components of the publicly-owned corporate world.
  The sizes of corporations may be broken down into many rankable dimensions
  such as annual revenue or number of employees worldwide.
  Market capitalization represents a kind of current collective belief
  in terms of money.

  The three examples above show some of the range of what size can mean, and why we
  cannot readily lift terminology out from one domain to apply to all.
  Size for a word in a corpus means the number of indistinguishable instances of that word
  (many identical entites---tokens);
  size for species means the number of `biological replications' of an individual type
  (many genetically similar entities of varying ages);
  and
  size for a corporation means a monetary value (one entity).

  %% grep here:
  %% ~/work/stories/2021-02ousiometer-books/data/2020-08gutenberg-jane-austen/narrativetimeseries
  %% grep ^the\$ pride_and_prejudice_narrativetimeseries.txt | wc

  Some further examples of component size
  are
  %% counts (e.g., of words in text corpora),
  rate (e.g., appearance of words in streaming text),
  physical dimensions (e.g., typical animal length or weight for a species),
  social popularity (e.g., number of plays of a song)
  scoring in sports by individual players,
  and so on.

  We make clear that we may have no knowledge of
  the underlying component sizes---we may only have rankings
  (e.g., book rankings provided by a seller,
  but with numbers of sales withheld).
  Our core instrument functions only on rankings,
  incorporating size data (if known) for minor diagnostics.
\end{excerpt}

\begin{reviewercomment}
  Define the notation $r_{\tau,1}$ and $r_{\tau,2}$ , which seems to be
  first used in Part B of Section II. I believe they are the ranks of
  component $\tau$ in Systems 1 and 2, but this should be clarified.
\end{reviewercomment}

The reviewer reads the notation correctly, even though, as the reviewer notes,
we did not formally provide a definition.

To correct this oversight, we have added the following sentence as
a standalone paragraph (now the third paragraph of Sec. II B):

\begin{excerpt}
In general, we will denote the rank of type
$\elementsymbol$
in system $\systema$
by
$\sizerank_{\elementsymbol,\indexaraw}$,
and
its rank
in system $\systemb$
by
$\sizerank_{\elementsymbol,\indexbraw}$.
\end{excerpt}

\begin{reviewercomment}
  In Figure 1, the caption has substantial overlap with the main
  text. Removing some of the overlap would help highlight the new
  information.
\end{reviewercomment}

It is our firm preference to make the captions reasonably complete
in themselves.
Readers may choose to work through the paper in different ways.

\begin{reviewercomment}
  In the first two paragraphs on page 6, some of the words that are
  described (such as ``Harambe'' and ``voted") do not seem to be in
  Figure 1.
\end{reviewercomment}

Many thanks for pointing this out---we appreciate the reviewer's
careful reading.

Generally speaking for rarer types, each run of the allotaxonometer
script may have different annotations automatically rendered.
There are a number of reasons, mainly if annotation spacing is changed,
or if a particular rank-rank pair is held by more than one type.
In the latter case, the annotated type will be randomly chosen,
and for some comparisons this means the rendered allotaxonometer will
always be different.
For example, for two days on Twitter in our data et,
the number of 1-grams that appear once on one day and not on the other
run into the millions.

We have adjusted the language describing
Fig.~1 to match words found at the edges of the histogram. `Gorilla'
appears in the figure, so we have been able to retain a description of Harambe for
context in the main text.

%% \done{double check that remade figure contains words referenced in the caption}

\begin{reviewercomment}
  Could you give some intuition of why you chose $\alpha = 0$ for the
  BCI example?
\end{reviewercomment}

We make the choice of $\alpha = 0$ by inspection.
In the appropriate supplementary PDF flipbook, the effect of tuning alpha can be observed.
Future work, which we discuss at the end of the paper, would seek to find
a method to determine an optimal alpha for specific comparisons.

A secondary aspect of the choice to use $\alpha = 0$ is simply to demonstrate
how such an allotaxonograph looks.
The BCI example, with its limited rank
turbulence and vertical shaping within the histogram, provides 
a good instance where $\alpha = 0$ is at least visually reasonable.
A choice of $\alpha = 0$
would make for a poor fit for the Twitter example (and for written texts in general).

We have revised and expanded the section reference by Reviewer 1.

\begin{excerpt}
  We numerically compare the 1985 and 2015 distributions by applying
  rank-turbulence divergence with $\alpha = 0$,
  finding $\rtdalphavarsystemsRank{0} = 0.077$.
  By inspection, we choose $\alpha=0$ here
  because of the match of the histogram
  with the verticality of the contour lines
  (we address optimal selection of $\alpha$ in our concluding remarks).
  The nature of the BCI example affords us an opportunity to
  demonstrate the limit of $\alpha=0$ for allotaxonometry,
  and is a secondary reason for including an example from ecology.

  In Flipbook \flipbooktrees, we show how the allotaxonometer
  performs with $\alpha$ varying away from 0 to $\infty$.
  The visual match on the contour lines continuously degrades.
\end{excerpt}

\begin{reviewercomment}
  Also, what is the $D_{0;\,\textnormal{rand}}^{\textnormal{R}}$ term
  mentioned in that example?
\end{reviewercomment}

We have now explained what we mean by a randomized version,
and explicitly connected this explanation to the notation.

The section (which directly continues on from the excerpt above)
now reads:
\begin{excerpt}
  The BCI example's histogram is far 
  from what we would expect of randomized systems
  (Fig.~1I).
  To see how RTD quantifies the difference between an observed system the randomized version,
  we construct a set of pairs of randomized systems,
  measuring rank-turbulence divergence for each.
  We do this by randomly permuting the species names within each system
  while leaving species counts fixed,
  thereby keeping the size-rank distributions the same.
  We can perform such a randomization for any system-system comparison
  (and we do so below again for baby names).

  We denote the average randomized divergence for two rankings 
  $\bigrank_{\indexaraw}$
  and
  $\bigrank_{\indexbraw}$
  as:
  \begin{gather}
    \rtdalphavarsystemsRankRand{\alpha}.
    \label{eq:rankturbdiv.randomizeddivergence}
  \end{gather}

  For the BCI example, we find that the score
  $\rtdalphavarsystemsRank{0} = 0.077$
  is well short of the randomized equivalent
  of
  $\rtdalphavarsystemsRankRand{0} = 0.376$
  (average of 100 randomizations; standard deviation $\sigma$=0.012).
\end{excerpt}

The quantity $\rtdalphavarsystemsRankRand{0}$
appears explicitly in the paper twice,
once in the BCI section (Sec. II C),
and once in the baby girl name section (Sec II D).

We have also made the first appearance of our general notation for RTD
a standalone equation (now Eq. 1).

\begin{excerpt}
  \begin{gather}
    \tag{1}
    \rtdalphasystemsOmega.
    \label{eq:rankturbdiv.rankturbdiv_defn}
  \end{gather}
\end{excerpt}

%% \done{Connect rand properly back to normalization and notation definition, or wherever}

\reviewerheader{2}

\begin{reviewercomment}
  It was my pleasure to review this article introducing a new
  framework to compare heavy-tailed ranked lists.  Specifically, the
  work motivates and derives a novel divergence measure for ranked
  lists, and then illustrates its usefulness across four examples
  including: Twitter word frequency, baby name popularity, species
  abundance, and firm sizes.

  The presentation is rich and clear, elucidating the central
  framework and offering wonderful insights into the motivating
  examples and applications.  The problem this article addresses is
  relevant to a broad range of data science applications.

  Therefore, I strongly recommend to publish the article in EPJ Data
  Science.
\end{reviewercomment}

We are delighted by Reviewer 2's very positive feedback.

\begin{reviewercomment}
  Reviewer comments: Even through the presentation is quite clear, it
  may benefit from an organization that extracts a few non-technical
  discussions from the rest of the text - i.e. a clearly identified
  general audience guide to allotaxonomy and the application of
  rank-turbulence.  This could support usage in fields outside of
  physics / data sciences.
\end{reviewercomment}

Thank you.

We will author a blog post and a Twitter thread to accompany
formal publication of the
manuscript that will advertise the utility of the instrument to other
disciplines.

We are also working on a Python and hopefully Julia version of the
allotaxonometer code (the original code was developed in Matlab) to
enable and encourage broader adoption.

We and others have also used allotaxonographs in a number of other
papers, and we now cite some of these in the manuscript's conclusion.

Our intent is that our foundational paper be thorough and detailed,
and will serve as a reference text, and we hope that the examples are illuminating
and inspirational.

\begin{reviewercomment}
  Relatedly, it would be helpful if the authors could elaborate on
  section IIE and the processes used to select different values of alpha
  for the different applications.  I'm imagining the framework being
  used by a non-expert who just keeps whatever default value of alpha is
  encoded instead of exploring the full range - or on the flip side,
  being overwhelmed with the many potential choices of alpha and not
  understanding what features of the histograms are being accentuated by
  the divergence in different limits. Can you distill a few guiding
  principles or questions to help motivate an appropriate selection?
\end{reviewercomment}

We have improved our discussion of the choice of $\alpha$ in the conclusion,
as well as given this discussion its own subsection (see excerpt below).

In this initial work, we justify the choice of alpha by looking at the shape of
rank-rank histogram as a guide.
The rank-rank histogram also immediately shows us when type (or lexical) turbulence is not
present (e.g., when the type rankings for the two systems are random shufflings of each other).

In short, our advice for readers is to use the visual help of the allotaxonographs.
In some of the supplementary PDF flipbooks, the effect of tuning alpha can be observed,
and readers can see how RTD behaves when the fit is good or bad.

In follow up work that has proved difficult, and which we discuss at the end of the paper, 
we  hope to identify a computational method to determine an optimal alpha for specific
comparisons.

\begin{excerpt}
  \subsection*{VI~C. Determination of $\alpha$}
  In our initial work, we have made the choice of the tuning parameter for
  rank-turbulence divergence, $\alpha$, a visually guided one.
  The user gains much from inspecting the rank-rank histogram alone,
  and, in our experience, is then readily able to choose an $\alpha$ for
  which the allotaxonometric contour lines best match the form of the histogram.
  A visually guided choice will be sufficient in cases of comparing two or a small
  number of systems.

  When rank turbulence presents as a scaling law---which is regularly the case
  for text corpora (e.g., Twitter, books)---we would want to be able to determine an optimal $\alpha$.
  While for generalized entropy approaches for single systems, the limit of
  linear scaling and Shannon's entropy demarcate the boundary
  between accentuating the common or the rare [7, 33, 53, 54], 
  we have found that for system comparisons, the optimal value of $\alpha$, if it exists,
  is dependent on the pair of systems being compared---there is no universal value.

  Generally speaking for text corpora, as two texts become more distinct,
  the optimal $\alpha$ increases.
  This is particularly clear for a corpus like
  Twitter, for example, where rank turbulence increases with the
  calendrical separation of two dates being studied.

  We have left open the
  possibility of an analytic
  connection between the rank-turbulence scaling described
  at the end of Sec. I B,
  and, to the extent that well-defined scaling is present,
  with an optimal $\alpha$ for rank-turbulence divergence.

  Even with an optimization method for determining $\alpha$, we urge readers to
  always look at the visuals provided by our allotaxonographs---the maps---for confirmation of fit.
\end{excerpt}

\begin{reviewercomment}
  The examples illustrating the effects of partial sampling were
  helpful, but it would be nice to see how the divergence measure
  itself changes with subsample size: does it converge or is it
  sensitive to the influx of different cross-ranks uncovered by larger
  subsample sizes?
\end{reviewercomment}

First, for each of the five main case studies,
we do in fact show the full allotaxonographs in the series of
Flipbooks S10, S11, S12, S13, and S14 in the
supplementary material.

From Sec IV. Guide to Flipbooks,
here is our description of these flipbooks:
\begin{excerpt}
  \textbf{Flipbook \flipbooktwittertrunc---Word use on Twitter, truncated:}
  Full series of allotaxonographs 
  corresponding to histograms of
  row 1 in Fig.~7 with $\alpha=1/3$.

  \smallskip
  \textbf{Flipbook \flipbooktreestrunc---Tree species abundance, truncated:}
  Full series of allotaxonographs 
  corresponding to histograms of
  row 2 in Fig.~7 with $\alpha=0$.

  \smallskip
  \textbf{Flipbook \flipbookgirlnamestrunc---Baby girl names, truncated:}
  Full series of allotaxonographs 
  corresponding to histograms of
  row 3 in Fig.~7 with $\alpha=\infty$.

  \smallskip
  \textbf{Flipbook \flipbookboynamestrunc---Baby boy names, truncated:}
  Full series of allotaxonographs 
  corresponding to histograms of
  row 4 in Fig.~7 with $\alpha=\infty$.

  \smallskip
  \textbf{Flipbook \flipbookcompaniestrunc---Market caps, truncated:}
  Full series of allotaxonographs 
  corresponding to histograms of
  row 5 in Fig.~7 with $\alpha=1/3$.
\end{excerpt}

Second, and however, we concur with Reviewer 2 that the figure of thumbnail
allotaxonographs in the paper (now Fig.~10, was Fig.~7) could be improved by adding
the rank-turbulence divergence scores, which we have now done (reproduced below).
We have slightly restructured the figure to make the log spacing
even for each sequence of subsamples.

\includegraphics[width=\textwidth]{figtruncation.pdf}

We have added the following to the caption for Fig.~10:
\begin{excerpt}
  The paths of convergence towards the divergence score
  vary and may be uneven if usually monotonic,
  depending on
  the systems being compared and the choice of $\alpha$.
\end{excerpt}

We have also restructured the caption slightly to point
the reader to the flipbooks in the third sentence rather than the last.

And as the last part of our response to Reviewer 2's comment,
we have added the following to the main text at the end
of Section III F. ``Truncation Effects for Rank-Based Allotaxonographs'':
\begin{excerpt}
  We see that the values of $\rtd{\alpha}$ for the truncation
  sequences approach the `true' value in largely monotonic, if different, ways.
  The value of $\rtd{1/3}$ for the Twitter study approaches from below,
  deceptively exhibiting a flat section up to $N$=$10^{6}$.
  The ecology example starts above and moves down towards
  the overall score of $\rtd{0}$=0.077.
  Baby names and markets caps similarly both start above their
  respective overall scores for $\rtd{\infty}$ and $\rtd{1/3}$,
  and move downwards, though their scores for strong truncation
  are close to one as they appear as disjoint systems
  (their histograms are predominantly `vee'-shaped).
  While the baby name scores drop slowly and not far
  (0.993 and 0.881 for 31 names for girls and boys
  down to 0.926 and 0.850 for all names),
  the market cap study only starts to
  gain more than the `vee' shape when $N$ is into the thousands.
  Because the market cap data comparison is a blend of large-scale turnovers
  around a relatively stable core, the drop is slow and then fast and further
  (0.931 for 31 corporations down to 0.441 for all corporations).

  Our work aside, we expect any divergence measure will likely vary as orders
  of magnitude more data is included.
  And we add that in certain circumstances,
  choosing to truncate a data set may be a well justified treatment of data.

  Finally, we note that while some form of truncation is a common measurement issue with real
  data for complex systems with many components, it is certainly not the only one.
  Exploring how other kinds of measurement errors affect rank-turbulence divergence
  would be a natural area of future work.
\end{excerpt}

\begin{reviewercomment}
  Further, how does the divergence measure work in
  the presence of unbalanced list sizes, say when comparing the word
  usage of English tweets between countries?  I noticed that some
  imbalance likely appears in the baby names example.  Does the
  skewness of the underlying histogram bias the divergence and
  resulting understanding of turbulent names?
\end{reviewercomment}

We have constructed rank-turbulence divergence
(and subsequently probability-turbulence divergence)
so that they inherently accommodate uneven systems sizes.
We don't see the measure as being a bias but rather
simply capable of revealing whatever differences there
are between the systems.
With familiarity, the shape of the histogram along with the balance bars
will give the user a rich understanding of how the systems differ
(regardless of what kind of divergence is being used).

We have added two more allotaxonographs for systems of strongly different sizes
as well as an accompanying discussion.
We chose to extend the baby name section because baby names in 1880 versus 2020
provide a good example of size disparity between two systems.

As we note earlier in response to Reviewer 1,
that the first two baby name allotaxonographs are for roughly
equal populations (1968 and 2018), and that the asymmetry of the
histogram is due to the larger set of distinct names used in 2018 relative to 1968.
The relevant excerpt (from the original manuscript):

\begin{excerpt}
  We note that the asymmetries of both histograms---their apparent
  right-side `heaviness'---are not due even in part
  to changes in overall numbers.
  Using total birth numbers (see Sec. V),
  the total number of girl names recorded in 1968 and 2018
  are comparable at 1,709,551 and 1,846,101 (7.99\% increase);
  for boys, these numbers are 1,775,997 and 1,928,871 (8.61\% increase).
  The number of year-exclusive names in the 1968 and 2018
  are strikingly different however:
  8,194 and 18,029 for girls (120\% increase),
  and 4,742 and 14,004 for boys (195\% increase).
  Two of the likely major factors which have lead to this explosion in name-space
  are immigration and a cultural shift towards parents creating novel names.
\end{excerpt}

The new figures (Figs.~6 and 7) and their captions, along with the discussion excerpted follow:

\includegraphics[width=\textwidth]{figallotaxonometer9000-babynames-rtd-010-alpha-infty-1-140-1880-vs-2020_noname.pdf}

Caption:

\begin{excerpt}
  \textbf{Allotaxonograph comparing US baby girl names for the years 1880 and 2020.}
  This figure is in part
  a demonstration of how allotaxonographs
  competently perform when the sizes of two systems differ strongly.
  For 1968 and 2018 in Fig.~5,
  the balance of total baby girl names 
  is an almost even 49.2\% and 50.8\%.
  By contrast, for 1880 and 2020, these percentages are 5.4\% versus 94.6\%.
  The choice of $\alpha=\infty$ again means that the top names for each year will
  dominate $\rtd{\infty}$, regardless of their rank in the comparison year
  (unless a name has equal rank in both years).
\end{excerpt}

\includegraphics[width=\textwidth]{figallotaxonometer9000-babynames-rtd-010-alpha-infty-2-140-1880-vs-2020_noname.pdf}

Caption:

\begin{excerpt}
  \textbf{Allotaxonograph comparing US baby boy names for the years 1880 and 2020,
    companion to Fig~6.}
  The year 1880 has 6.0\% of the total baby boy names of both years combined,
  while 2020 has 94.0\%.
  For 1968 and 2018 in Fig.~6,
  the equivalent numbers are 49.0\% and 51.0\%.
\end{excerpt}

Added discussion.

\begin{excerpt}
  To close out our study of baby names, we add two more allotaxonographs
  whose primary purpose is to show how our instrument performs when system sizes differ strongly.
  In Figs.~6 and 7,
  we compare US baby girl and boy names in 1880 and 2020, a 140 year gap.

  We make some observations about balances,
  the rank-turbulence divergence scores,
  the rank-rank histograms,
  and
  the changes in naming from 1880 to 2020.

  For the preceding allotaxonographs
  (Figs.~1--5),
  the largest difference for system sizes has been for
  Twitter in Fig.~1.
  The date 2016/11/09 carried 59.9\% of all tweets from the two dates combined,
  with the other 41.1\% on 2017/08/18
  (top balance bar, bottom right of the histogram).

  By contrast, of the total number of baby girls born in 1880 and 2020,
  the years separately account for 5.4\% and 94.6\% respectively,
  a factor of roughly 17-fold
  (new Fig.~6)
  For boys, these weights are similar at 6.0\% and 94.0\%, around 16-fold
  (new Fig.~7).

  Because of the large increase in registered babies being born,
  the two kinds of type balances are consequently more extreme.
  For example, of the combined types for the distinct baby girl names in 1880 and 2020,
  only 5.3\% were used in 1880, while 98.7\% were used in 2020.
  For exclusive types, 25.3\% of 1880's distinct baby girl names appeared
  only in 1880, while 96.0\% of 2020's were not used in 1880.

  For baby girl names, the value of $\rtd{\infty}$=9.31 for this 140 year comparison is slightly higher
  than that for the 50 year gap between 1968 and 2018, $\rtd{\infty}$=9.26,
  while for boys the increase is from $\rtd{\infty}$=0.850 to 0.900.

  In general, the rank-rank histograms of these disparately sized systems will show a strongly separated,
  highly dense line corresponding to exclusive types on the side of the large system.
  For both baby girl and boy names in 1880 and 2020,
  the separated line is around an order of magnitude from the main body of the histogram,
  and the component cells are high count ones.
  While this separation could occur for equal-sized systems if the type counts differ enough,
  the count density of the separated line will not be as strong.
  With familiarity, a glance at the balance bars will clarify these details.

  Rank-turbulence divergence with $\alpha$=$\infty$ 
  is a function only of the highest rank for each type
  (Eq.~11).
  As such, the main contributions for girls come from
  `Mary' ($\sizerank$=1 to 123)
  `Olivia' ($\sizerank$=234.5 to 1),
  while for boys
  the leaders are
  `John' ($\sizerank$=1 to 27)
  and
  `Liam' ($\sizerank$=7,643.5 to 1, not used in 1880).

  For evolving complex systems, allotaxonographs can help lead us to 
  examine time series for individual types that occupy interesting locations
  in the rank-rank histogram.
  For baby girl names, `Emma' stands out as a name that was
  enormously popular in both
  1880 ($\sizerank$=3)
  and
  2020 ($\sizerank$=2).
  But the story for Emma proves to be akin to that of Vonnegut's man-in-a-hole's
  emotional arc [64--66].
  Ranked third in 1880, `Emma' dropped at a gradually increasing rate over the next 90 years
  to a stable set of low ranks in the 1970s---the decade of `Jennifer'---bottoming out
  at $\sizerank$=463 in 1976.
  After first starting to revive in 1983,
  `Emma' rapidly rose back to 4th in 2002 and
  stayed in the top 3 from 2003--2021,
  six times atop with $\sizerank$=1.
\end{excerpt}

\begin{reviewercomment}
  The figures function as standalone illustrations of the concepts and
  enhance the intuition underlying the measure.  I especially like
  Figure 2, which manages to compress an impressive amount of
  information into a clear visualization.

  Furthermore, the
  Supplemental Flipbooks are fantastic and I hope the publication of
  this article is accompanied by a Twitter campaign sharing these as
  gifs. I'm already excited to explore the allotaxonomy of my own
  research questions!
\end{reviewercomment}

Thank you!

As we noted in the paper, we have an online site for the papers, data, and code that we have produced
(and will produce) around allotaxonometry. There are various gifs on display!
We will make more!

The top of the code page, for example, shows a video sliding through for
season points totals for NBA players, year to year:
\url{https://compstorylab.org/allotaxonometry/code/}

Our second paper on allotaxonometry (which introduces probability-turbulence divergence)
is also available there.
While conceived as an analog of RTD,
we show that PTD has a range of very interesting properties of its own.
We also show that PTD unifies and generalizes a wide array of existing probability-based
divergences.

The main site is here:
\url{https://storylab.w3.uvm.edu/allotaxonometry/}

\reviewerheader{3}

\begin{reviewercomment}
  Reviewer \#3: This is overall an excellent piece of research. The authors develop an innovative, solid method to compare rankings in systems, and demonstrate its use for word frequencies in social media. The methods are solid, the writing of very quality and the figures as well. It is without hesitation that I recommend this paper for publication as is.
\end{reviewercomment}

We greatly, greatly appreciate Reviewer 3's extremely positive assessment. Thank you!

### 2023-03-16 correspondence/2022-05-31response-epj-data-science.outtakes.tex
Greatly improved response-to-reviewers template with example; restructured cover letter naming

--- 2023-03-07 correspondence/2022-05-31epj-data-science-response.body.tex
@@
-\section*{Reply to Reviews}
-
-\textbf{Submission:} EPDS-D-20-00186 \\
-\textbf{Manuscript:} 
-Allotaxonometry and rank-turbulence divergence: A universal instrument for comparing complex systems
-\textbf{Authors:} Dodds et al.
-
-\generalcomment{Assessment from the Associate Editor:}{
-  The authors propose a very interesting method. We're sorry this review
-  took a long time, but eventually we were able to secure three highly
-  competent Reviewers. Reviewer 1 presents a few major reviewercomments
-  and recommended a major revision. Upon closer inspection, however,
-  those reviewercomments are mostly interpretational, but addressing
-  them will make the manuscript more readily accessible to a broader
-  readership. The reviewercomments of R1 are in line with the
-  reviewercomments from Reviewer 2, who also suggests the authors make a
-  few minor clarifications. Reviewer 3 raises no criticism. Taken
-  together, all reviews are very positive and only minor clarifications
-  are requested.
-}
-
-  We greatly thank the editor for the efforts and their overall positive evaluation,
-  and we thank the reviewers for their helpful and enthusiastic feedback.
-
-We have revised the manuscript to address all concerns, with individual responses below.
-We have also generally improved the manuscript with some minor edits throughout,
-as well as by marking up the allotaxonographs in Figures 1 and 2 to better facilitate our
-explanation of their elements.
-We have also adjusted some terminology for clarity,
-and to remove unneeded attributions to individual scientists.
-For example, we have largely replaced ``Zipf distributions'' with
-the more
-plainspoken ``type rankings''.
-
-We also apologize for the delay in our response.
-
-\reviewerheader{1}
-
-\reviewercomment{
-Reviewer \#1: Thank you for the opportunity to review this paper.
-It proposes new methods to compare either multiple systems or a single system over time, including graphs and a measure of divergence.
-The proposed methods are demonstrated on a variety of data examples.
-\\
-
-Major Reviewercomments:\\
-
-Can we use the rank-rank histogram to detect differences in the number of component types between the two systems?
-For example, can we use the histogram to figure out whether there are more species in 1985 vs.
-2015 in Figure 3?
-}
-
-As was presented in the manuscript, the total number of types was not made explicit.
-
-\todo{Point to balances}
-
-\todo{Add some text to paper about this.}
-
-\todo{In future iterations, we could add the number of tokens (if counts are the measure), number of types, and number of exclusive types to the inset showing balances.}
-
-\todo{Can see the imbalance from the histogram}
-
-\todo{The numbers can be reported in text accompanying figures}
-
-\reviewercomment{
-Also, can we use the graph to gauge the difference in size between the component type of rank r in System 1 and the component type of System 2 of the same rank?
-}
-
-For RTD, we wanted to build a tool that only works with ranked data.
-While in practice, we will have systems where sizes are typically known, the intent here
-was to build a fully ranked-based instrument in the spirit of well established
-rank-based statistical methods (e.g., the Spearman coefficient).
-
-In our second paper on allotaxonometry, we introduce and explore probability-turbulence divergence,
-where we make size (as measured by probability or rate) explicit.
-
-We note that the allotaxonometer code does include the ability to compare systems where we have sizes that are not counts or probabilities or cannot be interpreted as relative fractions
-in some way (e.g., heights of buildings).
-
-This manuscript is also in review and can be found here: \url{https://arxiv.org/abs/2008.13078}
-
 \todo{Zipf distributions comparing cities in 1750 to 2000 in the US}
@@
 
-\todo{Best thing would be to overlay the two Zipf distributions.
-We considered ways of adding Zipf distributions to the plots but we felt that the results would be too complicated (and we acknowledge that the allotaxonographs are complicated enough as is)}
+\todo{Bad idea: Joint distribution means that they won't be sensible
+  Zipf distributions with the 0 point. Plus binning.
+  Best thing would be to overlay the two Zipf distributions. We
+  considered ways of adding Zipf distributions to the plots but we
+  felt that the results would be too complicated (and we acknowledge
+  that the allotaxonographs are complicated enough as is)}
 
-\reviewercomment{
-How do we interpret the percentages in the "Balances" section at the bottom of Figure 1? Are these percentages unrelated to your alpha tuning parameter? 
-}
-
-The balances are independent of alpha, and indicate relative percentages
-of tokens, types, and exclusive types.
-\todo{We have added some text to make this more clear}
-and
-\todo{we have also expanded the balance section to included absolute numbers}.
-
-\reviewercomment{
-The paper states that the standard methods lack transparency and flexibility.
-It may be helpful to apply some standard methods in one of your case studies to demonstrate the differences from your method.
-}
-
-Yes, this is well spotted. We have separated out work where we compare
-methods (e.g., JSD vs PTD) and that is still in development, and had left
-residual guidance in the abstract.
-
-We have removed two sentences from the abstract that misleadingly pointed in this direction.
-
-\reviewercomment{
-How sensitive is your method to small perturbations?
-For example, if you fixed the errors in the Berkshire Hathaway and DowDuPont market caps in the last case study, how much do the divergence and histogram change?
-}
-
-In this first paper, we have explored the effects of truncation only.
-Generally, using ranks will give a more robust instrument, and is part of our motivation.
-
-We have now noted in the manuscript the following:
-
-"Small errors or perturbations of counts for a Zipf distribution 
-will generally not affect the ranks of the dominant types."
-
-\reviewercomment{
-What is the computing time to calculate the divergence and generate the histograms in your four case studies?
-}
-
-Generally, the calculations and graphics require a few seconds of compute time.
-The computational bottleneck in producing the RTD calculations is confined entirely to the counting of elements to produce ranks.
-For the Twitter data, for example, enumerating $n$-grams for each day (20GB of compressed text) can take many CPUs running continuously to keep up with real time, with more than one million previously unseen types encountered on a given day.
-
-\reviewercomment{
-
-Minor Reviewercomments:
-
-The terms "Zipf" and "1-gram" should be defined.
-}
-
-We have made the initial reference to Zipf more clear, connecting only to "Zipf's law".
-We have now defined Zipf distribution and $n$-grams in the manuscript. 
-We have replaced ``Zipf ranking'' with ``size ranking'' throughout, as we do not need
-to name this simple kind of ranking after an individual.
-
-Added to the manuscript:
-`For ease of reference, we will refer to
-``component rankings by decreasing size'' as ``size rankings.'''
-
-Definition/clarification of 1-grams now appears in a couple of places, e.g.:
-
-`For Twitter, subsampling $n$-grams---phrases containing $n$ words---allows for robust estimation of the rates of common $n$-grams but not rare ones.'
-
-\reviewercomment{
-In Paragraph 2 of Part B in the Introduction, it would be helpful to get a concrete example of a component type.
-}
-
-\todo{Do}
-
-\reviewercomment{
-
-Define the notation $r_(\tau,1)$ and $r_(\tau,2)$ , which seems to be first used in Part B of Section II. I believe they are the ranks of component τ in Systems 1 and 2, but this should be clarified.
-}
-
-\todo{Do}
-
-\reviewercomment{
-In Figure 1, the caption has substantial overlap with the main text. Removing some of the overlap would help highlight the new information.
-}
-
-It is our strong preference to make the captions reasonably complete in themselves. Readers may choose to work through the paper in different ways.
-
-\reviewercomment{
-In the first two paragraphs on page 6, some of the words that are described (such as "Harambe" and "voted") do not seem to be in Figure 1.
-}
-
-Thanks for pointing this out, we have adjusted the language describing Figure 1 to match words found at the edges of the histogram.
-`Gorilla' appears in the figure, so we have retained description of Harambe for context in the main text.
-
-\todo{double check that remade figure contains words referenced in the caption}
-
-\reviewercomment{
-Could you give some intuition of why you chose $\alpha = 0$ for the BCI example? Also, what is the $D_{0;rand}^R$ term mentioned in that example?
-}
-
-We make the choice of $\alpha = 0$ by inspection.
-In the appropriate supplementary PDF flipbook, the effect of tuning alpha can be observed.
-Future work, which we discuss at the end of the paper, would seek to find
-a method to determine an optimal alpha for specific comparisons.
-
-% \rtdalphavarsystemsRankRandw$
-
-\todo{Connect rand properly back to normalization and notation definition, or wherever}
-
-Note that we have made the first appearance of our notation for RTD into a standalone equation (the first equation, shifting following equations up).
-
-\reviewerheader{2}
-
-\reviewercomment{
-It was my pleasure to review this article introducing a new framework to compare heavy-tailed ranked lists.
-Specifically, the work motivates and derives a novel divergence measure for ranked lists, and then illustrates its usefulness across four examples including: Twitter word frequency, baby name popularity, species abundance, and firm sizes.
-The presentation is rich and clear, elucidating the central framework and offering wonderful insights into the motivating examples and applications.
-The problem this article addresses is relevant to a broad range of data science applications.
-Therefore, I strongly recommend to publish the article in EPJ Data Science.
-}
-
-We thank Reviewer 2 for their generous feedback.
-
-\reviewercomment{
-Reviewercomments:
-Even through the presentation is quite clear, it may benefit from an organization that extracts a few non-technical discussions from the rest of the text - i.e.
-a clearly identified general audience guide to allotaxonomy and the application of rank-turbulence.
-This could support usage in fields outside of physics / data sciences.
-}
-
-Thank you.
-
-We are authoring a blog post to accompany publication of the
-manuscript that will advertise the utility of the instrument to other
-disciplines.
-
-We are also working on julia and python versions of the
-allotaxonometer code (presently 6000 lines of Matlab) to encourage
-broader adoption.
-
-We and others have also used allotaxonographs in a number of other
-papers, and we now cite these in the manuscript's conclusion.
-
-%% nWe and others have also used allotaxonographs in a number of other papers, and we now cite these in the manuscript
-%% \cite{gothard2021a,stupinksi2021a,minot2021a,ring2020a,alshaabi2021c,dodds2020n}
-
-\reviewercomment{
-
-Relatedly, it would be helpful if the authors could elaborate on
-section IIE and the processes used to select different values of alpha
-for the different applications.  I'm imagining the framework being
-used by a non-expert who just keeps whatever default value of alpha is
-encoded instead of exploring the full range - or on the flip side,
-being overwhelmed with the many potential choices of alpha and not
-understanding what features of the histograms are being accentuated by
-the divergence in different limits. Can you distill a few guiding
-principles or questions to help motivate an appropriate selection?
-}
-
-In the present work, we justify the choice of alpha by looking at the shape of distribution as a guide.
-
-In the appropriate supplementary PDF flipbook, the effect of tuning alpha can be observed.
-
-In future work, which we discuss at the end of the paper, we will seek
-to find a method to determine an optimal alpha for specific
-comparisons.
-
-\reviewercomment{
-The examples illustrating the effects of partial sampling were helpful, but it would be nice to see how the divergence measure itself changes with subsample size: does it converge or is it sensitive to the influx of different cross-ranks uncovered by larger subsample sizes?
-Further, how does the divergence measure work in the presence of unbalanced list sizes, say when comparing the word usage of English tweets between countries?
-I noticed that some imbalance likely appears in the baby names example.
-Does the skewness of the underlying histogram bias the divergence and resulting understanding of turbulent names?
-}
-
-\todo{Performance under sampling: should obey whatever decay rate rank-ordered data of particular scaling law experiences, independent of this methodology.}
-
-\todo{We've seen undersampled tweet data, looks too sparse near center, too many hapax words appearing in bar chart on right.}
-
-\reviewercomment{
-  The figures function as standalone illustrations of the concepts and enhance the intuition underlying the measure.
-I especially like Figure 2, which manages to compress an impressive amount of information into a clear visualization.
-Furthermore, the Supplemental Flipbooks are fantastic and I hope the publication of this article is accompanied by a Twitter campaign sharing these as gifs.
-I'm already excited to explore the allotaxonomy of my own research questions!
-}
-
-Thank you!
-
-As we noted in the paper, we have an online site for the papers, data, and code that we have produced
-(and will produce) around allotaxonometry. There are various gifs on display:
-
-\url{https://storylab.w3.uvm.edu/allotaxonometry/}
-
-\reviewerheader{3}
-
-\reviewercomment{
-  Reviewer \#3: This is overall an excellent piece of research.
-The authors develop an innovative, solid method to compare rankings in systems, and demonstrate its use for word frequencies in social media.
-The methods are solid, the writing of very quality and the figures as well.
-It is without hesitation that I recommend this paper for publication as is.
-}
-
-We greatly, greatly appreciate Reviewer 3's extremely positive assessment. Thank you!
+\noway{we have also expanded the balance section to included absolute numbers}.

### 2023-07-12 localizify.pl
Getting this straighted out

@@ -52,8 +52,12 @@ foreach $line (@lines) {
 # but then you would have failed as a human anyway.
 foreach $i (0..$#figures) {
     $figure = $figures[$i];
-	    
-    $filenames = `find figures -follow -name $figure -print 2>/dev/null`;
+
+    $filenames = `find figures -type d -name archive -prune -follow -o -name $figure -print 2>/dev/null`;
+
+    ## uses prune per here:
+    ## https://stackoverflow.com/questions/4210042/how-do-i-exclude-a-directory-when-using-find
+
     @filenames = split("\n",$filenames);
     @filenames = grep { $_ !~ m/$localfiguresdir/ } @filenames;
     

### 2023-07-12 make-single-latex-file.pl

@@ -35,15 +35,9 @@ open (FILE,"$file") or die "can't open $file: $!\n";
 open (OUTFILE,">$outfile") or die "can't open $outfile: $!\n";
 
 undef $tex;
-foreach $line (<FILE>) {
-    # Clean out comments
-    # For full comments, will remove entire lines, otherwise to end of line
-    if ($line =~ m/^\s*%%.*$/) {
-	$line = "";
-    }
-    
+foreach $line (<FILE>)
+{
     $line =~ s/%%.*$//;
-    
     if ($line =~ m/\\currfile/) {
 	$line = "";
     }
@@ -53,23 +47,14 @@ foreach $line (<FILE>) {
     if ($line =~ m/\\input/)
     {
 	$line =~ m/\\input\{(.*?)\}/;
-	## print "$1\n";
+#	print "$1\n";
 
 	$inputfile = $1;
-	unless ($inputfile =~ m/\.tex$/ or $inputfile =~ m/\.txt$/) { # could be txt or tex
+	unless ($inputfile =~ m/\.tex$/) {
 	    $inputfile = $inputfile.".tex";
 	}
 
-	if (-e "$inputfile") {
-	    $inputfile_fullpath = $inputfile;
-	    ## print "beeply $inputfile_fullpath\n";
-	} else { # otherwise go find it:
-	    $tmp = `find . -follow -name $inputfile -print 2>/dev/null`;
-	    @tmp = split("\n",$tmp);
-	    $inputfile_fullpath = $tmp[0];
-	}
-	
-	open (INPUT,"$inputfile_fullpath") or die "can't open $inputfile_fullpath: $!\n";
+	open (INPUT,"$inputfile") or die "can't open $inputfile: $!\n";
 	undef $/;
 	$input = <INPUT>;
 	$/ = "\n";
@@ -78,40 +63,32 @@ foreach $line (<FILE>) {
 	# special treatment for $realbasefile.title.tex
 	if ($line =~ m/\.title/) {
 	    $input =~ s/%.*?$//msg;
-	    $input =~ s/^\s*?//;
-	    $input =~ s/\s*?$//;
+	    $input =~ s/^\s*//;
+	    $input =~ s/\s*$//;
 	    chomp($input);
 	}
 
 	$line =~ s/\\input\{(.*?)\}/$input/;
 
-	# one extra layer, process all input commands present
+# one extra layer
 	while ($line =~ m/\\input\{(.*?)\}/) {
[…]

### 2023-07-12 make-zip-file-for-overleaf.sh

@@ -2,7 +2,7 @@ zip \
     -FS \
     --exclude=make-zip-file* \
     --exclude=*.zip \
-    overleaf.zip \
+    `basename $PWD`-overleaf.zip \
     * \
     correspondence/* \
     correspondence/figures/* \

### 2025-01-27 correspondence/2022-05-31response-epj-data-science.body.tex
Solid update; moving to manuscipt format

@@
 \begin{center}
-  \includegraphics[width=0.3\textwidth]{figallotax-baby-girl-name-balances.pdf}
+  \includegraphics[width=0.3\textwidth]{figures/localized/figallotax-baby-girl-name-balances.pdf}
 \end{center}
@@
 \begin{center}
-  \includegraphics[width=0.4\columnwidth]{allotaxonograph_balances2_flattened-crop.pdf}
+  \includegraphics[width=0.4\columnwidth]{figures/localized/allotaxonograph_balances2_flattened-crop.pdf}
 \end{center}
@@
 
-\includegraphics[width=\textwidth]{figallotaxonometer9000-siblis_alphasweep_05-2007-Q4_2018-Q4_marketcaps004_noname.pdf}
-
-\includegraphics[width=\textwidth]{figallotaxonometer9000-siblis_alphasweep_06-2007-Q4_2018-Q4_marketcaps004_corrected_noname.pdf}
+\includegraphics[width=\textwidth]{figures/localized/figallotaxonometer9000-siblis_alphasweep_05-2007-Q4_2018-Q4_marketcaps004_noname.pdf}
+
+\includegraphics[width=\textwidth]{figures/localized/figallotaxonometer9000-siblis_alphasweep_06-2007-Q4_2018-Q4_marketcaps004_corrected_noname.pdf}
 
@@
 
-\includegraphics[width=\textwidth]{figtruncation.pdf}
+\includegraphics[width=\textwidth]{figures/localized/figtruncation.pdf}
 
@@
 
-\includegraphics[width=\textwidth]{figallotaxonometer9000-babynames-rtd-010-alpha-infty-1-140-1880-vs-2020_noname.pdf}
+\includegraphics[width=\textwidth]{figures/localized/figallotaxonometer9000-babynames-rtd-010-alpha-infty-1-140-1880-vs-2020_noname.pdf}
 
@@
 
-\includegraphics[width=\textwidth]{figallotaxonometer9000-babynames-rtd-010-alpha-infty-2-140-1880-vs-2020_noname.pdf}
+\includegraphics[width=\textwidth]{figures/localized/figallotaxonometer9000-babynames-rtd-010-alpha-infty-2-140-1880-vs-2020_noname.pdf}
 

### 2025-01-27 correspondence/localizify.pl

@@ -1,0 +1,190 @@
+#!/usr/bin/perl
+
+## adapted to the PoCSverse
+
+#############################################
+## localizes figures and text snippet inputs
+#############################################
+
+@files = @ARGV;
+
+$usage = "
+usage: localizify.pl foo1.tex [foo2.tex foo3.tex] ...
+
+does two things:
+
+1. finds and localizes all figures found in input files
+
+- searches for all matches to \includegraphics
+- also searches for
+ - everything in pocsverse-forced-figures.tex
+ - \changelecturelogo
+- will not search if file exists in figures/localized-links
+- finds first match in figures tree
+  and stores in figures/localized
+- ignores figures/localized and figures/localized-links
+- ignores figure names with variables/commands in them (force include these)
+- creates symbolic link in figures/localized-links
+- copies everything in figures/localized-links to figures/localized
+
+2. finds and localizes all input text files in *.settings.tex
+
+- signal to do this by explicitly using \input{inputs/localized/...}
+- stores in inputs/localized/
+- expects snippets to be in ./figures tree only
+
+";
+
+##############################
+## figures
+##############################
+
+$localfiguresdir = "figures/localized";
+`mkdir -p $localfiguresdir`;
+
+$localfigureslinkdir = "figures/localized-links";
+`mkdir -p $localfigureslinkdir`;
+
+$localinputsdir = "inputs/localized";
+`mkdir -p $localinputsdir`;
+
+$localinputslinkdir = "inputs/localized-links";
+`mkdir -p $localinputslinkdir`;
+
+## figures
+$text = `grep -h "figures/localized" @files 2>/dev/null`;
+
+## print $text;
+
+@lines = split('\n',$text);
+foreach $line (@lines) {
+
+    if ($line =~ m/^\s*%.*$/) {
+        $line = "";
+    }
+
+##    print $line."\n";
+
+    ## simpler, better: just look for figures/localized
+    if ($line =~ m/\{figures\/localized\/(.*?)\}/) {
+ 	$figfile = $1;
+ 	unless ($figfile =~ m/\\/) {
+ 	    push @figures, $figfile;
+ 	}
+     }
+    
[…]

### 2025-01-27 correspondence/make-2022-05-31response-epj-data-science.sh

@@ -1,0 +1,3 @@
+localizify.pl 2022-05-31response-epj-data-science.body.tex
+pdflatex 2022-05-31response-epj-data-science
+pdflatex 2022-05-31response-epj-data-science

### 2025-01-27 localizify.pl

@@ -1,23 +1,35 @@
 #!/usr/bin/perl
 
+## adapted to the PoCSverse
+
 #############################################
 ## localizes figures and text snippet inputs
 #############################################
 
+if ($#ARGV<0) {
+    exit;
+}
+
+@files = @ARGV;
+
 $usage = "
-usage: localizify.pl
+usage: localizify.pl foo1.tex [foo2.tex foo3.tex] ...
 
 does two things:
 
-1. finds and localizes all figures found in:
+1. finds and localizes all figures found in input files
 
-*.body.tex *.supplementary.tex *.appendix.tex
-
-- searches for all matches to includegraphics
+- searches for all matches to \includegraphics
+- also searches for
+ - everything in pocsverse-forced-figures.tex
+ - \changelecturelogo
+- will not search if file exists in figures/localized-links
 - finds first match in figures tree
   and stores in figures/localized
-- ignores figures/localized
-- overwrites figures
+- ignores figures/localized and figures/localized-links
+- ignores figure names with variables/commands in them (force include these)
+- creates symbolic link in figures/localized-links
+- copies everything in figures/localized-links to figures/localized
 
 2. finds and localizes all input text files in *.settings.tex
 
@@ -34,18 +46,51 @@ does two things:
 $localfiguresdir = "figures/localized";
 `mkdir -p $localfiguresdir`;
 
+$localfigureslinkdir = "figures/localized-links";
+`mkdir -p $localfigureslinkdir`;
+
 $localinputsdir = "inputs/localized";
 `mkdir -p $localinputsdir`;
 
-$text = `grep includegraphics *.body.tex *.supplementary.tex *.appendix.tex 2>/dev/null`;
+$localinputslinkdir = "inputs/localized-links";
+`mkdir -p $localinputslinkdir`;
+
+## figures
+$text = `grep -h "figures/localized" @files 2>/dev/null`;
+
+## print $text;
 
 @lines = split('\n',$text);
 foreach $line (@lines) {
-    $line =~ s/\%.*$//;
 
-    if ($line =~ m/includegraphics\[.*?\]\{$localfiguresdir\/(.*?)\}/) {
-	push @figures, $1;
[…]

### 2025-01-27 make-single-latex-file.pl

@@ -124,6 +124,9 @@ $tex =~ s/\n\n+/\n\n/msg;
 $tex =~ s/^\s+//;
 $tex =~ s/\s+$/\n/;
 
+# remove figures/localized/ from includegraphics
+$tex =~ s/figures\/localized\///msg;
+
 print OUTFILE $tex;
 
 close FILE;

### 2025-01-27 makemanuscript.pl

@@ -1,0 +1,166 @@
+#!/usr/bin/perl
+
+## set up name in settings file
+`make-name-match-settingsfile.pl`;
+
+## localize figures and inputs (for better sharing and also overcomes legacy figure search slowness)
+`./localizify.pl`;
+
+## local directory (for global recording of writing data)
+## works because paper/book directory names are in the format YYYY-MM-descriptive-tags
+## downfall: directory names may change but that represents a rebuilding
+$fulldir = $ENV{'PWD'};
+($localdir = $ENV{'PWD'}) =~ s/^.*\///;
+
+foreach $argument (@ARGV) {
+    if (($argument eq "-quick") or ($argument eq "-q"))
+    {
+	$quickswitch = 1;
+    } elsif ($argument =~ m/\.body\.tex$/) {
+	@bodies = (@bodies, $argument);
+    }
+}
+
+if ($#bodies eq -1) {
+    $bodylist = `ls *.body.tex`;
+    @bodies = split(/\s+/,$bodylist);
+}
+
+($sec,$min,$hour,$numday,$month,$year,$wday,$yday,$isdst) = localtime(time);
+$monthnum = $month + 1;
+$year += 1900;
+$timestamp = "$year $monthnum $numday $hour $min $sec";
+
+foreach $body (@bodies) {
+    ($filename = $body) =~ s/\.body\.tex$//;
+    ($supplementaryfilename = $body) =~ s/\.body\.tex$/\.supplementary\.tex/;
+    ($appendiexfilename = $body) =~ s/\.body\.tex$/\.appendix\.tex/;
+    if (-e "$filename-manuscript.tex") {
+
+	## make local bibliography
+	## will need to be adjusted for more elaborate inputs
+	$texfiles = join(" ",<$filename.*.tex>);
+	$command = "make-local-bibfile.pl $filename.bib $texfiles";
+        system($command);
+
+	## make single line versions of title, abstract, logline
+	$command = "./singlelineify.pl $filename.title.tex\n";
+        system($command);
+	$command = "./singlelineify.pl $filename.abs.tex\n";
+        system($command);
+	$command = "./singlelineify.pl $filename.logline.tex\n";
+        system($command);
+
[…]

### 2025-06-26 paper-template-newpax.tex
General updates

\documentclass{article}

%% load the lua code
\directlua{require("newpax")}

%% write .newpax files for newpax.sty

%% for references to the main data set
\newcommand{\Ncharacters}{2000}
\newcommand{\Ntraits}{464}
\newcommand{\Nstories}{341}

%%   newpax.writenewpax("figures/localized/Arrested-Development-Lucille-Bluth-\Ncharacters-\Ntraits-\Nstories")
%%   newpax.writenewpax("figures/localized/Buffy-the-Vampire-Slayer-Buffy-Summers-\Ncharacters-\Ntraits-\Nstories")
%%   newpax.writenewpax("figures/localized/rational--whimsical-\Ncharacters-\Ntraits-\Nstories")

\directlua
    {
      newpax.writenewpax("figures/localized/stories-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/traits-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/archetypes-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/essential-traits-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/orderly--chaotic-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Its-Always-Sunny-in-Philadelphia-Charlie-Kelly-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Buffy-the-Vampire-Slayer-Buffy-Summers-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Harry-Potter-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/Harry-Potter-dimensions-3-vs-4-\Ncharacters-\Ntraits-\Nstories")
      newpax.writenewpax("figures/localized/comparison-Archer-Malory-Archer-vs-Arrested-Development-Lucille-Bluth-2000-464-341")
      newpax.writenewpax("figures/localized/comparison-Archer-Sterling-Archer-vs-Bobs-Burgers-Bob-Belcher-2000-464-341")
    }

\begin{document}
\end{document}

%% - these don't work properly because the cropped pdfs include all of the original
%% - moving to pngs
%% 
%%       newpax.writenewpax("figures/localized/orderly--chaotic-\Ncharacters-\Ntraits-\Nstories")
%%       newpax.writenewpax("figures/localized/orderly--chaotic-\Ncharacters-\Ntraits-\Nstories-title")
%%       newpax.writenewpax("figures/localized/Buffy-the-Vampire-Slayer-Buffy-Summers-\Ncharacters-\Ntraits-\Nstories-title")
%%       newpax.writenewpax("figures/localized/outcast-angel-hero-header")
%%       newpax.writenewpax("figures/localized/outcast-angel-hero-switcher")
%%       newpax.writenewpax("figures/localized/essential-trait-007-positive-header")

### 2025-06-26 log/convert-month-numbers.pl

@@ -1,0 +1,39 @@
+#!/usr/bin/perl
+
+## adjust month numbers before the start of 2025
+
+@files = <pre-corrected-2025/progress-*-log>;
+
+`mkdir -p pre-corrected-2025`;
+`mkdir -p corrected-2025`;
+
+foreach $file (@files) {
+
+    print $file."\n";
+
+    ($outfile = $file) =~ s/pre-corrected/corrected/;
+
+    ##    print $file."\n";
+    ##    print $outfile."\n";
+
+    open (FILE,"$file") or die "Can't open $file: $!\n";
+    open (OUTFILE,">$outfile") or die "Can't open $outfile: $!\n";
+    
+    foreach $line (<FILE>) {
+
+	$line =~ m/^(.*?) (.*?) (.*)$/;
+	$year = $1;
+	if ($year < 2025) {
+	    $monthnum = $2 + 1;
+	} else {
+	    $monthnum = $2;
+	}
+	$newline = $year." ".$monthnum." ".$3."\n";
+
+	print OUTFILE $newline;
+    }
+
+    close FILE;
+    close OUTFILE;
+}
+    

### 2025-06-26 makemanuscript.pl

@@ -1,5 +1,7 @@
 #!/usr/bin/perl
 
+$filetype = "manuscript";
+
 ## set up name in settings file
 `make-name-match-settingsfile.pl`;
 
@@ -43,13 +45,11 @@ foreach $body (@bodies) {
 	$command = "make-local-bibfile.pl $filename.bib $texfiles";
         system($command);
 
-	## make single line versions of title, abstract, logline
-	$command = "./singlelineify.pl $filename.title.tex\n";
-        system($command);
-	$command = "./singlelineify.pl $filename.abs.tex\n";
-        system($command);
-	$command = "./singlelineify.pl $filename.logline.tex\n";
-        system($command);
+	## onelineify title, logline, and abstract
+	`singlelineify.pl $filename.title.tex`;
+	`singlelineify.pl $filename.logline.tex`;
+	`singlelineify.pl $filename.abs.tex`;
+	`singlelineify.pl $filename.acknowledgments.tex`;
 
 	($logfilename = "log/progress-$body") =~ s/\.body\.tex$/-log/;
 	print "processing $filename...\n";
@@ -73,7 +73,7 @@ foreach $body (@bodies) {
 	}
 	    
 	if ($pdfswitch == 1) {
-	    `lualatex $filename-manuscript 1>&2`;
+	    `pdflatex $filename-manuscript 1>&2`;
 	}
 	else {
 	    `latex $filename-manuscript 1>&2`;
@@ -81,8 +81,8 @@ foreach $body (@bodies) {
 	unless ($quickswitch == 1) {
 	    `bibtex $filename-manuscript 1>&2`;
 	    if ($pdfswitch == 1) {
-		`lualatex $filename-manuscript 1>&2`;
-		`lualatex $filename-manuscript 1>&2`;
+		`pdflatex $filename-manuscript 1>&2`;
+		`pdflatex $filename-manuscript 1>&2`;
 	    }
 	    else {
 		`latex $filename-manuscript 1>&2`;
@@ -139,7 +139,6 @@ foreach $body (@bodies) {
 	    `touch $centrallogfile`;
 	    open (CENTRALLOGFILE,">>$centrallogfile") or die "Can't open $centrallogfile: $!\n";
 
-	    $filetype = "manuscript";
 	    print CENTRALLOGFILE "$timestamp $numtodos $numpages $numbytes $numlines $numwords $numchars $fulldir $filename $filetype \n";
 	    close CENTRALLOGFILE;
 	    
[…]

### 2025-06-26 makerevtex4.pl

@@ -1,5 +1,7 @@
 #!/usr/bin/perl
 
+$filetype = "revtex4-article";
+
 ## set up name in settings file
 `make-name-match-settingsfile.pl`;
 
@@ -9,8 +11,8 @@
 ## local directory (for global recording of writing data)
 ## works because paper/book directory names are in the format YYYY-MM-descriptive-tags
 ## downfall: directory names may change but that represents a rebuilding
-$fulldir = `pwd`;
-($localdir = `pwd`) =~ s/^.*\///;
+$fulldir = $ENV{'PWD'};
+($localdir = $ENV{'PWD'}) =~ s/^.*\///;
 
 foreach $argument (@ARGV) {
     if (($argument eq "-quick") or ($argument eq "-q"))
@@ -27,6 +29,7 @@ if ($#bodies eq -1) {
 }
 
 ($sec,$min,$hour,$numday,$month,$year,$wday,$yday,$isdst) = localtime(time);
+$monthnum = $month + 1;
 $year += 1900;
 $timestamp = "$year $month $numday $hour $min $sec";
 
@@ -131,10 +134,17 @@ foreach $body (@bodies) {
 	chomp $numwords;
 	chomp $numchars;
 
-		print "$numlines $numwords $numchars";
-
+	print "$numlines $numwords $numchars";
 
 	if (($numpages ne "") and ($numbytes ne "")) {
+	    $centrallogfile = sprintf("$ENV{HOME}/papers/log/writing/%d-%02d.txt",$year,$monthnum);
+	    `touch $centrallogfile`;
+	    open (CENTRALLOGFILE,">>$centrallogfile") or die "Can't open $centrallogfile: $!\n";
+
+	    print CENTRALLOGFILE "$timestamp $numtodos $numpages $numbytes $numlines $numwords $numchars $fulldir $filename $filetype \n";
+	    close CENTRALLOGFILE;
+
+	    
 	    open (LOGFILE, ">>$logfilename") or die "Can't open $logfilename: $!\n";
 	    print LOGFILE "$timestamp $numtodos $numpages $numbytes $numlines $numwords $numchars\n";
 	    close LOGFILE;

### 2025-06-26 singlelineify.pl

@@ -1,11 +1,20 @@
 #!/usr/bin/perl
 
-$usage = "singlelineify.pl foo1.tex [foo2.tex ...]
+## turns input LaTeX files into single lines
+## which are then useable for submission, websites, etc.
+##
+## (have long done this by hand)
+
+## peter sheridan dodds
+## https://github.com/petersheridandodds
+## MIT License
+
+$usage = "one-lineify.pl foo1.tex [foo2.tex ...]
 
 - create foo1.txt [foo2.txt ...]
   which are single line versions of the tex versions.
 
-- removes all empty lines, newlines, all commented out sections.
+- removes all empty lines, newlines, all commented out sections, textblocks
 
 - for use with 
   .title.tex
@@ -20,23 +29,40 @@ if ($#ARGV < 0) {
     exit;
 }
 
-foreach $file (<@ARGV>) {
+foreach $file (@ARGV) {
 
     if ($file =~ m/\.tex$/) {
+
 	($outfile = $file) =~ s/\.tex$/.txt/;
 
 	undef $/;
-	open (FILE, "$file") or die "Can't open $file: $!\n";
-	$text = <FILE>;
-	$text =~ s/\\\\//msg;
-	$text =~ s/^\s*\%.*\n//msg;
-	$text =~ s/\n/ /msg;
-	$text =~ s/ +/ /g;
+	open (FILE, $file) or die "Can't open $file: $!\n";
+	$latex = <FILE>;
+	close FILE;
 
-	open (OUTFILE,">$outfile") or die "Can't open $outfile: $!\n";
-	print "Converting $file to $outfile ...\n";
-	print OUTFILE $text."\n";
+	$latex =~ s/^\s*\%.*$//msg;
+	$latex =~ s/\\begin{textblock}//msg;
+	$latex =~ s/\\end{textblock}//msg;
+	$latex =~ s/\\item//msg;
+	$latex =~ s/\\bigskip//msg;
+	$latex =~ s/\\medskip//msg;
+	$latex =~ s/\\smallskip//msg;
+	$latex =~ s/\\\\//msg;
+	$latex =~ s/^\s*\n//msg;
+	$latex =~ s/\n/ /g;
+	$latex =~ s/ +/ /g;
+	$latex =~ s/^ +//;
+	$latex =~ s/ +$//;
+
+	$latex = $latex."\n";
+
+	open (OUTFILE, ">$outfile") or die "Can't open $outfile: $!\n";
[…]
